### 2019-02-10 paper/ms.tex
Initial Commit to Repo

\documentclass[twocolumn]{aastex62}
\usepackage{times}
\usepackage{epsfig}
\usepackage{amsmath, amsthm, amssymb}
\usepackage{color}
\usepackage{microtype}
\usepackage{float}
\usepackage{natbib}
\usepackage{verbatim}
\usepackage{placeins}
\graphicspath{{./figures/}}
\usepackage{wrapfig}
\hypersetup{backref,breaklinks,colorlinks,citecolor=blue}
\usepackage[all]{hypcap}
\usepackage{definitions}

\bibliographystyle{apj}

\newcommand{\subdate}{\today}
\newcommand{\shortauth}{Barker et al.}
\newcommand{\slugcom}{Draft - \today}
%\newcommand{\slugcom}{Submitted to ApJL on \subdate}
%\slugcomment{\slugcom}

\lefthead{\sc \footnotesize \slugcom \hfill \shortauth}
\righthead{\sc \footnotesize \slugcom \hfill \shortauth}
%}

\bibliographystyle{aasjournal}
\received{\today}
\shorttitle{Discontinuous Galerkin methods with a nuclear EOS}

\begin{document}

\title{\thornado-hydro: Generalizing discontinuous galerkin methods for a nuclear
equation of state for supernova hydrodynamics}
\author{Brandon Barker}
\affiliation{Department of Physics and Astronomy, University of Tennessee Knoxville, TN 37996}

\author{Eirik Endeve}
\affiliation{Department of Physics and Astronomy, University of Tennessee Knoxville, TN 37996}
\affiliation{Computer Science and Mathematics Division, Oak Ridge National Laboratory, TN 37831}
\affiliation{Joint Institute for Computational Sciences, Oak Ridge National Laboratory, TN 37831}

%\begin{abstract}
%\end{abstract}
\begin{abstract}
\todo{Fill In Abstract}
A problem of high importance in computational astrophysics is obtaining accurate
solutions to the Euler equations of hydrodynamics. We are interested in solving
the Euler equations in the context of core collapse supernovae (CCSNe).
The toolkit for high-order neutrino-radiation hydrodynamics (\thornado)
is being developed for CCSNe simulations and related problems utilizing a spatial discretization
based on the discontinuous Galerkin (DG) method.
Because the Euler equations form a hyperbolic set of partial differential equations,
they may be linearized to advection equations by the introduction of
chracteristic variables related to the eigen-decomposition.
This use of characteristic variables also increases the efficiency of the slope
limiting process. However, the addition of the nuclear equation of state equation
to the Euler equations makes the decompisition nontrivial.
We introduce the framework for the characteristic decomposition of the Euler equations
with the inclusion of the nuclear EOS terms and
present results from some initial tests.
\end{abstract}

\keywords{supernovae: general -- hydrodynamics -- convection -- turbulence -- equation of state -- methods: numerical -- stars: massive -- stars: evolution}

\section{Introduction}
\label{Intro}
\todo{Fill in introduction. Contextualize the work.}
The core collapse supernovae (CCSNe) explosion mechanism is fundamentally three-dimensional
in nature \citep[see e.g.,][]{blondin:2006, muller:2012, oconnor:2018b}. Alongside general
reletavistic gravity, complex nuclear equations of state (EOS), and
neutrino transport, hydrodynamics must be accurately modelled,
creating a very challenging problem. Hydrodynamic instabilities are critical
in aiding the explosion, with complex phenomena such as turbulence and
convection play key roles in the CCSN mechanism
\citep{murphy:2011, murphy:2013, couch:2013, couch:2015a, radice:2016, mabanta:2018}.
The accurate and efficient modelling of the supernova hydrodynamics is crucial if the
explosion is to be realistically modelled.
For reviews of the CCSN mechanism, see \citet{bethe:1990, janka:2007, janka:2012a, janka:2016, burrows:2013, hix:2014, muller:2016, couch:2017}.

\todo{Talk about \thornado, DG methods, Euler Equations, possibly reference other codes that
employ different methods and those methods' drawbacks.}

\section{Euler Equations of Gas Dynamics in Cartesian Coordinates}
\todo{Use this form of the Euler Equations?}
The non-relativistic Euler equations of gas dynamics
\citep[see, e.g.,][for details]{leveque:2002} in cartesian coordinates are given by
the equations of conservation of mass,
\beq
  \pderiv{\rho}{t} + \pderiv{}{x_i}(\rho u_i) = 0
  \label{eq:massConservation}
\eeq
conservation of momentum,
\beq
	\pderiv{(\rho u_i)}{t} + \pderiv{}{x_i}(\rho u^2 + p) = 0
  \label{eq:momentumConservation}
\eeq
conservation of energy,
\beq
  \pderiv{E}{t} + \pderiv{}{x_i}((E+p)u_i) = 0
  \label{eq:energyConservation}
\eeq
and conservation of \textcolor{red}{electron fraction}\footnote{Is this Correct? I don't think so. What is conserved?}
\beq
  \pderiv{D}{t} + \pderiv{}{x_i}(D u_i) = 0
  \label{eq:electronConservation}
\eeq
where $\rho$ represents mass density, $u$ the fluid velocity, $p$ the fluid
pressure, $D=\rho y_e$ is the \textcolor{red}{what is this?},
and $E=e+\frac{1}{2}\rho u^2$ the total energy (internal plus kinetic).
The system of equations is closed by the EOS which links the pressure to
the quantities evolved by Equations
\eqref{eq:massConservation}-\eqref{eq:electronConservation}. The inclusion of
Equation \eqref{eq:electronConservation} is because we require a nuclear EOS.
We may rewrite Equations \eqref{eq:massConservation}-\eqref{eq:electronConservation}
is a more convenient way:

\beq
  \pderiv{\mathbf{U}}{t}
  +\pderiv{\mathbf{F}(\mathbf{U})}{x}
  =0.
\eeq
where $\mathbf{U} =(\rho,\rho u,E, D)^{T}$ is the vector of conserved quantities
and $\mathbf{F}(\mathbf{U})=(\rho u,\rho u^{2}+p,(E+p)u, uD)^{T}$
is the flux vector. We can rewrite this further:

\beq
  \pderiv{\mathbf{U}}{t}
  + \pderiv{\mathbf{U}}{x} \pderiv{\mathbf{F(\mathbf{U})}}{\mathbf{U}}
  = 0
  \label{charEq}
\eeq
Because the Euler equations for a systemer of hyperblic
partial differential equations \citep[see, e.g.,][]{leveque:1992}, we can decompose the
Jacobian of the flux vector as

\beq
  \pderiv{\mathbf{F(\mathbf{U})}}{\mathbf{U}} =
  R \Lambda R^{-1}
\eeq
where the columns of $R$ contain the right eigenvectors of the Jacobian,
the rows of $R^{-1}$ contain the left eigenvectors, and
$\Lambda$ is a diagonal matrix containing the eigenvalues of the Jacobian.
At this point, we will introduce the characteristic variable
$\mathbf{w} = R^{-1}\mathbf{U}$. Multiplying both sides of Equation
\eqref{charEq} by $R^{-1}$, we are able to linearize the system of equations to
a system of advection equations

\beq
  \pderiv{\mathbf{w}}{t} +
  \Lambda \pderiv{\mathbf{w}}{x_{i}}
  = 0
\eeq
Solutions to these advection equations are far simpler to obtain than the
previous equations. Limiting may then be applied to dampen oscillations in the
solutions for the characteristic variables $\mathbf{w}$, which are then
transformed back to the solution of the conserved variables $\mathbf{U}$
\citep[see e.g.,][for a description]{cockburn:1998, schaal:2015a}.

\todo{More discussion of limiting?}

\hrulefill

\section{Derivation of the Jacobian and Eigenvectors for the 1D Ideal Case}

\todo{Define new variables to rewrite some things in more compact forms.}

We assume that $p = p(\tau, e, n)$, where $\tau = \frac{1}{\rho}$. Let the vector of conserved variables be $\textbf{U} = \{\rho, m, E\}$,
where $m = \rho u$. The flux vector is $\textbf{F}(\textbf{U}) = \{m, m^{2} \tau + p, (E+p)m \tau\}$.

\small{
\begin{align}
	\pderiv{\mathbf{F}(\mathbf{U})}{\mathbf{U}}
	=
	\left[
		\begin{array}{ccc}
			0 & 1 & 0  \\
			-(p_{\tau}+ p_{\epsilon}E)\tau^{2} + u^{2}(p_{\epsilon}\tau -1)       & u(2-p_{\epsilon}\tau)         & p_{\epsilon}\tau       \\
			-u(\tau(-p_{\epsilon}u^{2} + p_{\tau}\tau + p_{\epsilon}E\tau) + H) & H - p_{\epsilon}u^{2}\tau  & u(1+p_{\epsilon}\tau)  \\
		\end{array}
	\right]
\end{align}
}

Here, $H = (E+p)\tau$. The eigenvalues are given by $\lambda = \{u - \tau \sqrt{p p_{\epsilon} - p_{\tau}}, u,  u + \tau \sqrt{ p p_{\epsilon} - p_{\tau}}\}$. In the ideal case where
$\tau = \rho^{-1}$, $ E = \frac{p}{\gamma -1} + \frac{\rho v^{2}}{2}$, $ p_{\epsilon} = \frac{\gamma-1}{\tau}$, and $p_{\tau} = -\frac{(\gamma - 1)e}{\tau^{2}}$,
the sound speed $c$ reduces to $\sqrt{\frac{p \gamma}{\rho}}$. \\

The eigenvectors are given by \\

\begin{align}
	\left[
		\begin{array}{c|c|c}
			1 & 1 & 1 \\
			u - c & u & u + c \\
			\frac{\tau( c( p-p_{\epsilon}u^{2} +E) +u( -p_{\tau} + p_{\epsilon}E)\tau)}{c + p_{\epsilon}u\tau} & \frac{\tau(p_{\tau} + p_{\epsilon}E)}{p_{\epsilon}} &
			\frac{\tau( -c( p-p_{\epsilon}u^{2} +E) +u( p_{\tau} + p_{\epsilon}E)\tau)}{-c + p_{\epsilon}u\tau} \\
		\end{array}
	\right]
\end{align}

which in the ideal case described above reduces to  the eigenvectors given in \citet{schaal:2015a}\\

\begin{align}
	\left[
		\begin{array}{c|c|c}
			1 & 1 & 1 \\
			u - c & u & u + c \\
			h - cu & k & h + cu \\
		\end{array}
	\right]
\end{align}

where $h = \frac{c^{2}}{\gamma - 1} + k$ and $k = \frac{1}{2}(u^{2} + v^{2} + w^{2})$.

\section{Derivation of Jacobian and Eigenvalues for the 1D Nuclear Case}

We assume that $p = p(\tau, e, n)$, where $\tau = \frac{1}{\rho}$. Let the vector of conserved variables be $\textbf{U} = \{\rho, m, E, D\}$,
where $m = \rho u$ and $D = \rho y_{e}$. The flux vector is $\textbf{F}(\textbf{U}) = \{m, m^{2}\tau + p, (E+p)m\tau, Dm\tau\}$, where $E=e+\frac{\rho v^{2}}{2}$.

The pressure differential is given by

\begin{equation}
  dp = \left(\pderiv{p}{\rho}\right)d\rho + \left(\pderiv{p}{\epsilon}\right)d\epsilon + \left(\pderiv{p}{D}\right)dD
\end{equation}

The internal energy differential can be written as

\begin{equation}
\begin{split}
	d\epsilon & = \left(\pderiv{e}{\rho}\right)d\rho + \left(\pderiv{\epsilon}{m}\right)dm + \left(\pderiv{e}{E}\right)dE \\
  & = (\epsilon \tau + \frac{1}{2} m^2 \tau^3 )d\rho - m \tau^2 dm + \tau dE
\end{split}
\end{equation}

Plugging this into the previous expression gives with a little work

\begin{equation}
\begin{split}
  dp & = -\tau (p_\tau \tau - p_{\epsilon}(\frac{1}{2} m^2 \tau^2 - \epsilon))d\rho - p_{\epsilon}m\tau^2 dm + p_{\epsilon}\tau dE + \tau P_{y_e}dD\\
     & = -\tau (p_\tau \tau - p_{\epsilon}(E\tau-v^2)d\rho -p_{\epsilon}v\tau dm + p_{\epsilon}\tau dE + \tau P_{y_e}dD
\end{split}
\end{equation}

where $p_{x} = \pderiv{p}{x}$

The elements of the Jacobian $\pderiv{\mathbf{F}(\mathbf{U})}{\mathbf{U}}$ are given by:

\begin{equation}
  \pderiv{f_1}{u_1} = \pderiv{m}{\rho} = 0
\end{equation}

\begin{equation}
  \pderiv{f_1}{u_2} = \pderiv{m}{m} = 1
\end{equation}

\begin{equation}
  \pderiv{f_1}{u_3} = \pderiv{m}{E} = 0
\end{equation}

\begin{equation}
  \pderiv{f_1}{u_4} = \pderiv{m}{D} = 0
\end{equation}

\begin{equation}
\begin{split}
  \pderiv{f_2}{u_1} &= \pderiv{}{\rho}\left(m^2\tau + p\right) = -m^2\tau^2 + \left( \pderiv{p}{\rho} \right) \\
	& =  -v^2 -\tau(p_{\tau}\tau - p_{\epsilon}(E\tau-v^2))
\end{split}
\end{equation}

\begin{equation}
\begin{split}
  \pderiv{f_2}{u_2} & = \pderiv{}{m}\left( m^2 \tau + p \right) = 2m\tau + \left(\pderiv{p}{m}\right) \\
  & = v(2-p_{\epsilon}\tau)
\end{split}
\end{equation}

\begin{equation}
\begin{split}
  \pderiv{f_2}{u_3} & = \pderiv{}{E}\left(m^2 \tau + p\right) = \left( \pderiv{p}{E}\right) \\
  & = \tau p_{\epsilon}
\end{split}
\end{equation}

\begin{equation}
\begin{split}
	\pderiv{f_2}{u_4} = \pderiv{}{D}\left(m^2 \tau + p\right) = p_D
\end{split}
\end{equation}

\begin{equation}
\begin{split}
  \pderiv{f_3}{u_1} & = \pderiv{}{\rho}\left(E+p)m\tau\right) = - E m \tau^2 + m\tau \pderiv{p}{\rho} + mp\pderiv{\tau}{\rho} \\
	& =  -v(H + p_{\tau}\tau^2 - p_{\epsilon}\tau(E\tau-v^2))
\end{split}
\end{equation}

\begin{equation}
\begin{split}
  \pderiv{f_3}{u_2} & = \pderiv{}{m}\left((E+p)m\tau\right) = m\tau\left(\pderiv{p}{m}\right) + \left(E+p\right) \tau \\
  & = H - p_{\epsilon}v^{2}\tau
\end{split}
\end{equation}

\begin{equation}
\begin{split}
  \pderiv{f_3}{u_3} = \pderiv{}{E}\left( (E+p)m\tau\right) = v(1 +\tau p_{\epsilon})
\end{split}
\end{equation}

\begin{equation}
\begin{split}
	\pderiv{f_3}{u_4} = \pderiv{}{n}\left( (E+p)m\tau\right) = v p_{D} = v \tau p_{y_e}
\end{split}
\end{equation}

\begin{equation}
\begin{split}
	\pderiv{f_4}{u_1} = \pderiv{}{\rho}\left( Dm\tau\right) = -v\tau D
\end{split}
\end{equation}

\begin{equation}
\begin{split}
	\pderiv{f_4}{u_2} = \pderiv{}{m}\left( Dm\tau\right) = D\tau
\end{split}
\end{equation}

\begin{equation}
\begin{split}
	\pderiv{f_4}{u_3} = \pderiv{}{E}\left( Dm\tau\right) = 0
\end{split}
\end{equation}

\begin{equation}
\begin{split}
  \pderiv{f_4}{u_4} = \pderiv{}{n}\left( Dm\tau\right) = v
\end{split}
\end{equation}

\begin{align}
	\pderiv{\mathbf{F}(\mathbf{U})}{\mathbf{U}}
	=
	\left[
		\begin{array}{cccc}
			0 & 1 & 0 & 0 \\
			-v^2 -\tau(p_{\tau}\tau - p_{\epsilon}(E\tau-v^2)) & v(2-p_{\epsilon}\tau) & \tau p_{\epsilon} & \tau p_{y_e} \\
			-v(H + p_{\tau}\tau^2 - p_{\epsilon}\tau(E\tau-v^2)) & H - p_{\epsilon}v^{2}\tau & v(1 +\tau p_{\epsilon}) & v \tau p_{y_e} \\
			- v y_{e} & y_{e} & 0 & v \\
		\end{array}
	\right]
\end{align}

The eigenvalues of this Jacobian are then $\lambda = \{u - \sqrt{\tau (p_{y_e}y_{e} + p p_{\epsilon}\tau - p_{\tau}\tau)}, u, u,
u + \sqrt{\tau (p_{y_e}y_{e} + p p_{\epsilon}\tau - p_{\tau}\tau)}\}$. In the
non-nuclear regime where $p_{y_e}=0$, this reduces to the eigenvalues given in \citep{colella:1985}. It also
simplifies to the ideal case given in LeVeque 1992.\\

The eigenvectors are then given by \\

\begin{align}
	\left[
		\begin{array}{c|c|c|c}
			1 & 1 & 1 & 1 \\
			u - c & u & u & u + c \\
			h_{n} - cu & (\frac{p_{\tau}}{p_{\epsilon}}+E)\tau & 0 &
			h_{n} +cu	\\
			n\tau & 0 & \frac{\tau^{2}(p_{\tau} + p_{\epsilon}E)}{p_{n}} & n\tau \\
		\end{array}
	\right]
\end{align}
where $h_{n} = \frac{c^{2}}{p_{\epsilon}\tau} + k_{n}$ and $k_{n} = \frac{(-np_{n} + (p_{\tau} + p_{\epsilon}E)\tau)}{p_{\epsilon}}$.

\section{Derivation of Jacobian and Eigenvalues for the 3D Nuclear Case}

We assume that $p = p(\tau, e, n)$, where $\tau = \frac{1}{\rho}$. Let the vector of conserved variables be $\textbf{U} = \{\rho, m_1, m_2, m_3, E, D\}$,
where $m_i = \rho v_i$ and $D = \rho y_{e}$. The flux vector is $\textbf{F}(\textbf{U}) = \{m_{1}, m_{1}^{2}\tau + p, m_{1}m_{2}\tau, m_{1}m_{3}\tau, (E+p)m_{1}\tau, Dm_{1}\tau\}$.

\beq
	\pderiv{\mathbf{F}(\mathbf{U})}{\mathbf{U}}
	=
		\begin{bmatrix}
			0 & 1 & 0 & 0 & 0 & 0 \\
			-v_{1}^{2} -p_{\tau}\tau^{2} - p_{\epsilon}\tau(E\tau - v^{2}) & v_{1}(2-p_{\epsilon}\tau)  & -p_{\epsilon}v_{2}\tau & -p_{\epsilon}v_{3}\tau  & p_{\epsilon}\tau  & \tau p_{ye} \\
			-v_{1}v_{2} & v_2 & v_1 & 0 & 0 & 0 \\
			-v_{1}v_{3} & v_{3} & 0 & v_{1} & 0 & 0 \\
			v_{1}(-H - p_{\tau}\tau^{2} -p_{\epsilon}\tau(E\tau - v^{2})) & H - p_{\epsilon}v_{1}^{2}\tau  & -p_{\epsilon}v_{1}v_{2}\tau & -p_{\epsilon}v_{1}v_{3}\tau  & v_{1}(1+p_{\epsilon}\tau) & v_{1} \tau p_{ye} \\
			-v_{1} y_{e} & y_{e} & 0 &0 & 0 & v_{1} \\
		\end{bmatrix}
\eeq

The eigenvalues of this Jacobian are then $\lambda = \{v_{1} - \sqrt{\tau (n p_{n} + p p_{\epsilon}\tau - p_{\tau}\tau)}, v_{1}, v_{1}, v_{1}, v_{1},
v_{1} + \sqrt{\tau (n p_{n} + p p_{\epsilon}\tau - p_{\tau}\tau)}\}$.

The eigenvectors are then given by \\

\beq
		\begin{bmatrix}
			1 & 1 & 1 & 1 & 1 & 1 \\
			v_{1} - c  & v_{1} & v_{1} & v_{1} & v_{1} & v_{1}+c \\
			v_{2} & \frac{\beta}{p_{\epsilon}v_{2}} & 0 & 0 & 0 & v_{2} \\
			v_{3} & 0 & \frac{\beta}{p_{\epsilon}v_{3}} & 0 & 0 & v_{3} \\
			h_{n} -cv_{1} & 0 & 0 & -v_{2}^{2} - v_{3}^{2} + (\frac{p_{\tau}}{p_{\epsilon}} + E)\tau & 0 & h_{n} + cv_{1} \\
			n\tau & 0 & 0 & 0 & \frac{-\tau \beta}{p_{n}} & n\tau    \\
		\end{bmatrix}
\eeq

where $h_{n}$ is defined as before, and $\beta = -p_{\tau}\tau + p_{\epsilon}(v_{2}^{2} + v_{3}^{2} -E\tau)$.

\section{Derivatives}

Derivatives of pressure with respect to $\tau$, e, and n. \\
\begin{footnotesize}
\begin{eqnarray}
	\frac{\partial{p}}{\partial{e}} &=& \left(\frac{\partial{e}}{\partial{T}}\right)^{-1}\left(\frac{\partial{p}}{\partial{T}}\right) \\
	\frac{\partial{p}}{\partial{n}} &=& \frac{m_B}{\rho} \left[ \frac{\partial{p}}{\partial{y_e}} -
          \frac{\partial{e}}{\partial{y_e}}\left(\frac{\partial{e}}{\partial{T}}\right)^{-1}\left(\frac{\partial{p}}{\partial{T}}\right)\right]\\
	\frac{\partial{p}}{\partial{\tau}} &=& -\tau^{-2} \left[ \frac{\partial{p}}{\partial{\rho}} - \frac{y_e}{\rho} \left( \frac{\partial{p}}{\partial{y_e}} -
          \frac{\partial{e}}{\partial{y_e}} \left(\frac{\partial{e}}{\partial{T}}\right)^{-1}\left(\frac{\partial{p}}{\partial{T}}\right)\right) -
          \frac{\partial{e}}{\partial{\rho}}\left(\frac{\partial{e}}{\partial{T}}\right)^{-1}\left(\frac{\partial{p}}{\partial{T}}\right)\right]
\end{eqnarray}
\end{footnotesize}

\acknowledgements
We thank people.

\bibliography{ms}

\end{document}

### 2019-02-12 paper/ms.tex
Started numerical methods section

@@
 
+... organized as follows:
+
 \section{Euler Equations of Gas Dynamics in Cartesian Coordinates}
@@
 \beq
-  \pderiv{\rho}{t} + \pderiv{}{x_i}(\rho u_i) = 0
+  \pderiv{\rho}{t} + \pderiv{}{x}(\rho u) = 0
   \label{eq:massConservation}
@@
 \beq
-	\pderiv{(\rho u_i)}{t} + \pderiv{}{x_i}(\rho u^2 + p) = 0
+	\pderiv{(\rho u)}{t} + \pderiv{}{x}(\rho u^2 + p) = 0
   \label{eq:momentumConservation}
@@
 \beq
-  \pderiv{E}{t} + \pderiv{}{x_i}((E+p)u_i) = 0
+  \pderiv{E}{t} + \pderiv{}{x}((E+p)u) = 0
   \label{eq:energyConservation}
 \eeq
-and conservation of \textcolor{red}{electron fraction}\footnote{Is this Correct? I don't think so. What is conserved?}
+and conservation of \textcolor{red}{electron density}
 \beq
-  \pderiv{D}{t} + \pderiv{}{x_i}(D u_i) = 0
+  \pderiv{D}{t} + \pderiv{}{x}(D u) = 0
   \label{eq:electronConservation}
@@
 and $\mathbf{F}(\mathbf{U})=(\rho u,\rho u^{2}+p,(E+p)u, uD)^{T}$
-is the flux vector. We can rewrite this further:
+is the flux vector.
+
+\section{Numerical Implementation}
+\label{DG}
+
+\todo{Add intro paragraph} \\
+\textcolor{red}{Subsections or no?}
+
+\subsection{The Discontinuous Galerkin Methods}
+In our solver we have chosen the discontinuous Galerkin (DG) method
+for our spatial discretization. In this section we will briefly discuss our
+implementation of the DG method, introducing notation and concepts. Recall that
+we seek solutions to the Euler equations of hydrodynamics, constituting a
+conservation law of the form
+
+\beq
+  \partial_{t} \mathbf{U} + \partial_{x} \mathbf{F}(\mathbf{U}) = 0
+\eeq
+
+We can rewrite this further:
 
@@
   \pderiv{\mathbf{w}}{t} +
-  \Lambda \pderiv{\mathbf{w}}{x_{i}}
+  \Lambda \pderiv{\mathbf{w}}{x}
   = 0

### 2019-02-19 paper/ms.tex
Finished draft of DG subsection

@@
 We may rewrite Equations \eqref{eq:massConservation}-\eqref{eq:electronConservation}
-is a more convenient way:
+in a more convenient way:
 
@@
   +\pderiv{\mathbf{F}(\mathbf{U})}{x}
-  =0.
+  =0,
 \eeq
@@
   \partial_{t} \mathbf{U} + \partial_{x} \mathbf{F}(\mathbf{U}) = 0
+  \label{eq:ode}
+\eeq
+Where $\mathbf{U}$ is the evolved state vector and $\mathbf{F}(\mathbf{U})$ is
+the flux. In order to solve Equation \eqref{eq:ode}, we divide the computational
+domain $D\subset \mathbb{R}$ into a disjoint union $\mathcal{T}$ of open elements
+$\bK$ such that $D = \cup_{\bK \in \cT}\bK$. Each element $\bK$ is a box in the
+coordinates
+\beq
+  \bK=\{\,\vect{x} = (\xL,x_H) : x \in K \},
+\eeq
+We let the approximation space $\mathbb{V}^{k}$ for the DG method
+be constructed from the tensor product of one-dimensional
+polynomials of degree $k$. Note that functions in $\mathbb{V}^{k}$
+can be discontinuous across element interfaces. The DG problem is then to find
+$\mathbf{U}_h \in \mathbb{V}^{k}$ which approximates $\mathbf{U}$ in Equation
+\eqref{eq:ode}, such that $\forall$ $\phi \in \mathbb{V}^{k}$ and $\bK \in \cT$
+
+\beq
+  \partial_{t} \int_{\bK}\mathbf{U}_h\, \phi\, dx +
+  \int_{\bK}\partial_{x}\, \mathbf{F(\mathbf{U})}\, \phi\, dx = 0.
+\label{eq:semidiscreteDG_almost}
+\eeq
+Integrating the second term by parts, this becomes
+\beq
+\begin{split}
+  & \partial_{t} \int_{\bK}\mathbf{U_{h}}\, \phi\, dx +
+   \widehat{\mathbf{F}}(\mathbf{U}_h)\, \phi^{-} \big|_{x_{H}} \\
+  & + \widehat{\mathbf{F}}(\mathbf{U}_h)\, \phi^{+} \big|_{x_{L}} +
+   \int_{\bK}\mathbf{F(\mathbf{U}_h)}\, \partial_{x}\phi\, dx = 0.
+\label{eq:semidiscreteDG}
+\end{split}
+\eeq
+In Eq.~\eqref{eq:semidiscreteDG}, $\widehat{\vect{F}}(\vect{U}_{h})$ is a
+numerical flux approximating the flux on the boundary of $\bK$.
+The numerical flux function is evaluated using values from
+both boundaries of an element; i.e.,
+\begin{equation}
+  \widehat{\vect{F}}(\vect{U}_{h})=\vect{f}(\vect{U}_{h}(x^{-},\tilde{\bx}),\vect{U}_{h}(x^{+},\tilde{\bx})),
+\end{equation}
+where superscripts $-/+$
+indicate that the function is evaluated to the immediate left/right of the
+interface. We use the Harten-Lax-van Leer (HLL) flux \citep{harten:1983} or
+the HLLC flux \citep{toro:1994,mignone:2005} for all the numerical
+experiments presented in Section \ref{sec:tests}.
+
+In each element $K$, we use a nodal representation of the conserved variables
+$\mathbf{U}$:
+
+\beq
+\begin{split}
+  \mathbf{U}(x,t) \approx\, & \mathbf{U}_h (x,t) =
+  \sum_{i=1}^{N}\vect{U}_{i}(t)\,\ell_{i}(x),
+  \quad\text{where}\quad \\
+  & \ell_{i}(\eta)=
+  \prod_{\substack{j=1\\j\ne i}}^{N}\f{\eta-\eta_{j}}{\eta_{i}-\eta_{j}}
+  \label{eq:conservedNodalExpansion}
+\end{split}
+\eeq
+are Lagrange polynomials defined on $I = \{ \eta : \eta \in (-0.5,0.5) \}$,
+and are constructed to interpolate the node set
+$S_{N}=\{\eta_{i}\}_{i=1}^{N}\subset I$. The spatial coordinate $x$ and the
+reference coordinate $\eta$ are related by the
+mapping $x(\eta)=\xL+(0.5+\eta)\,\Delta x$.
+Then, for any $\eta_{j}\in S_{N}$, $\ell_{i}(\eta_{j})=\delta_{ij}$,
+so that $\vect{U}_{h}(x(\eta_{j}),t)=\vect{U}_{j}(t)$.
+First we define the $M$-point quadrature $Q_{M}:C^{0}(I)\to\mathbb{R}$
+with abscissas $\hat{S}_{M}=\{\eta_{q}\}_{q=1}^{M}$ and weights
+$\{w_{q}\}_{q=1}^{M}$, normalized such that $\sum_{q=1}^{M}w_{q}=1$ in order to
+evaluate the integrals in Eq.~\eqref{eq:semidiscreteDG}..
+We use the $M$-point Legendre-Gauss quadrature, which is exact for
+polynomials of degree $\le 2M-1$.
+Then, if $P_{h}(x)$ is such a polynomial, we have
+\beq
+  \f{1}{\Delta x}\int_{K}P_{h}(x)\,dx=\int_{I}P_{h}(\eta)\,d\eta=\sum_{q=1}^{M}w_{q}\,P_{h}(\eta_{q}).
 \eeq
 
+For the sake of efficiency we let $M=N$ and $S_{N}=\hat{S}_{N}$, which is a
+spectral-type nodal collocation DG approximation \citep{bassi:2013}.
+Inserting Eq.~\eqref{eq:conservedNodalExpansion} into Eq.~\eqref{eq:semidiscreteDG},
+letting $\phi(x)=\ell_{k}(x)$, using the quadratures defined above,
+we obtain\footnote{Are these summations?}
+\begin{align}
+  \pd{}{t}\int_{K}\vect{U}_{h}\,\phi\,dx
+  &\approx w_{k}\,\pd{}{t}\vect{U}_{k}\,\Delta x
+  \label{eq:timeDerivativeTerm}
+\end{align}
+for the time derivative piece.
+Similarly, the last term on the left-hand side of Eq.~\eqref{eq:semidiscreteDG}
+becomes
+\beq
+  \int_{K}\vect{F}(\vect{U}_{h})\,\pderiv{\phi}{x}\,dx
+  \approx \sum_{q=1}^{N}w_{q}\,\vect{F}(\vect{U}_{q})\,\pderiv{\ell_{k}}{\eta}(\eta_{q}).
+  \label{eq:volumeTerm}
+\eeq
+Now we may combine Eqs.~\eqref{eq:timeDerivativeTerm}-\eqref{eq:volumeTerm} in
+Eq.~\eqref{eq:semidiscreteDG} resulting in the semi-discrete form
+\beq
+\begin{split}
+  \deriv{\vect{U}_{k}}{t}
+  & =-\f{1}{w_{k}\Delta x}
+  \Big\{
+  \Big[\,
+    \widehat{\vect{F}}\ell_{k}\big|_{x_{H}}
+     -\widehat{\vect{F}}\ell_{k}\big|_{\xL}
+  \,\Big] \\
+  & -\sum_{q=1}^{N}w_{q}\,\vect{F}(\vect{U}_{q})\,\pderiv{\ell_{k}}{\eta}(\eta_{q})
+  \Big\}.
+  \label{eq:semidiscreteDiscretized}
+\end{split}
+\eeq
+We now have a system of ordinary
+differential equations (ODEs), which may be evolved in time with an ODE solver.
+In Section~\ref{sec:tests} we use the third-order strong
+stability-preserving Runge-Kutta (SSP-RK3) method \citep{shu:1988}.
+
+\subsection{Slope Limiting}
+
+\subsection{Characteristic Decomposition}
+Experience has shown...
 We can rewrite this further:
@@
   = 0
-  \label{charEq}
+  \label{eq:charEq}
 \eeq
@@
 $\mathbf{w} = R^{-1}\mathbf{U}$. Multiplying both sides of Equation
-\eqref{charEq} by $R^{-1}$, we are able to linearize the system of equations to
+\eqref{eq:charEq} by $R^{-1}$, we are able to linearize the system of equations to
 a system of advection equations

### 2019-02-19 paper/ms.tex
Small 'Slope Limiting Subsection' Draft

@@
 \newcommand{\slugcom}{Draft - \today}
+\newcommand{\Pos}{\textnormal{\tiny\textsc{Pos}}}
+\newcommand{\TVD}{\textnormal{\tiny\textsc{Tvd}}}
 %\newcommand{\slugcom}{Submitted to ApJL on \subdate}
@@
 \subsection{Slope Limiting}
+A common feature of numerical ODEs is unphysical oscillations in the solutions.
+It is therfore of great interest in the DG algorithm to implement slope limiting
+of the polynomial $\mathbf{U}_h$. We use the total variation diminishing (TVD)
+slope limiter \citep[see, e.g.,][]{cockburn:1998} in conjunction with the
+troubled cell indicator discussed in \citet{fu:2017} to prevent excessive limiting.
+
+The general $s$-stage Runge-Kutta time stepping algorithm, including
+the limiting process, can then be summarized as in \citet{cockburn:2001}:
+\begin{itemize}
+  \item[1.] Set $\bar{\vect{U}}^{(0)} = \bar{\vect{U}}^{n}$,
+  \item[2.] For $i=1,\ldots,s$ compute:
+  \begin{equation}
+  \small  \bar{\vect{U}}^{(i)}
+    = \Lambda^{\Pos}\Big\{\Lambda^{\TVD}\Big\{\sum_{j=0}^{i-1}\alpha_{ij}\,\bar{\vect{U}}^{(j)}+\beta_{ij}\,\Delta t\,\bar{\vect{F}}\big(\bar{\vect{U}}^{(j)}\big)\Big\}\Big\},
+    \label{eq:rkStages}
+  \end{equation}
+  \item[3.] Set $\bar{\vect{U}}^{n+1}=\bar{\vect{U}}^{(s)}$.
+\end{itemize}
+Above, the TVD limiter and the `positivity' limiter preventing unphysical
+states are denoted by the operators $\Lambda^{\TVD}\{\}$
+and $ \Lambda^{\Pos}\{\}$, respectively. The SSP-RK3 coefficients $\alpha_{ij}$
+and $\beta_{ij}$ may be found in Table~2.1 in \citet{cockburn:2001}.
 

### 2019-02-19 paper/ms.tex
Draft: numerics, done.

@@
   =0,
+  \label{eq:conservation}
 \eeq
@@
 \subsection{Characteristic Decomposition}
-Experience has shown...
-We can rewrite this further:
-
+Experience has shown that the slope limiting described in the previous section
+is more effecient when performed on the so-called `characteristic variables`
+as opposed to the conserved variables $\mathbf{U}_h$
+\citep[see, e.g.,][for a description]{cockburn:1998}. We begin by
+rewriting Eq.~\ref{eq:conservation} as follows
 \beq
@@
   + \pderiv{\mathbf{U}}{x} \pderiv{\mathbf{F(\mathbf{U})}}{\mathbf{U}}
-  = 0
+  = 0.
   \label{eq:charEq}
@@
   \pderiv{\mathbf{F(\mathbf{U})}}{\mathbf{U}} =
-  R \Lambda R^{-1}
+  R \Lambda R^{-1},
 \eeq
@@
   \Lambda \pderiv{\mathbf{w}}{x}
-  = 0
+  = 0.
 \eeq
@@
 \citep[see e.g.,][for a description]{cockburn:1998, schaal:2015a}.
-
-\todo{More discussion of limiting?}
 

### 2019-03-04 paper/ms.tex
Some Updates, Added notebook to create FORTRAN expressions.

@@
 In Section~\ref{sec:tests} we use the third-order strong
-stability-preserving Runge-Kutta (SSP-RK3) method \citep{shu:1988}.
+stability-preserving Runge-Kutta (SSP-RK3) method\footnote{Change later if necessary} \citep{shu:1988}.
 
@@
 
+\section{Derivation of Jacobian and Eigensystem for the 1D Nuclear Case}
+
+We assume that pressure $p = p(\tau, e, n)$, where $\tau = \frac{1}{\rho}$ is the specific volume.
+Let the vector of conserved variables be $\textbf{U} = \{\rho, m, E, D\}$,
+where $m = \rho u$ and $D = \rho y_{e}$.
+The flux vector is $\textbf{F}(\textbf{U}) = \{m, m^{2}\tau + p, (E+p)m\tau, Dm\tau\}$, where $E=e+\frac{\rho v^{2}}{2}$.
+
+The pressure differential is given by
+\begin{equation}
+  dp = \left(\pderiv{p}{\rho}\right)d\rho + \left(\pderiv{p}{\epsilon}\right)d\epsilon + \left(\pderiv{p}{D}\right)dD.
+\end{equation}
+The internal energy differential can be written as
+
+\begin{equation}
+\begin{split}
+	d\epsilon & = \left(\pderiv{e}{\rho}\right)d\rho + \left(\pderiv{\epsilon}{m}\right)dm + \left(\pderiv{e}{E}\right)dE \\
+  & = (\epsilon \tau + \frac{1}{2} m^2 \tau^3 )d\rho - m \tau^2 dm + \tau dE.
+\end{split}
+\end{equation}
+
+Plugging this into the previous expression gives with a little work
+
+\begin{equation}
+\begin{split}
+  dp & = -\tau (p_\tau \tau - p_{\epsilon}(\frac{1}{2} m^2 \tau^2 - \epsilon))d\rho - p_{\epsilon}m\tau^2 dm + p_{\epsilon}\tau dE + \tau P_{y_e}dD\\
+     & = -\tau (p_\tau \tau - p_{\epsilon}(E\tau-v^2)d\rho -p_{\epsilon}v\tau dm + p_{\epsilon}\tau dE + \tau P_{y_e}dD
+\end{split}
+\end{equation}
+
+where $p_{x} := \pderiv{p}{x}$
+
+The flux Jacobian $\pderiv{\mathbf{F}(\mathbf{U})}{\mathbf{U}}$ may then be shown to be:
+
+% \begin{equation}
+%   \pderiv{f_1}{u_1} = \pderiv{m}{\rho} = 0
+% \end{equation}
+%
+% \begin{equation}
+%   \pderiv{f_1}{u_2} = \pderiv{m}{m} = 1
+% \end{equation}
+%
+% \begin{equation}
+%   \pderiv{f_1}{u_3} = \pderiv{m}{E} = 0
+% \end{equation}
+%
+% \begin{equation}
+%   \pderiv{f_1}{u_4} = \pderiv{m}{D} = 0
+% \end{equation}
+%
+% \begin{equation}
+% \begin{split}
+%   \pderiv{f_2}{u_1} &= \pderiv{}{\rho}\left(m^2\tau + p\right) = -m^2\tau^2 + \left( \pderiv{p}{\rho} \right) \\
+% 	& =  -v^2 -\tau(p_{\tau}\tau - p_{\epsilon}(E\tau-v^2))
+% \end{split}
+% \end{equation}
+%
+% \begin{equation}
+% \begin{split}
+%   \pderiv{f_2}{u_2} & = \pderiv{}{m}\left( m^2 \tau + p \right) = 2m\tau + \left(\pderiv{p}{m}\right) \\
+%   & = v(2-p_{\epsilon}\tau)
+% \end{split}
+% \end{equation}
+%
+% \begin{equation}
+% \begin{split}
+%   \pderiv{f_2}{u_3} & = \pderiv{}{E}\left(m^2 \tau + p\right) = \left( \pderiv{p}{E}\right) \\
+%   & = \tau p_{\epsilon}
+% \end{split}
+% \end{equation}
+%
+% \begin{equation}
+% \begin{split}
+% 	\pderiv{f_2}{u_4} = \pderiv{}{D}\left(m^2 \tau + p\right) = p_D
+% \end{split}
+% \end{equation}
+%
+% \begin{equation}
+% \begin{split}
+%   \pderiv{f_3}{u_1} & = \pderiv{}{\rho}\left(E+p)m\tau\right) = - E m \tau^2 + m\tau \pderiv{p}{\rho} + mp\pderiv{\tau}{\rho} \\
+% 	& =  -v(H + p_{\tau}\tau^2 - p_{\epsilon}\tau(E\tau-v^2))
+% \end{split}
+% \end{equation}
+%
+% \begin{equation}
+% \begin{split}
+%   \pderiv{f_3}{u_2} & = \pderiv{}{m}\left((E+p)m\tau\right) = m\tau\left(\pderiv{p}{m}\right) + \left(E+p\right) \tau \\
+%   & = H - p_{\epsilon}v^{2}\tau
+% \end{split}
+% \end{equation}
+%
+% \begin{equation}
+% \begin{split}
+%   \pderiv{f_3}{u_3} = \pderiv{}{E}\left( (E+p)m\tau\right) = v(1 +\tau p_{\epsilon})
+% \end{split}
+% \end{equation}
+%
+% \begin{equation}
+% \begin{split}
+% 	\pderiv{f_3}{u_4} = \pderiv{}{n}\left( (E+p)m\tau\right) = v p_{D} = v \tau p_{y_e}
+% \end{split}
+% \end{equation}
+%
+% \begin{equation}
+% \begin{split}
+% 	\pderiv{f_4}{u_1} = \pderiv{}{\rho}\left( Dm\tau\right) = -v\tau D
+% \end{split}
+% \end{equation}
+%
+% \begin{equation}
+% \begin{split}
+% 	\pderiv{f_4}{u_2} = \pderiv{}{m}\left( Dm\tau\right) = D\tau
+% \end{split}
+% \end{equation}
+%
+% \begin{equation}
+% \begin{split}
+% 	\pderiv{f_4}{u_3} = \pderiv{}{E}\left( Dm\tau\right) = 0
+% \end{split}
+% \end{equation}
+%
+% \begin{equation}
+% \begin{split}
+%   \pderiv{f_4}{u_4} = \pderiv{}{n}\left( Dm\tau\right) = v
+% \end{split}
+% \end{equation}
+
+\begin{align}
+	\pderiv{\mathbf{F}(\mathbf{U})}{\mathbf{U}}
+	=
+	\left[
+		\begin{array}{cccc}
+			0 & 1 & 0 & 0 \\
+			-v^2 -\tau(p_{\tau}\tau - p_{\epsilon}(E\tau-v^2)) & v(2-p_{\epsilon}\tau) & \tau p_{\epsilon} & \tau p_{y_e} \\
+			-v(H + p_{\tau}\tau^2 - p_{\epsilon}\tau(E\tau-v^2)) & H - p_{\epsilon}v^{2}\tau & v(1 +\tau p_{\epsilon}) & v \tau p_{y_e} \\
+			- v y_{e} & y_{e} & 0 & v \\
+		\end{array}
+	\right]
+\end{align}
+
+The eigenvalues of this Jacobian are then $\lambda = \{u - \sqrt{\tau (p_{y_e}y_{e} + p p_{\epsilon}\tau - p_{\tau}\tau)}, u, u,
+u + \sqrt{\tau (p_{y_e}y_{e} + p p_{\epsilon}\tau - p_{\tau}\tau)}\}$. In the
+non-nuclear regime where $p_{y_e}=0$, this reduces to the eigenvalues given in \citep{colella:1985}. It also
+simplifies to the ideal case given in LeVeque 1992.\\
+
+The eigenvectors are then given by \\
+
+\begin{align}
+	\left[
+		\begin{array}{c|c|c|c}
+			1 & 1 & 1 & 1 \\
+			u - c & u & u & u + c \\
+			h_{n} - cu & (\frac{p_{\tau}}{p_{\epsilon}}+E)\tau & 0 &
+			h_{n} +cu	\\
+			n\tau & 0 & \frac{\tau^{2}(p_{\tau} + p_{\epsilon}E)}{p_{n}} & n\tau \\
+		\end{array}
+	\right]
+\end{align}
+where $h_{n} = \frac{c^{2}}{p_{\epsilon}\tau} + k_{n}$ and $k_{n} = \frac{(-np_{n} + (p_{\tau} + p_{\epsilon}E)\tau)}{p_{\epsilon}}$.
+
 \hrulefill
@@
 where $h = \frac{c^{2}}{\gamma - 1} + k$ and $k = \frac{1}{2}(u^{2} + v^{2} + w^{2})$.
-
-\section{Derivation of Jacobian and Eigenvalues for the 1D Nuclear Case}
-
-We assume that $p = p(\tau, e, n)$, where $\tau = \frac{1}{\rho}$. Let the vector of conserved variables be $\textbf{U} = \{\rho, m, E, D\}$,
-where $m = \rho u$ and $D = \rho y_{e}$. The flux vector is $\textbf{F}(\textbf{U}) = \{m, m^{2}\tau + p, (E+p)m\tau, Dm\tau\}$, where $E=e+\frac{\rho v^{2}}{2}$.
-
-The pressure differential is given by
-
-\begin{equation}
-  dp = \left(\pderiv{p}{\rho}\right)d\rho + \left(\pderiv{p}{\epsilon}\right)d\epsilon + \left(\pderiv{p}{D}\right)dD
-\end{equation}
-
-The internal energy differential can be written as
-
-\begin{equation}
-\begin{split}
-	d\epsilon & = \left(\pderiv{e}{\rho}\right)d\rho + \left(\pderiv{\epsilon}{m}\right)dm + \left(\pderiv{e}{E}\right)dE \\
-  & = (\epsilon \tau + \frac{1}{2} m^2 \tau^3 )d\rho - m \tau^2 dm + \tau dE
-\end{split}
-\end{equation}
-
-Plugging this into the previous expression gives with a little work
-
-\begin{equation}
-\begin{split}
-  dp & = -\tau (p_\tau \tau - p_{\epsilon}(\frac{1}{2} m^2 \tau^2 - \epsilon))d\rho - p_{\epsilon}m\tau^2 dm + p_{\epsilon}\tau dE + \tau P_{y_e}dD\\
-     & = -\tau (p_\tau \tau - p_{\epsilon}(E\tau-v^2)d\rho -p_{\epsilon}v\tau dm + p_{\epsilon}\tau dE + \tau P_{y_e}dD
-\end{split}
-\end{equation}
-
-where $p_{x} = \pderiv{p}{x}$
-
-The elements of the Jacobian $\pderiv{\mathbf{F}(\mathbf{U})}{\mathbf{U}}$ are given by:
-
-\begin{equation}
-  \pderiv{f_1}{u_1} = \pderiv{m}{\rho} = 0
-\end{equation}
-
-\begin{equation}
-  \pderiv{f_1}{u_2} = \pderiv{m}{m} = 1
-\end{equation}
-
-\begin{equation}
-  \pderiv{f_1}{u_3} = \pderiv{m}{E} = 0
-\end{equation}
-
-\begin{equation}
-  \pderiv{f_1}{u_4} = \pderiv{m}{D} = 0
-\end{equation}
-
-\begin{equation}
-\begin{split}
-  \pderiv{f_2}{u_1} &= \pderiv{}{\rho}\left(m^2\tau + p\right) = -m^2\tau^2 + \left( \pderiv{p}{\rho} \right) \\
-	& =  -v^2 -\tau(p_{\tau}\tau - p_{\epsilon}(E\tau-v^2))
-\end{split}
-\end{equation}
-
-\begin{equation}
-\begin{split}
-  \pderiv{f_2}{u_2} & = \pderiv{}{m}\left( m^2 \tau + p \right) = 2m\tau + \left(\pderiv{p}{m}\right) \\
-  & = v(2-p_{\epsilon}\tau)
-\end{split}
-\end{equation}
-
-\begin{equation}
-\begin{split}
-  \pderiv{f_2}{u_3} & = \pderiv{}{E}\left(m^2 \tau + p\right) = \left( \pderiv{p}{E}\right) \\
-  & = \tau p_{\epsilon}
-\end{split}
-\end{equation}
-
-\begin{equation}
-\begin{split}
-	\pderiv{f_2}{u_4} = \pderiv{}{D}\left(m^2 \tau + p\right) = p_D
-\end{split}
-\end{equation}
-
-\begin{equation}
-\begin{split}
-  \pderiv{f_3}{u_1} & = \pderiv{}{\rho}\left(E+p)m\tau\right) = - E m \tau^2 + m\tau \pderiv{p}{\rho} + mp\pderiv{\tau}{\rho} \\
-	& =  -v(H + p_{\tau}\tau^2 - p_{\epsilon}\tau(E\tau-v^2))
-\end{split}
-\end{equation}
-
-\begin{equation}
-\begin{split}
-  \pderiv{f_3}{u_2} & = \pderiv{}{m}\left((E+p)m\tau\right) = m\tau\left(\pderiv{p}{m}\right) + \left(E+p\right) \tau \\
-  & = H - p_{\epsilon}v^{2}\tau
-\end{split}
-\end{equation}
-
-\begin{equation}
-\begin{split}
-  \pderiv{f_3}{u_3} = \pderiv{}{E}\left( (E+p)m\tau\right) = v(1 +\tau p_{\epsilon})
-\end{split}
-\end{equation}
-
-\begin{equation}
-\begin{split}
-	\pderiv{f_3}{u_4} = \pderiv{}{n}\left( (E+p)m\tau\right) = v p_{D} = v \tau p_{y_e}
-\end{split}
-\end{equation}
-
-\begin{equation}
-\begin{split}
-	\pderiv{f_4}{u_1} = \pderiv{}{\rho}\left( Dm\tau\right) = -v\tau D
-\end{split}
-\end{equation}
-
-\begin{equation}
-\begin{split}
-	\pderiv{f_4}{u_2} = \pderiv{}{m}\left( Dm\tau\right) = D\tau
-\end{split}
-\end{equation}
-
-\begin{equation}
-\begin{split}
-	\pderiv{f_4}{u_3} = \pderiv{}{E}\left( Dm\tau\right) = 0
-\end{split}
-\end{equation}
-
-\begin{equation}
-\begin{split}
-  \pderiv{f_4}{u_4} = \pderiv{}{n}\left( Dm\tau\right) = v
-\end{split}
-\end{equation}
-
-\begin{align}
-	\pderiv{\mathbf{F}(\mathbf{U})}{\mathbf{U}}
-	=
-	\left[
-		\begin{array}{cccc}
-			0 & 1 & 0 & 0 \\
-			-v^2 -\tau(p_{\tau}\tau - p_{\epsilon}(E\tau-v^2)) & v(2-p_{\epsilon}\tau) & \tau p_{\epsilon} & \tau p_{y_e} \\
-			-v(H + p_{\tau}\tau^2 - p_{\epsilon}\tau(E\tau-v^2)) & H - p_{\epsilon}v^{2}\tau & v(1 +\tau p_{\epsilon}) & v \tau p_{y_e} \\
-			- v y_{e} & y_{e} & 0 & v \\
-		\end{array}
-	\right]
-\end{align}
-
-The eigenvalues of this Jacobian are then $\lambda = \{u - \sqrt{\tau (p_{y_e}y_{e} + p p_{\epsilon}\tau - p_{\tau}\tau)}, u, u,
-u + \sqrt{\tau (p_{y_e}y_{e} + p p_{\epsilon}\tau - p_{\tau}\tau)}\}$. In the
-non-nuclear regime where $p_{y_e}=0$, this reduces to the eigenvalues given in \citep{colella:1985}. It also
-simplifies to the ideal case given in LeVeque 1992.\\
-
-The eigenvectors are then given by \\
-
-\begin{align}
-	\left[
-		\begin{array}{c|c|c|c}
-			1 & 1 & 1 & 1 \\
-			u - c & u & u & u + c \\
-			h_{n} - cu & (\frac{p_{\tau}}{p_{\epsilon}}+E)\tau & 0 &
-			h_{n} +cu	\\
-			n\tau & 0 & \frac{\tau^{2}(p_{\tau} + p_{\epsilon}E)}{p_{n}} & n\tau \\
-		\end{array}
-	\right]
-\end{align}
-where $h_{n} = \frac{c^{2}}{p_{\epsilon}\tau} + k_{n}$ and $k_{n} = \frac{(-np_{n} + (p_{\tau} + p_{\epsilon}E)\tau)}{p_{\epsilon}}$.
 

### 2019-03-07 paper/ms.tex
New Eigenvectors, added to paper. Updated notebooks.

@@
-\documentclass[twocolumn]{aastex62}
+\documentclass[onecolumn]{aastex62}
 \usepackage{times}
@@
 \usepackage{float}
+\usepackage{stfloats}
 \usepackage{natbib}
@@
   \pderiv{\mathbf{F(\mathbf{U})}}{\mathbf{U}} =
-  R \Lambda R^{-1},
+  \mathcal{R} \Lambda \mathcal{R}^{-1},
 \eeq
@@
 
-\section{Derivation of Jacobian and Eigensystem for the 1D Nuclear Case}
+\section{Derivation of Jacobian and Eigenvalues for the 3D Nuclear Case}
 
-We assume that pressure $p = p(\tau, e, n)$, where $\tau = \frac{1}{\rho}$ is the specific volume.
-Let the vector of conserved variables be $\textbf{U} = \{\rho, m, E, D\}$,
-where $m = \rho u$ and $D = \rho y_{e}$.
-The flux vector is $\textbf{F}(\textbf{U}) = \{m, m^{2}\tau + p, (E+p)m\tau, Dm\tau\}$, where $E=e+\frac{\rho v^{2}}{2}$.
+We assume that $p = p(\tau, e, n)$, where $\tau = \frac{1}{\rho}$.
+Let the vector of conserved variables be $\textbf{U} = \{\rho, m_1, m_2, m_3, E, D\}$,
+where $m_i = \rho v_i$ and $D = \rho Y_{e}$.
+The flux vector is $\textbf{F}(\textbf{U}) =
+\{m_{1}, m_{1}^{2}\tau + p, m_{1}m_{2}\tau, m_{1}m_{3}\tau,
+(E+p)m_{1}\tau, Dm_{1}\tau\}$. \\
 
-The pressure differential is given by
-\begin{equation}
-  dp = \left(\pderiv{p}{\rho}\right)d\rho + \left(\pderiv{p}{\epsilon}\right)d\epsilon + \left(\pderiv{p}{D}\right)dD.
-\end{equation}
-The internal energy differential can be written as
-
-\begin{equation}
-\begin{split}
-	d\epsilon & = \left(\pderiv{e}{\rho}\right)d\rho + \left(\pderiv{\epsilon}{m}\right)dm + \left(\pderiv{e}{E}\right)dE \\
-  & = (\epsilon \tau + \frac{1}{2} m^2 \tau^3 )d\rho - m \tau^2 dm + \tau dE.
-\end{split}
-\end{equation}
-
-Plugging this into the previous expression gives with a little work
-
-\begin{equation}
-\begin{split}
-  dp & = -\tau (p_\tau \tau - p_{\epsilon}(\frac{1}{2} m^2 \tau^2 - \epsilon))d\rho - p_{\epsilon}m\tau^2 dm + p_{\epsilon}\tau dE + \tau P_{y_e}dD\\
-     & = -\tau (p_\tau \tau - p_{\epsilon}(E\tau-v^2)d\rho -p_{\epsilon}v\tau dm + p_{\epsilon}\tau dE + \tau P_{y_e}dD
-\end{split}
-\end{equation}
-
-where $p_{x} := \pderiv{p}{x}$
-
-The flux Jacobian $\pderiv{\mathbf{F}(\mathbf{U})}{\mathbf{U}}$ may then be shown to be:
-
-% \begin{equation}
-%   \pderiv{f_1}{u_1} = \pderiv{m}{\rho} = 0
-% \end{equation}
-%
-% \begin{equation}
-%   \pderiv{f_1}{u_2} = \pderiv{m}{m} = 1
-% \end{equation}
-%
-% \begin{equation}
-%   \pderiv{f_1}{u_3} = \pderiv{m}{E} = 0
-% \end{equation}
-%
-% \begin{equation}
-%   \pderiv{f_1}{u_4} = \pderiv{m}{D} = 0
-% \end{equation}
-%
-% \begin{equation}
-% \begin{split}
-%   \pderiv{f_2}{u_1} &= \pderiv{}{\rho}\left(m^2\tau + p\right) = -m^2\tau^2 + \left( \pderiv{p}{\rho} \right) \\
-% 	& =  -v^2 -\tau(p_{\tau}\tau - p_{\epsilon}(E\tau-v^2))
-% \end{split}
-% \end{equation}
-%
-% \begin{equation}
-% \begin{split}
-%   \pderiv{f_2}{u_2} & = \pderiv{}{m}\left( m^2 \tau + p \right) = 2m\tau + \left(\pderiv{p}{m}\right) \\
-%   & = v(2-p_{\epsilon}\tau)
-% \end{split}
-% \end{equation}
-%
-% \begin{equation}
-% \begin{split}
-%   \pderiv{f_2}{u_3} & = \pderiv{}{E}\left(m^2 \tau + p\right) = \left( \pderiv{p}{E}\right) \\
-%   & = \tau p_{\epsilon}
-% \end{split}
-% \end{equation}
-%
-% \begin{equation}
-% \begin{split}
-% 	\pderiv{f_2}{u_4} = \pderiv{}{D}\left(m^2 \tau + p\right) = p_D
-% \end{split}
-% \end{equation}
-%
-% \begin{equation}
-% \begin{split}
-%   \pderiv{f_3}{u_1} & = \pderiv{}{\rho}\left(E+p)m\tau\right) = - E m \tau^2 + m\tau \pderiv{p}{\rho} + mp\pderiv{\tau}{\rho} \\
-% 	& =  -v(H + p_{\tau}\tau^2 - p_{\epsilon}\tau(E\tau-v^2))
-% \end{split}
-% \end{equation}
-%
-% \begin{equation}
-% \begin{split}
-%   \pderiv{f_3}{u_2} & = \pderiv{}{m}\left((E+p)m\tau\right) = m\tau\left(\pderiv{p}{m}\right) + \left(E+p\right) \tau \\
-%   & = H - p_{\epsilon}v^{2}\tau
-% \end{split}
-% \end{equation}
-%
-% \begin{equation}
-% \begin{split}
-%   \pderiv{f_3}{u_3} = \pderiv{}{E}\left( (E+p)m\tau\right) = v(1 +\tau p_{\epsilon})
-% \end{split}
-% \end{equation}
-%
-% \begin{equation}
-% \begin{split}
-% 	\pderiv{f_3}{u_4} = \pderiv{}{n}\left( (E+p)m\tau\right) = v p_{D} = v \tau p_{y_e}
-% \end{split}
-% \end{equation}
-%
-% \begin{equation}
-% \begin{split}
-% 	\pderiv{f_4}{u_1} = \pderiv{}{\rho}\left( Dm\tau\right) = -v\tau D
-% \end{split}
-% \end{equation}
-%
-% \begin{equation}
-% \begin{split}
-% 	\pderiv{f_4}{u_2} = \pderiv{}{m}\left( Dm\tau\right) = D\tau
-% \end{split}
-% \end{equation}
-%
-% \begin{equation}
-% \begin{split}
-% 	\pderiv{f_4}{u_3} = \pderiv{}{E}\left( Dm\tau\right) = 0
-% \end{split}
-% \end{equation}
-%
-% \begin{equation}
-% \begin{split}
-%   \pderiv{f_4}{u_4} = \pderiv{}{n}\left( Dm\tau\right) = v
-% \end{split}
-% \end{equation}
+\begin{table*}[ht!]
 
 \begin{align}
+  \centering
 	\pderiv{\mathbf{F}(\mathbf{U})}{\mathbf{U}}
-	=
-	\left[
-		\begin{array}{cccc}
-			0 & 1 & 0 & 0 \\
-			-v^2 -\tau(p_{\tau}\tau - p_{\epsilon}(E\tau-v^2)) & v(2-p_{\epsilon}\tau) & \tau p_{\epsilon} & \tau p_{y_e} \\
-			-v(H + p_{\tau}\tau^2 - p_{\epsilon}\tau(E\tau-v^2)) & H - p_{\epsilon}v^{2}\tau & v(1 +\tau p_{\epsilon}) & v \tau p_{y_e} \\
-			- v y_{e} & y_{e} & 0 & v \\
+	= \left[
+		\begin{array}{cccccc}
+			0 & 1 & 0 & 0 & 0 & 0 \\
+			-v_{1}^{2} -p_{\tau}\tau^{2} - p_{\epsilon}\tau(\epsilon - \frac{v^{2}}{2}) & v_{1}(2-p_{\epsilon}\tau)  & -p_{\epsilon}v_{2}\tau & -p_{\epsilon}v_{3}\tau  & p_{\epsilon}\tau  & p_{n} \\
+			-v_{1}v_{2} & v_2 & v_1 & 0 & 0 & 0 \\
+			-v_{1}v_{3} & v_{3} & 0 & v_{1} & 0 & 0 \\
+			v_{1}(-H - p_{\tau}\tau^{2} -p_{\epsilon}\tau(\epsilon - \frac{v^{2}}{2})) & H - p_{\epsilon}v_{1}^{2}\tau  & -p_{\epsilon}v_{1}v_{2}\tau & -p_{\epsilon}v_{1}v_{3}\tau  & v_{1}(1+p_{\epsilon}\tau) & v_{1} p_{n} \\
+			-v_{1} Y_{e}/m_{B} & Y_{e}/m_{B} & 0 & 0 & 0 & v_{1} \\
 		\end{array}
@@
 \end{align}
+\caption{}
+\end{table*}
 
-The eigenvalues of this Jacobian are then $\lambda = \{u - \sqrt{\tau (p_{y_e}y_{e} + p p_{\epsilon}\tau - p_{\tau}\tau)}, u, u,
-u + \sqrt{\tau (p_{y_e}y_{e} + p p_{\epsilon}\tau - p_{\tau}\tau)}\}$. In the
-non-nuclear regime where $p_{y_e}=0$, this reduces to the eigenvalues given in \citep{colella:1985}. It also
-simplifies to the ideal case given in LeVeque 1992.\\
-
-The eigenvectors are then given by \\
+\noindent The eigenvalues of this Jacobian are given by
 
 \begin{align}
+\Lambda =
+\begin{bmatrix}
+  v_{1} - c_{s} & 0 & 0& 0& 0& 0\\
+  0 & v_{1} & 0 & 0 & 0 & 0 \\
+  0 & 0 & v_{1} & 0 & 0 & 0 \\
+  0 & 0 & 0 & v_{1} & 0 & 0 \\
+  0 & 0 & 0 & 0 & v_{1} & 0 \\
+  0 & 0 & 0 & 0 & 0 & v_{1} + c_{s}
+\end{bmatrix}
+\end{align}
+where $c_{S} = \sqrt{\tau (n p_{n} + p p_{\epsilon}\tau - p_{\tau}\tau)}$ is
+the local sound speed. The right eigenvectors are then given by
+\begin{align*}
+  \mathcal{R}_{1} =
   \left[
-		\begin{array}{c|c|c|c}
-			1 & 1 & 1 & 1 \\
-			u - c & u & u & u + c \\
-			h_{n} - cu & (\frac{p_{\tau}}{p_{\epsilon}}+E)\tau & 0 &
-			h_{n} +cu	\\
-			n\tau & 0 & \frac{\tau^{2}(p_{\tau} + p_{\epsilon}E)}{p_{n}} & n\tau \\
+  \begin{array}{cccccc}
+   1 & 0 & 1 & 1 & 0 & 1 \\
+   v_{1}-c & 0 & v_{1} & v_{1} & 0 & c+v_{1} \\
+   v_{2} & 1 & 0 & 0 & 0 & v_{2} \\
+   v_{3} & 0 & 0 & 0 & 1 & v_{3} \\
+   h-c v_{1} & v_{2} & \beta & 0 & v_{3} & h+c v_{1} \\
+   n \tau  & 0 & 0 & \frac{\tau  \chi }{2 p_{n}} & 0 & n \tau  \\
   \end{array}
   \right]
-\end{align}
-where $h_{n} = \frac{c^{2}}{p_{\epsilon}\tau} + k_{n}$ and $k_{n} = \frac{(-np_{n} + (p_{\tau} + p_{\epsilon}E)\tau)}{p_{\epsilon}}$.
-
-\hrulefill
-
-\section{Derivation of the Jacobian and Eigenvectors for the 1D Ideal Case}
-
-\todo{Define new variables to rewrite some things in more compact forms.}
-
-We assume that $p = p(\tau, e, n)$, where $\tau = \frac{1}{\rho}$. Let the vector of conserved variables be $\textbf{U} = \{\rho, m, E\}$,
-where $m = \rho u$. The flux vector is $\textbf{F}(\textbf{U}) = \{m, m^{2} \tau + p, (E+p)m \tau\}$.
-
-\small{
-\begin{align}
-	\pderiv{\mathbf{F}(\mathbf{U})}{\mathbf{U}}
-	=
+\end{align*}
+where the follow definitions have been used:
+$h_{n} = \frac{c^2}{p_{\epsilon}\tau} + k $, $k = \frac{-n p_{n}
++ p_{\epsilon} (\frac{1}{2}v^2 + \epsilon) + p_{\tau}\tau}{p_{\epsilon}}$,
+$\delta_{1} = v_{1}^{2}-v_{2}^{2}-v_{3}^{2}$,
+$\chi = p_{\epsilon} ( \delta_{1} + 2\epsilon) + 2p_{\tau}\tau$, and
+$\beta = \frac{1}{2} (\delta_{1}+2 \epsilon +\frac{2 p_{\tau} \tau }{p_{\epsilon}})$.
+\begin{align*}
+  \mathcal{R}_{1}^{-1} = \frac{1}{c^2}
   \left[
-		\begin{array}{ccc}
-			0 & 1 & 0  \\
-			-(p_{\tau}+ p_{\epsilon}E)\tau^{2} + u^{2}(p_{\epsilon}\tau -1)       & u(2-p_{\epsilon}\tau)         & p_{\epsilon}\tau       \\
-			-u(\tau(-p_{\epsilon}u^{2} + p_{\tau}\tau + p_{\epsilon}E\tau) + H) & H - p_{\epsilon}u^{2}\tau  & u(1+p_{\epsilon}\tau)  \\
+  \begin{array}{cccccc}
+   \frac{1}{4} (2 c  v_{1}+\omega ) & \frac{1}{2} (-c- \phi_{1} ) & -\frac{1}{2} \phi_{2}
+    & -\frac{1}{2} \phi_{3}  & \frac{p_{\epsilon} \tau }{2} & \frac{p_{n}}{2}
+     \\
+   -\frac{v_{2} \omega }{2} & \phi_{1} v_{2}  & c^2+\phi_{2} v_{2}  &
+     \phi_{2} v_{3}  & -\phi_{2}  & -p_{n} v_{2} \\
+   \frac{2 \chi  c^2+\alpha  \omega }{2 \chi } & -\frac{\phi_{1} \alpha  }{\chi } &
+     -\frac{\phi_{2} \alpha  }{\chi } & -\frac{\phi_{3} \alpha }{\chi } &
+     \frac{p_{\epsilon} \alpha  \tau }{\chi } & \frac{p_{n} \left(\alpha  \tau -2 c^2\right)}{\tau  \chi } \\
+   -\frac{n p_{n} \omega }{\chi } & \frac{2 n p_{n} \phi_{1} }{\chi } & \frac{2 n p_{n}
+     \phi_{2} }{\chi } & \frac{2 n p_{n} \phi_{3} }{\chi } & -\frac{2 n
+     p_{n} p_{\epsilon} \tau }{\chi } & \frac{2 p_{n} \left(c^2-n p_{n} \tau \right)}{\tau  \chi } \\
+   -\frac{v_{3} \omega }{2} & \phi_{3} v_{1}  & \phi_{3} v_{2}   &
+     c^2+\phi_{3} v_{3}  & -\phi_{3}  & -p_{n} v_{3} \\
+   \frac{1}{4} (\omega -2 c  v_{1}) & \frac{1}{2} (c-\phi_{1} ) & -\frac{1}{2} \phi_{2}
+       & -\frac{1}{2} \phi_{3}  & \frac{p_{\epsilon} \tau }{2} & \frac{p_{n}}{2}
+     \\
   \end{array}
   \right]
-\end{align}
-}
+\end{align*}
+where $\phi_{i} = p_{\epsilon}\tau v_{i}$.
 
-Here, $H = (E+p)\tau$. The eigenvalues are given by $\lambda = \{u - \tau \sqrt{p p_{\epsilon} - p_{\tau}}, u,  u + \tau \sqrt{ p p_{\epsilon} - p_{\tau}}\}$. In the ideal case where
-$\tau = \rho^{-1}$, $ E = \frac{p}{\gamma -1} + \frac{\rho v^{2}}{2}$, $ p_{\epsilon} = \frac{\gamma-1}{\tau}$, and $p_{\tau} = -\frac{(\gamma - 1)e}{\tau^{2}}$,
-the sound speed $c$ reduces to $\sqrt{\frac{p \gamma}{\rho}}$. \\
-
-The eigenvectors are given by \\
-
-\begin{align}
-	\left[
-		\begin{array}{c|c|c}
-			1 & 1 & 1 \\
-			u - c & u & u + c \\
-			\frac{\tau( c( p-p_{\epsilon}u^{2} +E) +u( -p_{\tau} + p_{\epsilon}E)\tau)}{c + p_{\epsilon}u\tau} & \frac{\tau(p_{\tau} + p_{\epsilon}E)}{p_{\epsilon}} &
-			\frac{\tau( -c( p-p_{\epsilon}u^{2} +E) +u( p_{\tau} + p_{\epsilon}E)\tau)}{-c + p_{\epsilon}u\tau} \\
-		\end{array}
-	\right]
-\end{align}
-
-which in the ideal case described above reduces to  the eigenvectors given in \citet{schaal:2015a}\\
-
-\begin{align}
-	\left[
-		\begin{array}{c|c|c}
-			1 & 1 & 1 \\
-			u - c & u & u + c \\
-			h - cu & k & h + cu \\
-		\end{array}
-	\right]
-\end{align}
-
-where $h = \frac{c^{2}}{\gamma - 1} + k$ and $k = \frac{1}{2}(u^{2} + v^{2} + w^{2})$.
-
-\section{Derivation of Jacobian and Eigenvalues for the 3D Nuclear Case}
-
-We assume that $p = p(\tau, e, n)$, where $\tau = \frac{1}{\rho}$. Let the vector of conserved variables be $\textbf{U} = \{\rho, m_1, m_2, m_3, E, D\}$,
-where $m_i = \rho v_i$ and $D = \rho y_{e}$. The flux vector is $\textbf{F}(\textbf{U}) = \{m_{1}, m_{1}^{2}\tau + p, m_{1}m_{2}\tau, m_{1}m_{3}\tau, (E+p)m_{1}\tau, Dm_{1}\tau\}$.
-
-\beq
-	\pderiv{\mathbf{F}(\mathbf{U})}{\mathbf{U}}
-	=
-		\begin{bmatrix}
-			0 & 1 & 0 & 0 & 0 & 0 \\
-			-v_{1}^{2} -p_{\tau}\tau^{2} - p_{\epsilon}\tau(E\tau - v^{2}) & v_{1}(2-p_{\epsilon}\tau)  & -p_{\epsilon}v_{2}\tau & -p_{\epsilon}v_{3}\tau  & p_{\epsilon}\tau  & \tau p_{ye} \\
-			-v_{1}v_{2} & v_2 & v_1 & 0 & 0 & 0 \\
-			-v_{1}v_{3} & v_{3} & 0 & v_{1} & 0 & 0 \\
-			v_{1}(-H - p_{\tau}\tau^{2} -p_{\epsilon}\tau(E\tau - v^{2})) & H - p_{\epsilon}v_{1}^{2}\tau  & -p_{\epsilon}v_{1}v_{2}\tau & -p_{\epsilon}v_{1}v_{3}\tau  & v_{1}(1+p_{\epsilon}\tau) & v_{1} \tau p_{ye} \\
-			-v_{1} y_{e} & y_{e} & 0 &0 & 0 & v_{1} \\
-		\end{bmatrix}
-\eeq
-
-The eigenvalues of this Jacobian are then $\lambda = \{v_{1} - \sqrt{\tau (n p_{n} + p p_{\epsilon}\tau - p_{\tau}\tau)}, v_{1}, v_{1}, v_{1}, v_{1},
-v_{1} + \sqrt{\tau (n p_{n} + p p_{\epsilon}\tau - p_{\tau}\tau)}\}$.
-
-The eigenvectors are then given by \\
-
-\beq
-		\begin{bmatrix}
-			1 & 1 & 1 & 1 & 1 & 1 \\
-			v_{1} - c  & v_{1} & v_{1} & v_{1} & v_{1} & v_{1}+c \\
-			v_{2} & \frac{\beta}{p_{\epsilon}v_{2}} & 0 & 0 & 0 & v_{2} \\
-			v_{3} & 0 & \frac{\beta}{p_{\epsilon}v_{3}} & 0 & 0 & v_{3} \\
-			h_{n} -cv_{1} & 0 & 0 & -v_{2}^{2} - v_{3}^{2} + (\frac{p_{\tau}}{p_{\epsilon}} + E)\tau & 0 & h_{n} + cv_{1} \\
-			n\tau & 0 & 0 & 0 & \frac{-\tau \beta}{p_{n}} & n\tau    \\
-		\end{bmatrix}
-\eeq
-
-where $h_{n}$ is defined as before, and $\beta = -p_{\tau}\tau + p_{\epsilon}(v_{2}^{2} + v_{3}^{2} -E\tau)$.
+\todo{What is the physical interpretation of h and k? I tried to follow \citet{schaal:2015a},
+and in their eigenvectors (Ideal), h is specific enthalpy and k is specific kinetic energy.}
+The matrix of left eigenvectors is given by
 

### 2019-03-07 paper/ms.tex
Right and Left eigenvectors reduced and added to manuscript. Old notebooks deleted.

@@
 \usepackage{definitions}
+\usepackage{subfiles}
 
@@
 \begin{abstract}
-\todo{Fill In Abstract}
 A problem of high importance in computational astrophysics is obtaining accurate
@@
 pressure, $D=\rho y_e$ is the \textcolor{red}{what is this?},
-and $E=e+\frac{1}{2}\rho u^2$ the total energy (internal plus kinetic).
+and $E=\epsilon \rho +\frac{1}{2}\rho u^2$ the total specific energy (internal plus kinetic).
 The system of equations is closed by the EOS which links the pressure to
@@
 troubled cell indicator discussed in \citet{fu:2017} to prevent excessive limiting.
-
 The general $s$-stage Runge-Kutta time stepping algorithm, including
@@
 where the columns of $R$ contain the right eigenvectors of the Jacobian,
-the rows of $R^{-1}$ contain the left eigenvectors, and
+the rows of $\mathcal{R}^{-1}$ contain the left eigenvectors, and
 $\Lambda$ is a diagonal matrix containing the eigenvalues of the Jacobian.
 At this point, we will introduce the characteristic variable
-$\mathbf{w} = R^{-1}\mathbf{U}$. Multiplying both sides of Equation
-\eqref{eq:charEq} by $R^{-1}$, we are able to linearize the system of equations to
+$\mathbf{w} = \mathcal{R}^{-1}\mathbf{U}$. Multiplying both sides of Equation
+\eqref{eq:charEq} by $\mathcal{R}^{-1}$, we are able to linearize the system of equations to
 a system of advection equations
@@
 
-\begin{table*}[ht!]
+%\begin{table*}[ht!]
 
@@
 \end{align}
-\caption{}
-\end{table*}
+%\caption{}
+%\end{table*}
 
@@
 \end{align}
-where $c_{S} = \sqrt{\tau (n p_{n} + p p_{\epsilon}\tau - p_{\tau}\tau)}$ is
+where $c_{s} = \sqrt{\tau (n p_{n} + p p_{\epsilon}\tau - p_{\tau}\tau)}$ is
 the local sound speed. The right eigenvectors are then given by
@@
 
+\appendix
+
 \section{Derivatives}

### 2019-04-11 paper/derivatives.tex
Changes to eigenvectors

Derivatives of pressure with respect to $\tau$, $\epsilon$, and n. \\
\begin{footnotesize}
\begin{eqnarray}
	\frac{\partial{P}}{\partial{\epsilon}} &=& \left(\frac{\partial{\epsilon}}{\partial{T}}\right)^{-1}\left(\frac{\partial{P}}{\partial{T}}\right) \\
	\frac{\partial{P}}{\partial{n}} &=& \frac{m_B}{\rho} \left[ \frac{\partial{P}}{\partial{y_e}} -
          \frac{\partial{\epsilon}}{\partial{y_e}}\left(\frac{\partial{\epsilon}}{\partial{T}}\right)^{-1}\left(\frac{\partial{P}}{\partial{T}}\right)\right]\\
	\frac{\partial{P}}{\partial{\tau}} &=& -\tau^{-2} \left[ \frac{\partial{P}}{\partial{\rho}} - \frac{y_e}{\rho} \left( \frac{\partial{P}}{\partial{y_e}} -
          \frac{\partial{\epsilon}}{\partial{y_e}} \left(\frac{\partial{\epsilon}}{\partial{T}}\right)^{-1}\left(\frac{\partial{P}}{\partial{T}}\right)\right) -
          \frac{\partial{\epsilon}}{\partial{\rho}}\left(\frac{\partial{\epsilon}}{\partial{T}}\right)^{-1}\left(\frac{\partial{P}}{\partial{T}}\right)\right]
\end{eqnarray}
\end{footnotesize}

### 2019-04-11 paper/ms.tex
Changes to eigenvectors

@@
 
+\author{Anthony Mezzacappa}
+\affiliation{Department of Physics and Astronomy, University of Tennessee Knoxville, TN 37996}
+\affiliation{Joint Institute for Computational Sciences, Oak Ridge National Laboratory, TN 37831}
+
 %\begin{abstract}
 %\end{abstract}
+% nonlinear hyperbloic cons. laws. develop discontinuities from smooth ICs
+% this can lead to oscillatins which require some limiting process to smooth.
+% limiting in characteristic variables increases the efficiency of slope limiting
 \begin{abstract}
@@
 solutions to the Euler equations of hydrodynamics. We are interested in solving
-the Euler equations in the context of core collapse supernovae (CCSNe).
+the Euler equations in the context of core collapse supernovae.
 The toolkit for high-order neutrino-radiation hydrodynamics (\thornado)
@@
 based on the discontinuous Galerkin (DG) method.
-Because the Euler equations form a hyperbolic set of partial differential equations,
-they may be linearized to advection equations by the introduction of
-chracteristic variables related to the eigen-decomposition.
+The Euler equations form a hyperbolic set of partial differential equations.
+In the quasi-linear form, the system can be represented
+as a set of independent advection equations that can be limited separately.
 This use of characteristic variables also increases the efficiency of the slope
 limiting process. However, the addition of the nuclear equation of state equation
-to the Euler equations makes the decompisition nontrivial.
+to the Euler equations makes the decomposition nontrivial.
 We introduce the framework for the characteristic decomposition of the Euler equations
@@
 \section{Introduction}
-\label{Intro}
+\label{sec:Intro}
 \todo{Fill in introduction. Contextualize the work.}
@@
 neutrino transport, hydrodynamics must be accurately modelled,
-creating a very challenging problem. Hydrodynamic instabilities are critical
+creating a very challenging and compelling problem. Hydrodynamic instabilities are critical
 in aiding the explosion, with complex phenomena such as turbulence and
-convection play key roles in the CCSN mechanism
+convection playing key roles in the CCSN mechanism
 \citep{murphy:2011, murphy:2013, couch:2013, couch:2015a, radice:2016, mabanta:2018}.
@@
 explosion is to be realistically modelled.
-For reviews of the CCSN mechanism, see \citet{bethe:1990, janka:2007, janka:2012a, janka:2016, burrows:2013, hix:2014, muller:2016, couch:2017}.
+For in-depth reviews of the CCSN mechanism, see \citet{bethe:1990, janka:2007, janka:2012a, janka:2016, burrows:2013, hix:2014, muller:2016, couch:2017}.
 
@@
 
-... organized as follows:
+This paper will be organized as follows: in Section \ref{sec:eulereq} we briefly
+describe the Euler equations of gas dynamics. Section \ref{sec:DG} describes our
+numerical implementation of the DG method in the one dimensional (1D) case.
+Section \ref{sec:results} shows the results of 1D hydrodynamic simulations,
+and Section \ref{sec:eigen} presents the eigenvectors used in the eigen-decomposition
+of the Euler equations.
 
 \section{Euler Equations of Gas Dynamics in Cartesian Coordinates}
+\label{sec:eulereq}
 \todo{Use this form of the Euler Equations?}
 The non-relativistic Euler equations of gas dynamics
-\citep[see, e.g.,][for details]{leveque:2002} in cartesian coordinates are given by
+\citep[see, e.g.,][for details]{leveque:2002} in Cartesian coordinates
+in the absence of sources with a nuclear matter EOS are given by
 the equations of conservation of mass,
 \beq
-  \pderiv{\rho}{t} + \pderiv{}{x}(\rho u) = 0
+  \partial_{t} \rho + \divergence{} (\rho\,  \mathbf{v}) = 0
   \label{eq:massConservation}
@@
 \beq
-	\pderiv{(\rho u)}{t} + \pderiv{}{x}(\rho u^2 + p) = 0
+%	\partial_{t}(\rho \mathbf{v}) + \divergence{}(\rho u^2 + p) = 0
+  \partial_{t}(\rho\,  \mathbf{v}) + \divergence{}(\rho\,  \mathbf{v} \otimes \mathbf{v} + P\, \mathbf{I}) = 0
   \label{eq:momentumConservation}
@@
 \beq
-  \pderiv{E}{t} + \pderiv{}{x}((E+p)u) = 0
+  \partial_{t} E + \divergence{}\left[(E+P)\, \mathbf{v}\right] = 0
   \label{eq:energyConservation}
@@
 \beq
-  \pderiv{D}{t} + \pderiv{}{x}(D u) = 0
+  \partial_{t} D + \divergence{}(D\, \mathbf{v}) = 0
   \label{eq:electronConservation}
 \eeq
-where $\rho$ represents mass density, $u$ the fluid velocity, $p$ the fluid
-pressure, $D=\rho y_e$ is the \textcolor{red}{what is this?},
-and $E=\epsilon \rho +\frac{1}{2}\rho u^2$ the total specific energy (internal plus kinetic).
-The system of equations is closed by the EOS which links the pressure to
-the quantities evolved by Equations
+where $\rho$ represents mass density, $\mathbf{v}$ the fluid velocity vector,
+$P$ the fluid pressure, $D=\rho y_e$ where $y_e$ is the electron fraction,
+$E=\epsilon \rho +\frac{1}{2}\rho v^2$ the total energy (internal plus kinetic),
+$\epsilon$ is the specific internal energy, and $\mathbf{I}$ is the identity tensor.
 \eqref{eq:massConservation}-\eqref{eq:electronConservation}. The inclusion of
 Equation \eqref{eq:electronConservation} is because we require a nuclear EOS.
+These equations are closed by a tabulated EOS where the pressure is given by a function of
+density, temperature $T$ and the electron fraction: $P = P(\rho, T, y_e)$.
 We may rewrite Equations \eqref{eq:massConservation}-\eqref{eq:electronConservation}
@@
 \beq
-  \pderiv{\mathbf{U}}{t}
-  +\pderiv{\mathbf{F}(\mathbf{U})}{x}
-  =0,
+  \partial_{t}\mathbf{U} + \partial_{x}\mathbf{F}(\mathbf{U}) = 0,
   \label{eq:conservation}
 \eeq
-where $\mathbf{U} =(\rho,\rho u,E, D)^{T}$ is the vector of conserved quantities
-and $\mathbf{F}(\mathbf{U})=(\rho u,\rho u^{2}+p,(E+p)u, uD)^{T}$
+where $\mathbf{U} =(\rho,\rho \mathbf{v},E, D)^{T}$ is the vector of conserved quantities
+and $\mathbf{F}(\mathbf{U})=(\rho \mathbf{v},\rho u^{2}+p,(E+P)\mathbf{v}, D\mathbf{v})^{T}$
 is the flux vector.
@@
 \section{Numerical Implementation}
-\label{DG}
+\label{sec:DG}
 
@@
 for our spatial discretization. In this section we will briefly discuss our
-implementation of the DG method, introducing notation and concepts. Recall that
+implementation of the DG method, introducing notation and concepts.
+For simplicity, we will focus on the one dimensional (1D) case. Recall that
 we seek solutions to the Euler equations of hydrodynamics, constituting a
@@
 Where $\mathbf{U}$ is the evolved state vector and $\mathbf{F}(\mathbf{U})$ is
-the flux. In order to solve Equation \eqref{eq:ode}, we divide the computational
+the flux. In order to solve Equation \eqref{eq:ode} numerically, we divide the computational
 domain $D\subset \mathbb{R}$ into a disjoint union $\mathcal{T}$ of open elements
@@
 \beq
-  \bK=\{\,\vect{x} = (\xL,x_H) : x \in K \},
+  \bK \in (\xL,x_{R}),
 \eeq
@@
 \beq
-\begin{split}
-  & \partial_{t} \int_{\bK}\mathbf{U_{h}}\, \phi\, dx +
-   \widehat{\mathbf{F}}(\mathbf{U}_h)\, \phi^{-} \big|_{x_{H}} \\
-  & + \widehat{\mathbf{F}}(\mathbf{U}_h)\, \phi^{+} \big|_{x_{L}} +
+   \partial_{t} \int_{\bK}\mathbf{U_{h}}\, \phi\, dx +
+   \widehat{\mathbf{F}}(\mathbf{U}_h)\, \phi^{-} \big|_{x_{R}}
+   + \widehat{\mathbf{F}}(\mathbf{U}_h)\, \phi^{+} \big|_{x_{L}} +
    \int_{\bK}\mathbf{F(\mathbf{U}_h)}\, \partial_{x}\phi\, dx = 0.
 \label{eq:semidiscreteDG}
-\end{split}
 \eeq
@@
 \begin{equation}
-  \widehat{\vect{F}}(\vect{U}_{h})=\vect{f}(\vect{U}_{h}(x^{-},\tilde{\bx}),\vect{U}_{h}(x^{+},\tilde{\bx})),
+  \widehat{\vect{F}}(\vect{U}_{h})=\vect{f}(\vect{U}_{h}(x^{-}),\vect{U}_{h}(x^{+})),
 \end{equation}
@@
 indicate that the function is evaluated to the immediate left/right of the
-interface. We use the Harten-Lax-van Leer (HLL) flux \citep{harten:1983} or
-the HLLC flux \citep{toro:1994,mignone:2005} for all the numerical
-experiments presented in Section \ref{sec:tests}.
+interface. We use the Harten-Lax-van Leer (HLL) flux \citep{harten:1983}
+for all the numerical experiments presented in Section \ref{sec:results}.
 
@@
 $\{w_{q}\}_{q=1}^{M}$, normalized such that $\sum_{q=1}^{M}w_{q}=1$ in order to
-evaluate the integrals in Eq.~\eqref{eq:semidiscreteDG}..
+evaluate the integrals in Eq.~\eqref{eq:semidiscreteDG}.
 We use the $M$-point Legendre-Gauss quadrature, which is exact for
@@
 letting $\phi(x)=\ell_{k}(x)$, using the quadratures defined above,
-we obtain\footnote{Are these summations?}
+we obtain
 \begin{align}
@@
 \beq
-\begin{split}
   \deriv{\vect{U}_{k}}{t}
-  & =-\f{1}{w_{k}\Delta x}
+   =-\f{1}{w_{k}\Delta x}
   \Big\{
@@
      -\widehat{\vect{F}}\ell_{k}\big|_{\xL}
-  \,\Big] \\
-  & -\sum_{q=1}^{N}w_{q}\,\vect{F}(\vect{U}_{q})\,\pderiv{\ell_{k}}{\eta}(\eta_{q})
+  \,\Big]
+   -\sum_{q=1}^{N}w_{q}\,\vect{F}(\vect{U}_{q})\,\pderiv{\ell_{k}}{\eta}(\eta_{q})
   \Big\}.
   \label{eq:semidiscreteDiscretized}
-\end{split}
 \eeq
@@
 differential equations (ODEs), which may be evolved in time with an ODE solver.
-In Section~\ref{sec:tests} we use the third-order strong
-stability-preserving Runge-Kutta (SSP-RK3) method\footnote{Change later if necessary} \citep{shu:1988}.
+In Section~\ref{sec:results} we use the third-order strong
+stability-preserving Runge-Kutta (SSP-RK3) method \citep{shu:1988}.
 
-\subsection{Slope Limiting}
-A common feature of numerical ODEs is unphysical oscillations in the solutions.
+\subsection{Time Integration}
+A common feature of high-order numerical PDEs is unphysical oscillations in the solutions.
 It is therfore of great interest in the DG algorithm to implement slope limiting
@@
   \small  \bar{\vect{U}}^{(i)}
-    = \Lambda^{\Pos}\Big\{\Lambda^{\TVD}\Big\{\sum_{j=0}^{i-1}\alpha_{ij}\,\bar{\vect{U}}^{(j)}+\beta_{ij}\,\Delta t\,\bar{\vect{F}}\big(\bar{\vect{U}}^{(j)}\big)\Big\}\Big\},
+    = \Big\{\Lambda^{\TVD}\Big\{\sum_{j=0}^{i-1}\alpha_{ij}\,\bar{\vect{U}}^{(j)}+\beta_{ij}\,\Delta t\,\bar{\vect{F}}\big(\bar{\vect{U}}^{(j)}\big)\Big\}\Big\},
     \label{eq:rkStages}
@@
 \citep[see, e.g.,][for a description]{cockburn:1998}. We begin by
-rewriting Eq.~\ref{eq:conservation} as follows
+rewriting Eq.~\eqref{eq:conservation} in the quasi-linear form as follows
 \beq
   \pderiv{\mathbf{U}}{t}
-  + \pderiv{\mathbf{U}}{x} \pderiv{\mathbf{F(\mathbf{U})}}{\mathbf{U}}
+  + \pderiv{\mathbf{U}}{\mathbf{x}} \pderiv{\mathbf{F(\mathbf{U})}}{\mathbf{U}}
   = 0.
@@
 \eeq
-where the columns of $R$ contain the right eigenvectors of the Jacobian,
+where the columns of $\mathcal{R}$ contain the right eigenvectors of the Jacobian,
 the rows of $\mathcal{R}^{-1}$ contain the left eigenvectors, and
@@
 transformed back to the solution of the conserved variables $\mathbf{U}$
-\citep[see e.g.,][for a description]{cockburn:1998, schaal:2015a}.
+\citep[see e.g.,][for a description]{cockburn:1998, schaal:2015a}. While this
+process of characteristic limiting has been done for the ideal EOS
+\citep[][]{cockburn:1998}, we want
+to extend to the nuclear matter EOS case.
+
+\section{Numerical Results}
+\label{sec:results}
+figure figure figure.
+Sod: Definitely want a $y_{e}$ curve, with and without
+characteristic limiting. Perhaps density too, or something. Show comparison
+to the nX=10000 run.
+
+Maybe also do the Noh problem, include results from there.
 
 \section{Derivation of Jacobian and Eigenvalues for the 3D Nuclear Case}
+\label{sec:eigen}
 
-We assume that $p = p(\tau, e, n)$, where $\tau = \frac{1}{\rho}$.
+We assume that $P = P(\tau, \epsilon, n)$, where $\tau = \frac{1}{\rho}$.
 Let the vector of conserved variables be $\textbf{U} = \{\rho, m_1, m_2, m_3, E, D\}$,
-where $m_i = \rho v_i$ and $D = \rho Y_{e}$.
+where $m_i = \rho v_i$ and $D = \rho y_{e}$.
 The flux vector is $\textbf{F}(\textbf{U}) =
-\{m_{1}, m_{1}^{2}\tau + p, m_{1}m_{2}\tau, m_{1}m_{3}\tau,
-(E+p)m_{1}\tau, Dm_{1}\tau\}$. \\
+\{m_{1}, m_{1}^{2}\tau + P, m_{1}m_{2}\tau, m_{1}m_{3}\tau,
+(E+P)m_{1}\tau, Dm_{1}\tau\}$. The Jacobian of the flux vector is given by
 
@@
 			0 & 1 & 0 & 0 & 0 & 0 \\
-			-v_{1}^{2} -p_{\tau}\tau^{2} - p_{\epsilon}\tau(\epsilon - \frac{v^{2}}{2}) & v_{1}(2-p_{\epsilon}\tau)  & -p_{\epsilon}v_{2}\tau & -p_{\epsilon}v_{3}\tau  & p_{\epsilon}\tau  & p_{n} \\
+			-v_{1}^{2} -P_{\tau}\tau^{2} - P_{\epsilon}\tau(\epsilon - \frac{v^{2}}{2}) & v_{1}(2-P_{\epsilon}\tau)  & -P_{\epsilon}v_{2}\tau & -P_{\epsilon}v_{3}\tau  & P_{\epsilon}\tau  & P_{D_{e}} \\
 			-v_{1}v_{2} & v_2 & v_1 & 0 & 0 & 0 \\
 			-v_{1}v_{3} & v_{3} & 0 & v_{1} & 0 & 0 \\
-			v_{1}(-H - p_{\tau}\tau^{2} -p_{\epsilon}\tau(\epsilon - \frac{v^{2}}{2})) & H - p_{\epsilon}v_{1}^{2}\tau  & -p_{\epsilon}v_{1}v_{2}\tau & -p_{\epsilon}v_{1}v_{3}\tau  & v_{1}(1+p_{\epsilon}\tau) & v_{1} p_{n} \\
-			-v_{1} Y_{e}/m_{B} & Y_{e}/m_{B} & 0 & 0 & 0 & v_{1} \\
+			v_{1}(-H - P_{\tau}\tau^{2} -P_{\epsilon}\tau(\epsilon - \frac{v^{2}}{2})) & H - P_{\epsilon}v_{1}^{2}\tau  & -P_{\epsilon}v_{1}v_{2}\tau & -P_{\epsilon}v_{1}v_{3}\tau  & v_{1}(1+P_{\epsilon}\tau) & v_{1} P_{D_{e}} \\
+			-v_{1} y_{e} & y_{e} & 0 & 0 & 0 & v_{1} \\
 		\end{array}
@@
 
-\noindent The eigenvalues of this Jacobian are given by
+\noindent Where $H=(E+P)\tau$ is the specific enthalpy and $P_{x} = \partial_{x}P$\footnote{Maybe define these elsewhere to show what's constant}. The eigenvalues of this Jacobian are given by
 
@@
 \begin{bmatrix}
-  v_{1} - c_{s} & 0 & 0& 0& 0& 0\\
+  v_{1} - c_{s} & 0 & 0& 0& 0& 0 \\
   0 & v_{1} & 0 & 0 & 0 & 0      \\
@@
 \end{align}
-where $c_{s} = \sqrt{\tau (n p_{n} + p p_{\epsilon}\tau - p_{\tau}\tau)}$ is
-the local sound speed. The right eigenvectors are then given by
+where $c_{s} = \sqrt{\Gamma P \tau}$, for
+$\Gamma = \frac{\tau (P P_{\epsilon} - P_{\tau}) + P_{D_e} y_{e} \tau^{-1}}{P}$, is
+the local sound speed. In the ideal EOS regime this reduces to the sound speeds given \citet{colella:1985}.
+The right eigenvectors are then given by
 \begin{align*}
@@
    h-c v_{1} & v_{2} & \beta & 0 & v_{3} & h+c v_{1} \\
-   n \tau  & 0 & 0 & \frac{\tau  \chi }{2 p_{n}} & 0 & n \tau  \\
+   y_{e}  & 0 & 0 & \frac{\tau  \chi }{2 P_{D_{e}}} & 0 & y_{e}  \\
   \end{array}
@@
 where the follow definitions have been used:
-$h_{n} = \frac{c^2}{p_{\epsilon}\tau} + k $, $k = \frac{-n p_{n}
-+ p_{\epsilon} (\frac{1}{2}v^2 + \epsilon) + p_{\tau}\tau}{p_{\epsilon}}$,
+$h_{n} = \frac{c^2}{P_{\epsilon}\tau} + k $, $k = \frac{-y_{e} P_{D_{e}} \tau^{-1}
++ P_{\epsilon} (\frac{1}{2}v^2 + \epsilon) + P_{\tau}\tau}{P_{\epsilon}}$,
 $\delta_{1} = v_{1}^{2}-v_{2}^{2}-v_{3}^{2}$,
-$\chi = p_{\epsilon} ( \delta_{1} + 2\epsilon) + 2p_{\tau}\tau$, and
-$\beta = \frac{1}{2} (\delta_{1}+2 \epsilon +\frac{2 p_{\tau} \tau }{p_{\epsilon}})$.
+$\chi = P_{\epsilon} ( \delta_{1} + 2\epsilon) + 2P_{\tau}\tau$, and
+$\beta = \frac{1}{2} (\delta_{1}+2 \epsilon +\frac{2 P_{\tau} \tau }{P_{\epsilon}})$.
 \begin{align*}
@@
    \frac{1}{4} (2 c  v_{1}+\omega ) & \frac{1}{2} (-c- \phi_{1} ) & -\frac{1}{2} \phi_{2}
-    & -\frac{1}{2} \phi_{3}  & \frac{p_{\epsilon} \tau }{2} & \frac{p_{n}}{2}
+    & -\frac{1}{2} \phi_{3}  & \frac{P_{\epsilon} \tau }{2} & \frac{P_{D_{e}}}{2}
      \\
    -\frac{v_{2} \omega }{2} & \phi_{1} v_{2}  & c^2+\phi_{2} v_{2}  &
-     \phi_{2} v_{3}  & -\phi_{2}  & -p_{n} v_{2} \\
-   \frac{2 \chi  c^2+\alpha  \omega }{2 \chi } & -\frac{\phi_{1} \alpha  }{\chi } &
-     -\frac{\phi_{2} \alpha  }{\chi } & -\frac{\phi_{3} \alpha }{\chi } &
-     \frac{p_{\epsilon} \alpha  \tau }{\chi } & \frac{p_{n} \left(\alpha  \tau -2 c^2\right)}{\tau  \chi } \\
-   -\frac{n p_{n} \omega }{\chi } & \frac{2 n p_{n} \phi_{1} }{\chi } & \frac{2 n p_{n}
-     \phi_{2} }{\chi } & \frac{2 n p_{n} \phi_{3} }{\chi } & -\frac{2 n
-     p_{n} p_{\epsilon} \tau }{\chi } & \frac{2 p_{n} \left(c^2-n p_{n} \tau \right)}{\tau  \chi } \\
-   -\frac{v_{3} \omega }{2} & \phi_{3} v_{1}  & \phi_{3} v_{2}   &
-     c^2+\phi_{3} v_{3}  & -\phi_{3}  & -p_{n} v_{3} \\
+     \phi_{3} v_{2}  & -\phi_{2}  & -P_{D_{e}} v_{2}
+     \\
+   \frac{2 \chi  c^2+\alpha  \omega \tau^{-1} }{2 \chi } & -\frac{\phi_{1} \alpha  }{\chi \tau } &
+     -\frac{\phi_{2} \alpha  }{\chi \tau } & -\frac{\phi_{3} \alpha }{\chi \tau } &
+     \frac{P_{\epsilon} \alpha }{\chi } & \frac{P_{D_{e}} \left(\alpha -2 c^2\right)}{\tau \chi }
+      \\
+   -\frac{y_{e} P_{D_{e}} \omega }{\chi \tau } & \frac{2 y_{e} P_{D_{e}} \phi_{1} }{\chi \tau } & \frac{2 y_{e} P_{D_{e}}
+     \phi_{2} }{\chi \tau} & \frac{2 y_{e} P_{D_{e}} \phi_{3} }{\chi \tau} & -\frac{2 y_{e}
+     P_{D_{e}} P_{\epsilon} }{\chi } & \frac{2 P_{D_{e}} \left(c^2-y_{e} P_{D_{e}} \right)}{\tau  \chi }
+      \\
+   -\frac{v_{3} \omega }{2} & \phi_{1} v_{3}  & \phi_{2} v_{3}   &
+     c^2+\phi_{3} v_{3}  & -\phi_{3}  & -P_{D_{e}} v_{3}
+      \\
    \frac{1}{4} (\omega -2 c  v_{1}) & \frac{1}{2} (c-\phi_{1} ) & -\frac{1}{2} \phi_{2}
-       & -\frac{1}{2} \phi_{3}  & \frac{p_{\epsilon} \tau }{2} & \frac{p_{n}}{2}
+       & -\frac{1}{2} \phi_{3}  & \frac{P_{\epsilon} \tau }{2} & \frac{P_{D_{e}}}{2}
      \\
@@
 \end{align*}
-where $\phi_{i} = p_{\epsilon}\tau v_{i}$.
+where $\phi_{i} = P_{\epsilon}\,\tau\, v_{i}$ and
+$\omega = \tau\, (P_{\epsilon}\,(v^2 - 2\epsilon) - 2\,P_{\tau}\,\tau)$.
 
@@
 \section{Derivatives}
-
-Derivatives of pressure with respect to $\tau$, e, and n. \\
-\begin{footnotesize}
-\begin{eqnarray}
-	\frac{\partial{p}}{\partial{e}} &=& \left(\frac{\partial{e}}{\partial{T}}\right)^{-1}\left(\frac{\partial{p}}{\partial{T}}\right) \\
-	\frac{\partial{p}}{\partial{n}} &=& \frac{m_B}{\rho} \left[ \frac{\partial{p}}{\partial{y_e}} -
-          \frac{\partial{e}}{\partial{y_e}}\left(\frac{\partial{e}}{\partial{T}}\right)^{-1}\left(\frac{\partial{p}}{\partial{T}}\right)\right]\\
-	\frac{\partial{p}}{\partial{\tau}} &=& -\tau^{-2} \left[ \frac{\partial{p}}{\partial{\rho}} - \frac{y_e}{\rho} \left( \frac{\partial{p}}{\partial{y_e}} -
-          \frac{\partial{e}}{\partial{y_e}} \left(\frac{\partial{e}}{\partial{T}}\right)^{-1}\left(\frac{\partial{p}}{\partial{T}}\right)\right) -
-          \frac{\partial{e}}{\partial{\rho}}\left(\frac{\partial{e}}{\partial{T}}\right)^{-1}\left(\frac{\partial{p}}{\partial{T}}\right)\right]
-\end{eqnarray}
-\end{footnotesize}
+\input{derivatives}
 

### 2019-04-23 paper/ms.tex
Started numerical results section

@@
 slope limiter \citep[see, e.g.,][]{cockburn:1998} in conjunction with the
-troubled cell indicator discussed in \citet{fu:2017} to prevent excessive limiting.
+troubled cell indicator discussed in \citet{fu:2017} to prevent excessive limiting
+by flagged elements where limiting is needed.
 The general $s$-stage Runge-Kutta time stepping algorithm, including
@@
 \end{itemize}
-Above, the TVD limiter and the `positivity' limiter preventing unphysical
-states are denoted by the operators $\Lambda^{\TVD}\{\}$
-and $ \Lambda^{\Pos}\{\}$, respectively. The SSP-RK3 coefficients $\alpha_{ij}$
-and $\beta_{ij}$ may be found in Table~2.1 in \citet{cockburn:2001}.
+Above, the TVD limiter limiter preventing unphysical
+states is denoted by the operators $\Lambda^{\TVD}\{\}$ The SSP-RK3 coefficients $\alpha_{ij}$
+and $\beta_{ij}$ may be found in Table~2.1 in \citet{cockburn:2001}. For each step
+in Eq.~\eqref{eq:rkStages}, the TVD limiter is applied to elements flagged
+by the troubled cell indicator. For a more in-depth discussion of the
+limiting process used in \thornado, see \citet{endeve:2019}.
 
@@
 to extend to the nuclear matter EOS case.
-
-\section{Numerical Results}
-\label{sec:results}
-figure figure figure.
-Sod: Definitely want a $y_{e}$ curve, with and without
-characteristic limiting. Perhaps density too, or something. Show comparison
-to the nX=10000 run.
-
-Maybe also do the Noh problem, include results from there.
 
@@
 \end{align*}
-where $\phi_{i} = P_{\epsilon}\,\tau\, v_{i}$ and
-$\omega = \tau\, (P_{\epsilon}\,(v^2 - 2\epsilon) - 2\,P_{\tau}\,\tau)$.
+where $\phi_{i} = P_{\epsilon}\,\tau\, v_{i}$,
+$\omega = \tau\, (P_{\epsilon}\,(v^2 - 2\epsilon) - 2\,P_{\tau}\,\tau)$, and
+$\alpha = 2 y_{e} P_{D_{e}} - \tau \chi$.
 
 \todo{What is the physical interpretation of h and k? I tried to follow \citet{schaal:2015a},
-and in their eigenvectors (Ideal), h is specific enthalpy and k is specific kinetic energy.}
+and in their eigenvectors (Ideal), h is specific enthalpy and k is specific kinetic energy.
+What about my other defined variables? Should I redefined, or rename?}
 The matrix of left eigenvectors is given by
+
+\section{Preliminary Numerical Results}
+\label{sec:results}
+In this section we present preliminary numerical results obtained with the DG
+method implemented in \thornado  leveraging use of the characteristic
+decomposition. We use the SFHo EOS of nuclear matter \citep{steiner:2013}
+now commonly used in hig-fidelity CCSN simulations.
+These tests serve to gauge the performance of the
+DG characteristic decomposition implementation on a set of benchmarks as an initial
+assessment of its suitability for future CCSN simulations.
+
+\subsection{Sod Shock Tube}
+We present the classic 1D Sod shock tube Riemann problem \citep{sod:1978} in
+Cartesian coordinates. The computational domain is D = [-5,5] km
+with a discontinuity initially located at $x = 0$ km separating the left and right states
+
+\beq
+  \mathbf{U}_{L} = (10^{12} g/cm^3, 0\, km/s, 10^{32} ergs/cm^3, 0.4\rho)^T\,\,\,
+  \mathbf{U}_{R} = (1.25*10^{11} g/cm^3, 0\, km/s, 10^{31} ergs/cm^3, 0.3\rho)^T
+\eeq
+\footnote{How to specify this? Specifically $y_e$}
+
+The test is run until $t = 0.25$ms with 100 elements
+\\
+\todo{Componentwise vs Characteristic vs reference.}\\
+Plots
+\todo{Limiting Paramter Landscape.}\\
+plots.
+\todo{"Best" Parameters}\\
+Plots
+\todo{"Realistic" high density run}\\
+Plots
 

### 2019-05-01 paper/ms.tex
First Draft?

--- 2019-04-11 paper/ms.tex
@@
 \usepackage{verbatim}
-\usepackage{placeins}
+%\usepackage{placeins}
 \graphicspath{{./figures/}}
@@
 
+\usepackage{placeins}
+
 \bibliographystyle{apj}
@@
 \newcommand{\slugcom}{Draft - \today}
-\newcommand{\Pos}{\textnormal{\tiny\textsc{Pos}}}
 \newcommand{\TVD}{\textnormal{\tiny\textsc{Tvd}}}
+\newcommand{\TCI}{\textnormal{\tiny\textsc{TCI}}}
+
 %\newcommand{\slugcom}{Submitted to ApJL on \subdate}
@@
 
-%\begin{abstract}
-%\end{abstract}
-% nonlinear hyperbloic cons. laws. develop discontinuities from smooth ICs
-% this can lead to oscillatins which require some limiting process to smooth.
-% limiting in characteristic variables increases the efficiency of slope limiting
 \begin{abstract}
@@
 The toolkit for high-order neutrino-radiation hydrodynamics (\thornado)
-is being developed for CCSNe simulations and related problems utilizing a spatial discretization
+is being developed for CCSN simulations and related problems utilizing a spatial discretization
 based on the discontinuous Galerkin (DG) method.
@@
 with the inclusion of the nuclear EOS terms and
-present results from some initial tests.
+present results from some initial tests. The results confirm that performing limiting
+on the characteristic variables provides better numerical solutions and is
+suited to applications in CCSN simulations.
 \end{abstract}
 
-\keywords{supernovae: general -- hydrodynamics -- convection -- turbulence -- equation of state -- methods: numerical -- stars: massive -- stars: evolution}
+\keywords{supernovae: general -- hydrodynamics -- equation of state -- methods: numerical -- discontinuous Galerkin}
 
@@
 
-\subsection{Time Integration}
-A common feature of high-order numerical PDEs is unphysical oscillations in the solutions.
+\subsection{Slope Limiting}
+\label{sec:Limiting}
+A common feature of high-order numerical PDEs is unphysical oscillations in the solutions,
+particularly near discontinuities.
 It is therfore of great interest in the DG algorithm to implement slope limiting
@@
 slope limiter \citep[see, e.g.,][]{cockburn:1998} in conjunction with the
-troubled cell indicator discussed in \citet{fu:2017} to prevent excessive limiting.
+troubled cell indicator discussed in \citet{fu:2017} to prevent excessive limiting
+by flagging elements where limiting is needed. To do this, we need to reduce under- and
+overshootings of the higher-order solution at cell boundaries compared to
+the cell averages of neighbor cells. Recall from Eq.~\eqref{eq:conservedNodalExpansion}
+that in each cell our solution is expressed as
+
+\beq
+\mathbf{U}_h (x,t) =
+\sum_{l=1}^{N}c_{l}(t)\,P_{l}(x)
+\eeq
+
+\noindent where we have chosen to expand in the Legendre polynomials. We compare
+the weight $c_{1}$, which is proportional to the first derivative of the
+solution in the cell, to the neighboring cell averages by the following
+
+\beq
+  \widetilde{c}_{1} = \text{minmod}(c_{1},
+    \beta_{\TVD}(c_{0}^{+} - c_{0}), \beta_{\TVD}(c_{0} - c_{0}^{-}))
+  \label{eq:limiting}
+\eeq
+
+\noindent where $\widetilde{c}_{1}$ is the limited weight. Here, we used the minmod
+function defined as
+
+\beq
+  \text{minmod}(a_{1}, a_{2}, a_{3}) =
+   \begin{cases}
+      s\,\, \text{min}\{\abs{a_{1}}, \abs{a_{2}}, \abs{a_{3}}\} &
+        s = \text{sign}(a_{1})= \text{sign}(a_{2})= \text{sign}(a_{3}) \\
+      0 & \text{otherwise.}
+   \end{cases}
+\label{eq:minmod}
+\eeq
+
+Limiting is applied to each component of the vector of conserved quantities
+separately and is generally referred to as `Componentwise limiting.`
+The parameter $\beta_{\TVD}$ takes values in the closed interval $\left[1,2\right]$
+and scales the strength of the limiting. A minimal $\beta_{\TVD}$ corresponds with
+the total variation diminishing scheme \textcolor{red}{reference?} but is more
+dissipative than the maximal $\beta_{\TVD}$ case, which is more flexible but allows for more oscillations.
+Increasing $\beta_{\TVD}$ puts more weight on the neighboring cell averages,
+making the minmod function more likely to set $ \widetilde{c}_{1} = c_{1}$.
+
+In order to determine where slope limiting is necessary, we use the
+troubled cell indicator (TCI) \citep{fu:2017} to prevent excessive limiting:\footnote{Would this notation be altered for 1D?}
+
+\begin{equation}
+  I_{\bK}(G) = \f{\sum_{j}|G_{\bK}-G_{\bK}^{(j)}|}{\max_{j}|G_{\bK^{(j)}}^{(j)}|},
+  \label{eq:tci}
+\end{equation}
+
+\noindent where $G\in\vect{G}\subseteq\vect{U}$ and $\vect{G}$ consists of the
+first and final elements of $\vect{U}$. Here, the sum in the numerator and the
+max in the denominator are taken over the neighboring elements sharing a
+boundary with target element $\bK$, $G_{\bK}$ is the cell average in $\bK$,
+$G_{\bK}^{(j)}$ is the cell average
+computed by extrapolating the polynomial representation from the neighboring
+element $\bK^{(j)}$ into $\bK$, and $G_{\bK^{(j)}}^{(j)}$
+is the cell average native to neighbor element $\bK^{(j)}$. An element is flagged
+for limiting if, for any $G\in\vect{G}$, $I_{\bK}(G)>C_{\TCI}$, where
+$C_{\TCI}$ is a defined threshold.
+
+\subsection{Time Integration}
+\label{sec:TimeInt}
 The general $s$-stage Runge-Kutta time stepping algorithm, including
-the limiting process, can then be summarized as in \citet{cockburn:2001}:
+the limiting process, can be summarized as in \citet{cockburn:2001}:
 \begin{itemize}
@@
 \end{itemize}
-Above, the TVD limiter and the `positivity' limiter preventing unphysical
-states are denoted by the operators $\Lambda^{\TVD}\{\}$
-and $ \Lambda^{\Pos}\{\}$, respectively. The SSP-RK3 coefficients $\alpha_{ij}$
-and $\beta_{ij}$ may be found in Table~2.1 in \citet{cockburn:2001}.
+Above, the TVD limiter limiter preventing unphysical
+states is denoted by the operators $\Lambda^{\TVD}\{\}$. The SSP-RK3 coefficients $\alpha_{ij}$
+and $\beta_{ij}$ may be found in Table~2.1 in \citet{cockburn:2001}. For each step
+in Eq.~\eqref{eq:rkStages}, the TVD limiter is applied to elements flagged
+by the troubled cell indicator.\textcolor{red}{Discuss timestepping conditions?}
 
 \subsection{Characteristic Decomposition}
+\label{sec:characteristicDecomp}
 Experience has shown that the slope limiting described in the previous section
@@
 $\beta = \frac{1}{2} (\delta_{1}+2 \epsilon +\frac{2 P_{\tau} \tau }{P_{\epsilon}})$.
+The matrix of left eigenvectors is given by
+
 \begin{align*}
@@
 \end{align*}
-where $\phi_{i} = P_{\epsilon}\,\tau\, v_{i}$ and
-$\omega = \tau\, (P_{\epsilon}\,(v^2 - 2\epsilon) - 2\,P_{\tau}\,\tau)$.
+where $\phi_{i} = P_{\epsilon}\,\tau\, v_{i}$,
+$\omega = \tau\, (P_{\epsilon}\,(v^2 - 2\epsilon) - 2\,P_{\tau}\,\tau)$, and
+$\alpha = 2 y_{e} P_{D_{e}} - \tau \chi$.
 
-\todo{What is the physical interpretation of h and k? I tried to follow \citet{schaal:2015a},
-and in their eigenvectors (Ideal), h is specific enthalpy and k is specific kinetic energy.}
-The matrix of left eigenvectors is given by
+The left and right eigenvectors presented here are similar in form to those
+presented in \citet{schaal:2015}, but are considerably messier due to the
+restriction of a nuclear EOS.
+
+\section{Preliminary Numerical Results}
+\label{sec:results}
+In this section we present preliminary numerical results obtained with the DG
+method implemented in \thornado\, leveraging use of the characteristic
+decomposition. Unless otherwise noted, we use the SFHo EOS of
+nuclear matter \citep{steiner:2013} now commonly used in high-fidelity CCSN simulations.
+These tests serve to gauge the performance improvement of the
+DG characteristic decomposition implementation on a set of benchmarks as an initial
+assessment of its suitability for future CCSN simulations.
+
+\subsection{Sod Shock Tube}
+We present the classic 1D Sod shock tube Riemann problem \citep{sod:1978} in
+Cartesian coordinates. The computational domain is D = [-5,5] km
+with a discontinuity initially located at $x = 0$ km separating the left and right states
+
+\begin{align}
+  \mathbf{U}_{L} &= (10^{12}~\text{g~cm}^{-3}, 0\,, 3.712*10^{32}~\text{ergs~cm}^{-3}, 0.4*10^{12})^T\,\,\, \\
+  \mathbf{U}_{R} &= (1.25*10^{11}~\text{g~cm}^{-3}, 0\, , 3.015*10^{31}~\text{ergs~cm}^{-3}, 0.375*10^{11})^T.
+\end{align}
+\footnote{\textbf{Fix e in the state vectors}}
+\noindent The test is run until $t = 0.025$ms with 100 elements
+using $C_{\TCI} = 0.2$ and $\beta_{\TVD} = 2.0$. Results are plotted in
+Figure~\ref{fig:SodSedovOptimal}. Using the characteristic limiting described
+above, the DG method captures the features of the solution without introducing
+noticeable oscillations in the solutions near the discontinuities. In Figure~\ref{fig:shock}
+we plot elements in the $xt$-plane flagged by the troubled cell indicator for
+limiting.
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=32pc,height=30pc]{optimal.png}
+  \centering
+  \caption{\label{fig:SodSedovOptimal} Numerical solution of Sod's problem using
+    100 elements compared with a reference solution using 10000 elements.}
+\end{figure}
+%\clearpage
+\begin{wrapfigure}{L}{2.8in}
+  \includegraphics[width=18pc]{optimal_shock.png}
+  \caption{$xt$-plane of elements flagged for limiting by the troubled cell indicator.}
+  \label{fig:shock}
+\end{wrapfigure}
+
+\subsection{Optimal Limiting Parameters}
+\label{sec:param}
+In order to determine the optimal limiting parameters $C_{\TCI}$ and $\beta_{\TVD}$,
+we performed a suite of sixteen simulations of Sod's problem with various
+limiting parameters with $\beta_{\TVD}\in\left[1.0,2.0\right]$ and
+$C_{\TCI}\in\left[0.0,0.2\right]$. For each $\left(\beta_{\TVD},C_{\TCI}\right)$ pair,
+we computed the relative error and total variation \textbf{\textcolor{red}{EQUATIONS}} in density and electron
+fraction at $t = 0.025$ms. Results are plotted in Figure~\ref{fig:landscape}.
+We find that, as expected, increasing $\beta_{\TVD}$ tends to monotonically
+decrease the relative error while (non-monotonically) increasing the total variation due to the
+limiter allowing for more oscillations. Increasing $C_{\TCI}$ tends to decrease
+the relative error and increase the total variation.
+We have selected
+$\beta_{\TVD}=1.75$ and $C_{\TCI}=0.1$ to be the optimal parameters providing
+the best combination of relative error and total variation reduction. Unless
+otherwise noted, all following results will use this combination of limiting
+parameters.
+
+\begin{figure}[b]
+  \centering
+  \includegraphics[width=32pc, height=27.5pc]{./figures/error_tv_landscape.png}
+  \caption{\label{fig:landscape} Error - total variation landscapes for various
+  values of the limiting parameters $C_{TCI}$ and $\beta_{TVD}$.}
+\end{figure}
+
+\subsection{Improvement From Componentwise Limiting}
+\label{sec:optimal}
+The motivation for limiting on the characteristic variables is the potential
+improvement from componentwise limiting. In Figure~\ref{fig:SodSedovCW} we plot the
+solution at $t = 0.025$ms using 100 elements employing both characteristic
+limiting (blue) and componentwise limiting (red) compared to a reference
+solution computed with 10000 elements. We observe, for the componentwise limiting,
+noticable oscillations in density around the contact discontinuity as well
+as less resolved discontinuities. The electron fraction displays very
+large oscillations at the left side of the contact discontinuity, up to
+$\pm 0.02$ in the cell averaged electron fraction. This makes characteristic
+limiting particularly appealing for CCSN simulations where the electron fraction
+plays a critical role in the explosion dynamics.
+
+% \begin{figure}[h]
+%   \centering
+%   \includegraphics[width=32pc, height=32pc]{./figures/error_tv_landscape.png}
+%   \caption{\label{fig:landscape} Error - total variation landscapes for various
+%   values of the limiting parameters $C_{TCI}$ and $\beta_{TVD}$.}
+% \end{figure}
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=36pc]{./figures/characteristic_cw.png}
+  \caption{\label{fig:SodSedovCW} Numerical solution of Sod's problem using
+  100 elements with characteristic limiting (blue) and componentwise limiting
+  (red) compared with a reference solution using 10000 elements.}
+\end{figure}
+
+\subsection{High Density Regime}
+This test is included to verify the performance improvements of the characteristic
+limiting in a higher density regime. It is designed with the Sod problem in mind.
+The computational domain is $D = [-5,5]$ km
+with a discontinuity initially located at $x = 0$ km separating the left and right states
+
+\begin{align}
+  \mathbf{U}_{L} &= (10^{13}~\text{g~cm}^{-3}, 0\,, 3.712*10^{32}~\text{ergs~cm}^{-3}, 0.15*10^{12})^T\,\,\, \\
+  \mathbf{U}_{R} &= (1.25*10^{12}~\text{g~cm}^{-3}, 0\, , 3.015*10^{31}~\text{ergs~cm}^{-3}, 0.169*10^{12})^T.
+\end{align}
+
+\noindent The test is run until $t = 0.05$ms with 100 elements
+using $C_{\TCI} = 0.1$ and $\beta_{\TVD} = 1.2$. We chose a lower $\beta_{\TVD}$
+here, as larger values resulted in internal energies outside of the
+EOS table causing the test to fail, motivating the development of a
+positivity limiter compatible with a tabulated nuclear matter EOS.
+Inital state values were chosen to be consistent with
+high-fidelity CCSN simulation code \chimera \citep{bruenn:2018} data.
+Results are plotted in
+Figure~\ref{fig:SodSedovHighDens}. Our method captures the main features
+of the solution, but introduces more oscillations than in the lower density
+case, even when using the characteristic limiter.
+We notice an undershooting of the right side
+of the contact discontinuity in the electron fraction not previously present
+warranting further investigation.
+
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=36pc]{./figures/highDens.png}
+  \caption{\label{fig:SodSedovHighDens} Results for the high density Sod-like problem
+  computed using 100 elements and $C_{\TCI} = 0.1$ and $\beta_{\TVD} = 1.2$ at $t=0.05$ms.}
+\end{figure}
+
+\subsection{EOS Resolution Dependence}
+\label{sec:eosRes}
+In order to test the sensitivity of the limiter on the EOS table resolution,
+we repeated the tests in Section~\ref{sec:optimal} with a higher resolution
+SFHo EOS table. As before, tests are computed using 100 elements until
+$t = 0.025$ms with characteristic limiting and the optimal limiting parameters determined in
+Section~\ref{sec:param} and compared to a reference solution computed using
+10000 elements. Results for both tables are plotted in Figure~\ref{fig:SodSedovSFHoRes}.
+We find nearly no sensitivity to the table resolution, with the exception of
+minor differences in density to the left of the contact discontinuity.
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=36pc]{./figures/eos_res.png}
+  \caption{\label{fig:SodSedovSFHoRes} Comparison of numerical solutions to
+  Sod's problem using 100 elements with characteristic limiting and two
+  different EOS table resolutions.}
+\end{figure}
+
+\subsection{EOS Table Sensitivity}
+In order to probe the sensitivity of the method to the EOS table, we repeated
+the tests conducted in Section~\ref{sec:optimal} using three different EOS tables:
+the high resolution SFHo table from Section~\ref{sec:eosRes}, the SFHx table
+(\textbf{citation}), and the DD2 table (\textbf{citation}).
+As before, tests are computed using 100 elements until
+$t = 0.025$ms with characteristic limiting and the optimal limiting parameters determined in
+Section~\ref{sec:param}. We find nearly no sensitivity to the EOS table used in
+the low density regime, with the exception of very small variations in density
+to the left of the contact discontinuity. Future work includes testing the
+sensitivity to the EOS table in higher density regimes.
+\begin{figure}[h!]
+  \centering
+  \includegraphics[width=36pc]{./figures/eos_all.png}
+  \caption{\label{fig:SodSedovEOS} Comparison of numerical solutions to
+  Sod's problem using 100 elements with characteristic limiting and three
+  different EOS tables.}
+\end{figure}
+%\FloatBarrier
+
+\section{Conclusions}
+
+We have presented preliminary developments and numerical results for a
+characteristic limiter to be used with solvers of the non-relativistic
+Euler equations of gas dynamics in the {\bf t}oolkit for {\bf h}igh-{\bf or}der
+{\bf n}eutrino-r{\bf ad}iation hydr{\bf o}dynamics (\thornado). The results
+presented from a suite of 1D test problems, based around Sod's 1D shock
+tube problem, demonstrate the performance of the characteristic limiter
+compared to the componentwise limiter. Moreover, after performing a parameter
+study on the limiter parameters, we found optimal limiting
+parameters for use with the Sod problem. While the optimal limiting parameters
+tend to be problem dependent, we hope to investigate if such an optimal
+parameter exists for CCSN applications. Tests with various EOS tables and
+table resolutions showed little to no sensitivity to the table or its
+resolution in the tests studied. For future work, we will extend these studies
+to other higher density regimes applicable to the CCSN environment.
+Planned near-future work on the nuclear equation of state compatibility of
+\thornado\, includes the development of a positivity limiter and extension of
+the characteristic limtier to handle relativistic and curvilinear problems.
+
+\acknowledgements
+We thank people.
+
+\software{
+  \href{https://matplotlib.org/}{Matplotlib} \citep{hunter:2007a},
+  \href{http://www.numpy.org/}{NumPy} \citep{walt:2011},
+  \href{https://www.scipy.org/}{SciPy} \citep{jones:2001},
+  }
 
@@
 
-\acknowledgements
-We thank people.
-
 \bibliography{ms}

### 2019-05-01 paper/ms.tex
Draft: Sent.

@@
 \label{sec:Intro}
-\todo{Fill in introduction. Contextualize the work.}
 The core collapse supernovae (CCSNe) explosion mechanism is fundamentally three-dimensional
 in nature \citep[see e.g.,][]{blondin:2006, muller:2012, oconnor:2018b}. Alongside general
-reletavistic gravity, complex nuclear equations of state (EOS), and
+reletavistic gravity, complex nuclear matter equations of state (EOS), and
 neutrino transport, hydrodynamics must be accurately modelled,
@@
 \citep{murphy:2011, murphy:2013, couch:2013, couch:2015a, radice:2016, mabanta:2018}.
-The accurate and efficient modelling of the supernova hydrodynamics is crucial if the
+The accurate and efficient modelling of the supernova hydrodynamics is imperative if the
 explosion is to be realistically modelled.
-For in-depth reviews of the CCSN mechanism, see \citet{bethe:1990, janka:2007, janka:2012a, janka:2016, burrows:2013, hix:2014, muller:2016, couch:2017}.
+For in-depth reviews of the CCSN mechanism, see
+\citet{bethe:1990, janka:2007, janka:2012a, janka:2016, burrows:2013, hix:2014, muller:2016, couch:2017}.
 
-\todo{Talk about \thornado, DG methods, Euler Equations, possibly reference other codes that
-employ different methods and those methods' drawbacks.}
+Discontinuous Galerkin (DG) methods have been shown by others
+\citep[see e.g.,][]{radice:2011, schaal:2015a, zanotti:2015, dumbser:2018} to have a
+high potential for apllications to astrophysical problems. They are a desireable
+choice for modelling fluid flows in CCSNe. These methods combine elements of
+spectral and finite volume methods, and achieve high-order accuracy
+on a compact stencil.
 
-This paper will be organized as follows: in Section \ref{sec:eulereq} we briefly
-describe the Euler equations of gas dynamics. Section \ref{sec:DG} describes our
+The {\bf t}oolkit for {\bf h}igh-{\bf or}der {\bf n}eutrino-r{\bf ad}iation hydr{\bf o}dynamics (\thornado) is being developed to simulate neutrino-radiation hydrodynamics in CCSNe and related applications in nuclear astrophysics.
+The spatial discretization of solvers for hyperbolic partial differential equations in \thornado\, is based on the DG method.
+Whether the high-order approach will improve accuracy and efficiency of CCSN models remains to be demonstrated.
+In this paper we provide an initial description and encouraging demonstration
+of the DG method implemented in \thornado\, to solve the non-relativistic Euler equations.
+We focus on basic one-dimensional tests focusing on the implementation of limiters in \thornado.
+
+The DG slope limiting techniques implemented in \thornado\, are similar to those
+used by \cite{schaal:2015a}, but is supplemented with a troubled-cell
+indicator \cite{fu:2017} to prevent excessive limiting (e.g., around smooth extrema).
+It has been shown \citep[see e.g.,][]{schaal:2015a, cockburn:1998, cockburn:1989}
+that limiting on the chracteristic variables, which represent advected quantities,
+shows an improvement over limiting on the conserved variables.
+In this paper, we develop a chracteristic limiter for use with a general tabulated
+nuclear matter EOS in \thornado.
+
+This paper will be organized as follows: in Section~\ref{sec:eulereq} we briefly
+describe the Euler equations of gas dynamics. Section~\ref{sec:DG} describes our
 numerical implementation of the DG method in the one dimensional (1D) case.
-Section \ref{sec:results} shows the results of 1D hydrodynamic simulations,
-and Section \ref{sec:eigen} presents the eigenvectors used in the eigen-decomposition
-of the Euler equations.
+Section~\ref{sec:eigen} presents the eigenvectors used in the eigen-decomposition
+of the Euler equations, and Section~\ref{sec:results}
+shows the results of preliminary 1D hydrodynamic simulations.
 
@@
 \label{sec:eulereq}
-\todo{Use this form of the Euler Equations?}
 The non-relativistic Euler equations of gas dynamics
@@
 \beq
-  \partial_{t} D + \divergence{}(D\, \mathbf{v}) = 0
+  \partial_{t} D_{e} + \divergence{}(D_{e}\, \mathbf{v}) = 0
   \label{eq:electronConservation}
@@
 where $\rho$ represents mass density, $\mathbf{v}$ the fluid velocity vector,
-$P$ the fluid pressure, $D=\rho y_e$ where $y_e$ is the electron fraction,
+$P$ the fluid pressure, $D_{e}=\rho y_e$ where $y_e$ is the electron fraction,
 $E=\epsilon \rho +\frac{1}{2}\rho v^2$ the total energy (internal plus kinetic),
@@
 
-\todo{Add intro paragraph} \\
-\textcolor{red}{Subsections or no?}
-
 \subsection{The Discontinuous Galerkin Methods}
@@
 we seek solutions to the Euler equations of hydrodynamics, constituting a
-conservation law of the form
+hyperbolic conservation law of the form
 
@@
 the weight $c_{1}$, which is proportional to the first derivative of the
-solution in the cell, to the neighboring cell averages by the following
+solution in the cell, to the neighboring cell averages by the following \\
 
@@
 
-\noindent where $\widetilde{c}_{1}$ is the limited weight. Here, we used the minmod
-function defined as
+\noindent where $\widetilde{c}_{1}$ is the limited weight. The superscripts $-/+$
+indicate the cell averages of the neighboring cells to the  immediate left/right
+of the interface.
+Here, we use the minmod function defined as
 
@@
 states is denoted by the operators $\Lambda^{\TVD}\{\}$. The SSP-RK3 coefficients $\alpha_{ij}$
-and $\beta_{ij}$ may be found in Table~2.1 in \citet{cockburn:2001}. For each step
+and $\beta_{ij}$ may be found in Table~2.1 in \citet{cockburn:2001a}. For each step
 in Eq.~\eqref{eq:rkStages}, the TVD limiter is applied to elements flagged
-by the troubled cell indicator.\textcolor{red}{Discuss timestepping conditions?}
+by the troubled cell indicator (where $I_{\bK} > C_{\TCI}$).
+This algorithm is subject to a timestep stability condition $\Delta t_{\TVD}$.
+Let $(h/|\lambda|)=\min_{i\in\{1,\ldots,d\}}(\Delta x^{i}/|\lambda^{i}|)$, where $|\lambda^{i}|$ is the largest absolute eigenvalue of the flux Jacobian.
+Then, assuming Cartesian coordinates (e.g., \cite{cockburn:2001a}):
+\begin{equation}
+  \Delta t_{\TVD} \le \f{1}{d}\,\f{(h/|\lambda|)}{(2k+1)} .
+  \label{eq:timestepRestrictions}
+\end{equation}
+For $d=k=2$, the effective Courant-Friedrichs-Lewy (CFL) factor is $0.1$.
 
@@
 $\mathbf{w} = \mathcal{R}^{-1}\mathbf{U}$. Multiplying both sides of Equation
-\eqref{eq:charEq} by $\mathcal{R}^{-1}$, we are able to linearize the system of equations to
+\eqref{eq:charEq} by $\mathcal{R}^{-1}$ and simplifying, we are able to linearize the system of equations to
 a system of advection equations
@@
   \Lambda \pderiv{\mathbf{w}}{x}
-  = 0.
+  = 0. \\
 \eeq
@@
 process of characteristic limiting has been done for the ideal EOS
-\citep[][]{cockburn:1998}, we want
-to extend to the nuclear matter EOS case.
+\citep[][]{cockburn:1998}, we want to extend this process to the case of a
+tabulated nuclear matter EOS.
 
@@
 
-We assume that $P = P(\tau, \epsilon, n)$, where $\tau = \frac{1}{\rho}$.
-Let the vector of conserved variables be $\textbf{U} = \{\rho, m_1, m_2, m_3, E, D\}$,
-where $m_i = \rho v_i$ and $D = \rho y_{e}$.
+We assume that pressure $P = P(\tau, \epsilon, n)$, where $\tau = \frac{1}{\rho}$.
+Let the vector of conserved variables be $\textbf{U} = \{\rho, m_1, m_2, m_3, E, D_{e}\}$,
+where $m_i = \rho v_i$ and $D_{e} = \rho y_{e}$.
 The flux vector is $\textbf{F}(\textbf{U}) =
@@
 
-\noindent Where $H=(E+P)\tau$ is the specific enthalpy and $P_{x} = \partial_{x}P$\footnote{Maybe define these elsewhere to show what's constant}. The eigenvalues of this Jacobian are given by
+\noindent Where $H=(E+P)\tau$ is the specific enthalpy and $P_{x} = \partial_{x}P$\footnote{Maybe define these elsewhere to show what's constant}.
+The eigenvalues of the flux Jacobian are given by the diagonal matrix
 
@@
 where $c_{s} = \sqrt{\Gamma P \tau}$, for
-$\Gamma = \frac{\tau (P P_{\epsilon} - P_{\tau}) + P_{D_e} y_{e} \tau^{-1}}{P}$, is
-the local sound speed. In the ideal EOS regime this reduces to the sound speeds given \citet{colella:1985}.
-The right eigenvectors are then given by
+$\Gamma = \left(\tau (P P_{\epsilon} - P_{\tau}) + P_{D_e} y_{e} \tau^{-1}\right) P^{-1}$, is
+the local sound speed. In the ideal EOS regime this reduces to the sound speeds given in \citet{colella:1985}.
+The right eigenvectors are then given by the column vectors of the following matrix
 \begin{align*}
@@
 $\beta = \frac{1}{2} (\delta_{1}+2 \epsilon +\frac{2 P_{\tau} \tau }{P_{\epsilon}})$.
-The matrix of left eigenvectors is given by
+The left eigenvectors are given by the row vectors of the inverse matrix $\mathcal{L}_{1} = \mathcal{R}_{1}^{-1}$
 
@@
 $\alpha = 2 y_{e} P_{D_{e}} - \tau \chi$.
-
 The left and right eigenvectors presented here are similar in form to those
-presented in \citet{schaal:2015}, but are considerably messier due to the
+presented in \citet{schaal:2015a}, but are considerably messier due to the
 restriction of a nuclear EOS.
@@
 We present the classic 1D Sod shock tube Riemann problem \citep{sod:1978} in
-Cartesian coordinates. The computational domain is D = [-5,5] km
+Cartesian coordinates. This is an ideal first test as it is a fairly simple
+scenario but contains both a contact discontinuity and a shock to test the
+efficiency of our limiting scheme. The computational domain is D = [-5,5] km
 with a discontinuity initially located at $x = 0$ km separating the left and right states
 
-\begin{align}
+\begin{align*}
   \mathbf{U}_{L} &= (10^{12}~\text{g~cm}^{-3}, 0\,, 3.712*10^{32}~\text{ergs~cm}^{-3}, 0.4*10^{12})^T\,\,\, \\
   \mathbf{U}_{R} &= (1.25*10^{11}~\text{g~cm}^{-3}, 0\, , 3.015*10^{31}~\text{ergs~cm}^{-3}, 0.375*10^{11})^T.
-\end{align}
+\end{align*}
 \footnote{\textbf{Fix e in the state vectors}}
@@
 Figure~\ref{fig:SodSedovOptimal}. Using the characteristic limiting described
-above, the DG method captures the features of the solution without introducing
-noticeable oscillations in the solutions near the discontinuities. In Figure~\ref{fig:shock}
+above, the DG method captures the features of the solution,
+resolving well both the contact discontinuity and shock, without introducing
+noticeable oscillations in the numerical solutions near the discontinuities.
+In Figure~\ref{fig:shock}
 we plot elements in the $xt$-plane flagged by the troubled cell indicator for
-limiting.
+limiting, showing that the troubled cell indocator flagged cells around both
+the shock and contact discontinuity for limiting and maintained that
+for the full time of the test.
 
@@
 
-\begin{align}
+\begin{align*}
   \mathbf{U}_{L} &= (10^{13}~\text{g~cm}^{-3}, 0\,, 3.712*10^{32}~\text{ergs~cm}^{-3}, 0.15*10^{12})^T\,\,\, \\
   \mathbf{U}_{R} &= (1.25*10^{12}~\text{g~cm}^{-3}, 0\, , 3.015*10^{31}~\text{ergs~cm}^{-3}, 0.169*10^{12})^T.
-\end{align}
+\end{align*}
 
@@
 using $C_{\TCI} = 0.1$ and $\beta_{\TVD} = 1.2$. We chose a lower $\beta_{\TVD}$
-here, as larger values resulted in internal energies outside of the
+here, as larger values resulted in unphysical internal energies outside of the
 EOS table causing the test to fail, motivating the development of a
@@
 Inital state values were chosen to be consistent with
-high-fidelity CCSN simulation code \chimera \citep{bruenn:2018} data.
+high-fidelity CCSN simulation code \chimera\, \citep{bruenn:2018} data.
 Results are plotted in

### 2019-05-05 paper/ms.tex
Working through final revisions. Final push.

@@
 
-\title{\thornado-hydro: Generalizing discontinuous galerkin methods for a nuclear
+\title{\thornado-hydro: Generalizing discontinuous Galerkin methods for a nuclear
 equation of state for supernova hydrodynamics}
@@
 The toolkit for high-order neutrino-radiation hydrodynamics (\thornado)
-is being developed for CCSN simulations and related problems utilizing a spatial discretization
+is being developed for core-collapse supernova (CCSN) simulations and related problems utilizing a spatial discretization
 based on the discontinuous Galerkin (DG) method.
@@
 This use of characteristic variables also increases the efficiency of the slope
-limiting process. However, the addition of the nuclear equation of state equation
-to the Euler equations makes the decomposition nontrivial.
+limiting process for high-order DG methods. However, the addition of a
+tabulated nuclear equation of state to the Euler equations makes the decomposition nontrivial.
 We introduce the framework for the characteristic decomposition of the Euler equations
 with the inclusion of the nuclear EOS terms and
-present results from some initial tests. The results confirm that performing limiting
-on the characteristic variables provides better numerical solutions and is
+present results from some initial tests using a third-order scheme.
+The results confirm that performing limiting
+on the characteristic variables provides better numerical solutions and is perhaps
 suited to applications in CCSN simulations.
@@
 \citep[see e.g.,][]{radice:2011, schaal:2015a, zanotti:2015, dumbser:2018} to have a
-high potential for apllications to astrophysical problems. They are a desireable
+high potential for applications to astrophysical problems. They are a desireable
 choice for modelling fluid flows in CCSNe. These methods combine elements of
@@
 
-The {\bf t}oolkit for {\bf h}igh-{\bf or}der {\bf n}eutrino-r{\bf ad}iation hydr{\bf o}dynamics (\thornado) is being developed to simulate neutrino-radiation hydrodynamics in CCSNe and related applications in nuclear astrophysics.
+The {\bf t}oolkit for {\bf h}igh-{\bf or}der {\bf n}eutrino-r{\bf ad}iation hydr{\bf o}dynamics
+(\thornado)\footnote{\url{https://github.com/endeve/thornado}}is being developed to simulate neutrino-radiation
+hydrodynamics in CCSNe and related applications in nuclear astrophysics.
 The spatial discretization of solvers for hyperbolic partial differential equations in \thornado\, is based on the DG method.
@@
 used by \cite{schaal:2015a}, but is supplemented with a troubled-cell
-indicator \cite{fu:2017} to prevent excessive limiting (e.g., around smooth extrema).
+indicator \citep{fu:2017} to prevent excessive limiting (e.g., around smooth extrema).
 It has been shown \citep[see e.g.,][]{schaal:2015a, cockburn:1998, cockburn:1989}
-that limiting on the chracteristic variables, which represent advected quantities,
-shows an improvement over limiting on the conserved variables.
-In this paper, we develop a chracteristic limiter for use with a general tabulated
-nuclear matter EOS in \thornado.
+that limiting on the chracteristic variables (characteristic limiting),
+which represent advected quantities, shows an improvement over limiting
+directly on the conserved variables (componentwise limiting).
+In this paper, we develop a chracteristic
+limiter for use with a general tabulated nuclear matter EOS in \thornado.
+In the case of the ideal EOS, the eigenvectors and eigenvalues which diagonalize
+the flux Jacobian, which are necessary to transform to the characteristic variables,
+are known analytically. All of the thermodynamic derivatives have analytic forms.
+The inclusion of a tabulated nuclear EOS complicates matters, as derivatives
+must be appriximated by interpolating from the EOS table and the diagonalization
+becomes non-trovial.
+We provide explicit analytic expressions for the matrices which diagonalize the
+flux Jacobian of the Euler equations.
 
-This paper will be organized as follows: in Section~\ref{sec:eulereq} we briefly
-describe the Euler equations of gas dynamics. Section~\ref{sec:DG} describes our
+This paper is organized as follows: in Section~\ref{sec:eulereq} we briefly
+describe the Euler equations of gas dynamics including a conservation law
+appropriate for a nuclear EOS. Section~\ref{sec:DG} describes our
 numerical implementation of the DG method in the one dimensional (1D) case.
-Section~\ref{sec:eigen} presents the eigenvectors used in the eigen-decomposition
-of the Euler equations, and Section~\ref{sec:results}
-shows the results of preliminary 1D hydrodynamic simulations.
+Section~\ref{sec:eigen} presents the eigenvectors used in the diagonalization
+of the flux Jacobian of the Euler equations, and Section~\ref{sec:results}
+shows the results of preliminary 1D hydrodynamic Riemann problem similar to that
+of Sod \citep{sod:1978}.
 
@@
 The non-relativistic Euler equations of gas dynamics
-\citep[see, e.g.,][for details]{leveque:2002} in Cartesian coordinates
+\citep[see, e.g.,][for details in the case of an ideal EOS]{leveque:2002}
 in the absence of sources with a nuclear matter EOS are given by
@@
 \eeq
-and conservation of \textcolor{red}{electron density}
+and conservation of electrons
 \beq
@@
 \beq
-  \partial_{t}\mathbf{U} + \partial_{x}\mathbf{F}(\mathbf{U}) = 0,
+  \partial_{t}\mathbf{U} + \divergence{}\mathbf{F}(\mathbf{U}) = 0,
   \label{eq:conservation}
 \eeq
-where $\mathbf{U} =(\rho,\rho \mathbf{v},E, D)^{T}$ is the vector of conserved quantities
-and $\mathbf{F}(\mathbf{U})=(\rho \mathbf{v},\rho u^{2}+p,(E+P)\mathbf{v}, D\mathbf{v})^{T}$
+where $\mathbf{U} =(\rho,\rho \mathbf{v},E, D_{e})^{T}$ is the vector of conserved quantities
+and $\mathbf{F}(\mathbf{U})=(\rho \mathbf{v},
+\rho\, \mathbf{v} \otimes \mathbf{v} + P\, \mathbf{I},(E+P)\mathbf{v}, D_{e}\mathbf{v})^{T}$
 is the flux vector.
@@
 In our solver we have chosen the discontinuous Galerkin (DG) method
-for our spatial discretization. In this section we will briefly discuss our
+for our spatial discretization \citep[see e.g.,][]{schaal:2015, zhang:2010a, cockburn:1998, cockburn:1989}.
+In this section we will briefly discuss our
 implementation of the DG method, introducing notation and concepts.
@@
 \eeq
+where $\xL$ and $x_{R}$ are the left and right boundaries of the cell.
 We let the approximation space $\mathbb{V}^{k}$ for the DG method
-be constructed from the tensor product of one-dimensional
-polynomials of degree $k$. Note that functions in $\mathbb{V}^{k}$
+be polynomials of maximal degree $k$. Note that functions in $\mathbb{V}^{k}$
 can be discontinuous across element interfaces. The DG problem is then to find
@@
 
-In each element $K$, we use a nodal representation of the conserved variables
+In each element $\bK$, we use a nodal representation of the conserved variables
 $\mathbf{U}$:
 
-\beq
-\begin{split}
-  \mathbf{U}(x,t) \approx\, & \mathbf{U}_h (x,t) =
-  \sum_{i=1}^{N}\vect{U}_{i}(t)\,\ell_{i}(x),
-  \quad\text{where}\quad \\
-  & \ell_{i}(\eta)=
-  \prod_{\substack{j=1\\j\ne i}}^{N}\f{\eta-\eta_{j}}{\eta_{i}-\eta_{j}}
+\begin{equation}
+  \vect{U}(x,t)\approx
+  \vect{U}_{h}(x,t)=\sum_{i=1}^{N=k+1}\vect{U}_{i}(t)\,\ell_{i}(x),
+  \quad\text{where}\quad
+  \ell_{i}(\eta)=
+  \prod_{\substack{j=1\\j\ne i}}^{N=k+1}\f{\eta-\eta_{j}}{\eta_{i}-\eta_{j}}
   \label{eq:conservedNodalExpansion}
-\end{split}
-\eeq
+\end{equation}
+
 are Lagrange polynomials defined on $I = \{ \eta : \eta \in (-0.5,0.5) \}$,
@@
 so that $\vect{U}_{h}(x(\eta_{j}),t)=\vect{U}_{j}(t)$.
-First we define the $M$-point quadrature $Q_{M}:C^{0}(I)\to\mathbb{R}$
+We define the $M$-point quadrature $Q_{M}:C^{0}(I)\to\mathbb{R}$
 with abscissas $\hat{S}_{M}=\{\eta_{q}\}_{q=1}^{M}$ and weights
@@
 \beq
-  \f{1}{\Delta x}\int_{K}P_{h}(x)\,dx=\int_{I}P_{h}(\eta)\,d\eta=\sum_{q=1}^{M}w_{q}\,P_{h}(\eta_{q}).
+  \f{1}{\Delta x}\int_{\bK}P_{h}(x)\,dx=\int_{I}P_{h}(\eta)\,d\eta=\sum_{q=1}^{M}w_{q}\,P_{h}(\eta_{q}).
 \eeq
 
-For the sake of efficiency we let $M=N$ and $S_{N}=\hat{S}_{N}$, which is a
-spectral-type nodal collocation DG approximation \citep{bassi:2013}.
+We let $M=N$ and $S_{N}=\hat{S}_{N}$, which is a
+spectral-type nodal collocation DG approximation \citep{bassi:2013}, which is exact
+for Cartesian coordinates and Legendre-Gauss quadrature.
 Inserting Eq.~\eqref{eq:conservedNodalExpansion} into Eq.~\eqref{eq:semidiscreteDG},
@@
 \begin{align}
-  \pd{}{t}\int_{K}\vect{U}_{h}\,\phi\,dx
-  &\approx w_{k}\,\pd{}{t}\vect{U}_{k}\,\Delta x
+  \pd{}{t}\int_{\bK}\vect{U}_{h}\,\phi\,dx
+  &=w_{k}\,\pd{}{t}\vect{U}_{k}\,\Delta x
   \label{eq:timeDerivativeTerm}
 \end{align}
-for the time derivative piece.
+for the time derivative piece, where $\Delta x = x_{R} - \xL$.
 Similarly, the last term on the left-hand side of Eq.~\eqref{eq:semidiscreteDG}
@@
   \int_{K}\vect{F}(\vect{U}_{h})\,\pderiv{\phi}{x}\,dx
-  \approx \sum_{q=1}^{N}w_{q}\,\vect{F}(\vect{U}_{q})\,\pderiv{\ell_{k}}{\eta}(\eta_{q}).
+  = \sum_{q=1}^{N}w_{q}\,\vect{F}(\vect{U}_{q})\,\pderiv{\ell_{k}}{\eta}(\eta_{q}).
   \label{eq:volumeTerm}
@@
   \Big[\,
-    \widehat{\vect{F}}\ell_{k}\big|_{x_{H}}
+    \widehat{\vect{F}}\ell_{k}\big|_{x_{R}}
      -\widehat{\vect{F}}\ell_{k}\big|_{\xL}
@@
 \label{sec:Limiting}
-A common feature of high-order numerical PDEs is unphysical oscillations in the solutions,
-particularly near discontinuities.
+A common feature of high-order numerical PDEs is unphysical oscillations in the
+solutions around discontinuities.
 It is therfore of great interest in the DG algorithm to implement slope limiting
@@
 the cell averages of neighbor cells. Recall from Eq.~\eqref{eq:conservedNodalExpansion}
-that in each cell our solution is expressed as
+that in each cell our solution is expressed in the nodal form. It is convenient,
+however, for limiting purposes to express the solution in a modal representation:
 
@@
 \mathbf{U}_h (x,t) =
-\sum_{l=1}^{N}c_{l}(t)\,P_{l}(x)
+\sum_{l=1}^{N=k+1}c_{l}(t)\,P_{l}(x)
 \eeq
 
-\noindent where we have chosen to expand in the Legendre polynomials. We compare
-the weight $c_{1}$, which is proportional to the first derivative of the
-solution in the cell, to the neighboring cell averages by the following \\
+\noindent where $P_{l}(x)$ are the Legendre polynomials. These
+representations of the solution are equivalent by requiring weak equivalence
 
 \beq
-  \widetilde{c}_{1} = \text{minmod}(c_{1},
-    \beta_{\TVD}(c_{0}^{+} - c_{0}), \beta_{\TVD}(c_{0} - c_{0}^{-}))
+  \sum_{j=1}^{N}\int_{\bK} \left( \vect{U}_{j}(t)\, \ell_{j}(x) -
+    c_{j}(t)\, P_{j}(x) \right) \phi\, dx = 0 \quad \forall\,\phi \in \mathbb{V}^{k}.
+    \label{eq:NodalModal}
+\eeq
+
+\noindent Letting $\phi$ be the Lagrange interpolating polynomials $\ell_{i}$, we have
+
+\beq
+  \sum_{j=1}^{N}\int_{\bK}\ell_{i}(x)\, \ell_{j}(x)\, dx\, \vect{U}_{j}(t) =
+  \sum_{j=1}^{N}\int_{\bK}\ell_{i}(x)\, P_{j}(x)\, dx\, c_{j}(t).
+\eeq
+
+\noindent Now letting $ M_{ij} = \int_{\bK}\ell_{i}(x)\, \ell_{j}(x)\, dx$ be the mass matrix,
+$A_{ij} = \int_{\bK}\ell_{i}(x)\, P_{j}(x)\, dx$, $\bar{\vect{U}} = \{U_{1},...,U_{N}
+\}$ be the nodal coefficients, and $\bar{\vect{c}} = \{c_{1},...,c_{N}\}$
+be the modal coefficients, we see that the different representations are related by
+a linear transformation
+
+\beq
+  M \bar{\vect{U}} = A \bar{\vect{c}}
+\eeq
+To perform slope limiting, we compare
+the weight $c_{2}$, which is proportional to the first derivative of the
+solution in the cell, to the neighboring cell averages by the following
+
+\beq
+  \mathcal{M}\,\widetilde{c}_{2} = \text{minmod}(\mathcal{M}c_{2},
+    \beta_{\TVD}\mathcal{M}(c_{1}^{+} - c_{1}), \beta_{\TVD}\mathcal{M}(c_{1} - c_{1}^{-}))
   \label{eq:limiting}
@@
 
-\noindent where $\widetilde{c}_{1}$ is the limited weight. The superscripts $-/+$
+\noindent where $\widetilde{c}_{2}$ is the limited weight. The superscripts $-/+$
 indicate the cell averages of the neighboring cells to the  immediate left/right
-of the interface.
-Here, we use the minmod function defined as
+of the interface and $\mathcal{M}$ is a transformation matrix. For componentwise
+limiting, we let $\mathcal{M}$ be the identity matrix. Notice that
+
+\beq
+%  \begin{align}
+    \frac{1}{\Delta x}\int_{\bK}\vect{U}_{h}\,dx =
+      \frac{1}{\Delta x} \sum_{l=1}^{N}\int_{\bK}P_{l}(x)\,\cdot 1\, dx\, c_{l}
+      = \delta_{l1} c_{l} = c_{1}
+  \label{eq:cellAverage}
+%  \end{align}
+\eeq
+so $c_{1}$ is precisely the cell average. The minmod function in
+Eq.~\eqref{eq:limiting} is defined as
 
@@
 
-Limiting is applied to each component of the vector of conserved quantities
-separately and is generally referred to as `Componentwise limiting.`
 The parameter $\beta_{\TVD}$ takes values in the closed interval $\left[1,2\right]$
 and scales the strength of the limiting. A minimal $\beta_{\TVD}$ corresponds with
-the total variation diminishing scheme \textcolor{red}{reference?} but is more
-dissipative than the maximal $\beta_{\TVD}$ case, which is more flexible but allows for more oscillations.
+the total variation diminishing scheme but is more
+dissipative than the maximal $\beta_{\TVD}$ case, which is potentially more oscillatory.
 Increasing $\beta_{\TVD}$ puts more weight on the neighboring cell averages,
-making the minmod function more likely to set $ \widetilde{c}_{1} = c_{1}$.
+making the minmod function more likely to set $\widetilde{c}_{2} = c_{2}$, when
+no limiting is applied. If instead $\widetilde{c}_{2} \neq c_{2}$, the solution
+is truncated to first order
+
+\beq
+  \vect{U}_{h} \to \widetilde{\vect{U}}_{h} = c_{1}\, P_{1}(x) + \widetilde{c}_{2}\,P_{2}(x)
+\eeq
+and, by Eq.~\eqref{eq:minmod}, $\widetilde{c}_{2}$ can potentially be zero as well.
+Notice that, by Eq.~\eqref{eq:cellAverage}, the cell average is not altered by
+the limiting process, so the process is conservatory.
 
 In order to determine where slope limiting is necessary, we use the
-troubled cell indicator (TCI) \citep{fu:2017} to prevent excessive limiting:\footnote{Would this notation be altered for 1D?}
+troubled cell indicator (TCI) \citep{fu:2017} to prevent excessive limiting
 
@@
 
-\noindent where $G\in\vect{G}\subseteq\vect{U}$ and $\vect{G}$ consists of the
-first and final elements of $\vect{U}$. Here, the sum in the numerator and the
+\noindent where $G\in\vect{G}\subseteq\vect{U}$ and $\vect{G} = \left(\rho,E\right)$.
+Here, the sum in the numerator and the
 max in the denominator are taken over the neighboring elements sharing a
@@
 \label{sec:TimeInt}
+Eq.~\eqref{eq:semidiscreteDiscretized} presented a system of ODEs that must
+be evolved in time. We use the third-order strong stability-preserving
+Runge-Kutta (SSP-RK) scheme \citep{shu:1988}.
 The general $s$-stage Runge-Kutta time stepping algorithm, including
@@
   \small  \bar{\vect{U}}^{(i)}
-    = \Big\{\Lambda^{\TVD}\Big\{\sum_{j=0}^{i-1}\alpha_{ij}\,\bar{\vect{U}}^{(j)}+\beta_{ij}\,\Delta t\,\bar{\vect{F}}\big(\bar{\vect{U}}^{(j)}\big)\Big\}\Big\},
+    = \Lambda^{\TVD}\Big\{\sum_{j=0}^{i-1}\alpha_{ij}\,\bar{\vect{U}}^{(j)}+\beta_{ij}\,\Delta t\,\bar{\vect{F}}\big(\bar{\vect{U}}^{(j)}\big)\Big\},
     \label{eq:rkStages}
@@
 Above, the TVD limiter limiter preventing unphysical
-states is denoted by the operators $\Lambda^{\TVD}\{\}$. The SSP-RK3 coefficients $\alpha_{ij}$
+states is denoted by the operator $\Lambda^{\TVD}\{\}$. The SSP-RK3 coefficients $\alpha_{ij}$
 and $\beta_{ij}$ may be found in Table~2.1 in \citet{cockburn:2001a}. For each step
@@
 by the troubled cell indicator (where $I_{\bK} > C_{\TCI}$).
-This algorithm is subject to a timestep stability condition $\Delta t_{\TVD}$.
-Let $(h/|\lambda|)=\min_{i\in\{1,\ldots,d\}}(\Delta x^{i}/|\lambda^{i}|)$, where $|\lambda^{i}|$ is the largest absolute eigenvalue of the flux Jacobian.
-Then, assuming Cartesian coordinates (e.g., \cite{cockburn:2001a}):
+This algorithm is subject to a timestep stability condition $\Delta t_{\TVD}$
+where $|\lambda|$ is the magnitude of the largest eigenvalue of the flux Jacobian.
+Then (e.g., \cite{cockburn:2001a}):
 \begin{equation}
-  \Delta t_{\TVD} \le \f{1}{d}\,\f{(h/|\lambda|)}{(2k+1)} .
+  \Delta t_{\TVD} \le \f{1}{d}\,\f{(\Delta x/|\lambda|)}{(2k+1)}
   \label{eq:timestepRestrictions}
 \end{equation}
-For $d=k=2$, the effective Courant-Friedrichs-Lewy (CFL) factor is $0.1$.
+where d is the dimension. For $d=1,k=2$ (third-order scheme) the effective Courant-Friedrichs-Lewy (CFL) factor is $0.2$.
+In the numerical experiments presented in Section~\ref{sec:results} we use
+a more conservative timestep restriction of $\Delta t = 0.1 (\Delta x/|\lambda|)$.
 
@@
 as opposed to the conserved variables $\mathbf{U}_h$
-\citep[see, e.g.,][for a description]{cockburn:1998}. We begin by
+\citep[see, e.g.,][for a description]{cockburn:1998}.
+Inspired by theory for linear systems, we begin by
 rewriting Eq.~\eqref{eq:conservation} in the quasi-linear form as follows
@@
   \pderiv{\mathbf{U}}{t}
-  + \pderiv{\mathbf{U}}{\mathbf{x}} \pderiv{\mathbf{F(\mathbf{U})}}{\mathbf{U}}
+  + \pderiv{\mathbf{F(\mathbf{U})}}{\mathbf{U}}\pderiv{\mathbf{U}}{x}
   = 0.
@@
 \eeq
-Because the Euler equations for a systemer of hyperblic
+Because the Euler equations form a system of hyperblic
 partial differential equations \citep[see, e.g.,][]{leveque:1992}, we can decompose the
@@
 $\Lambda$ is a diagonal matrix containing the eigenvalues of the Jacobian.
+For hyperbolic systems, the eigenvalues are real and the eigenvectors form a
+complete set \citep[see e.g.,][]{leveque:1992}.
 At this point, we will introduce the characteristic variable
@@
 \eeq
-Solutions to these advection equations are far simpler to obtain than the
-previous equations. Limiting may then be applied to dampen oscillations in the
+Solutions to these now decoupled advection equations are far simpler to obtain than the
+previous equations. Recall in Eq.~\eqref{eq:minmod} that the limiting cell
+averages were transformed by a matrix $\mathcal{M}$. If we let
+$\mathcal{M}=\mathcal{R}^{-1}$ then limiting is performed on the characteristic variables.
+Limiting may then be applied to dampen oscillations in the
 solutions for the characteristic variables $\mathbf{w}$, which are then
 transformed back to the solution of the conserved variables $\mathbf{U}$
-\citep[see e.g.,][for a description]{cockburn:1998, schaal:2015a}. While this
+\citep[see e.g.,][for a description]{cockburn:1998, schaal:2015a}.
+While this
 process of characteristic limiting has been done for the ideal EOS
@@
 
-\section{Derivation of Jacobian and Eigenvalues for the 3D Nuclear Case}
+\section{Jacobian and Eigensystem for the Nuclear Case}
 \label{sec:eigen}
 
-We assume that pressure $P = P(\tau, \epsilon, n)$, where $\tau = \frac{1}{\rho}$.
+We assume that pressure $P = P(\tau, \epsilon, D_{e})$, where $\tau = \frac{1}{\rho}$.
 Let the vector of conserved variables be $\textbf{U} = \{\rho, m_1, m_2, m_3, E, D_{e}\}$,
@@
 \{m_{1}, m_{1}^{2}\tau + P, m_{1}m_{2}\tau, m_{1}m_{3}\tau,
-(E+P)m_{1}\tau, Dm_{1}\tau\}$. The Jacobian of the flux vector is given by
+(E+P)m_{1}\tau, D_{e}m_{1}\tau\}$. The Jacobian of the flux vector is given by
 
@@
 
-\noindent Where $H=(E+P)\tau$ is the specific enthalpy and $P_{x} = \partial_{x}P$\footnote{Maybe define these elsewhere to show what's constant}.
+\noindent Where $H=(E+P)\tau$ is the specific enthalpy and
+\beq
+%  \begin{align}
+    P_{\epsilon}  = \left(\frac{\partial{P}}{\partial{\epsilon}}\right)_{\tau,D_{e}}, \quad
+    P_{D_e}  = \left(\frac{\partial{P}}{\partial{D_{e}}}\right)_{\tau, \epsilon}, \quad
+    P_{\tau}  = \left(\frac{\partial{P}}{\partial{\tau}}\right)_{\epsilon, D_{e}}.
+%  \end{align}
+\eeq
 The eigenvalues of the flux Jacobian are given by the diagonal matrix
@@
 \end{align}
-where $c_{s} = \sqrt{\Gamma P \tau}$, for
+where $c_{s} = \sqrt{\Gamma P \tau}$, with
 $\Gamma = \left(\tau (P P_{\epsilon} - P_{\tau}) + P_{D_e} y_{e} \tau^{-1}\right) P^{-1}$, is
-the local sound speed. In the ideal EOS regime this reduces to the sound speeds given in \citet{colella:1985}.
+the local sound speed. In the case of the general EOS $P=P(\tau,\epsilon)$,
+this reduces to the sound speeds given in \citet{colella:1985}.
 The right eigenvectors are then given by the column vectors of the following matrix
@@
 \end{align*}
-where the follow definitions have been used:
+where the following definitions have been used:
 $h_{n} = \frac{c^2}{P_{\epsilon}\tau} + k $, $k = \frac{-y_{e} P_{D_{e}} \tau^{-1}
@@
 The left and right eigenvectors presented here are similar in form to those
-presented in \citet{schaal:2015a}, but are considerably messier due to the
+presented in \citet{schaal:2015a}, but are considerably more complicated due to the
 restriction of a nuclear EOS.

### 2019-05-06 paper/ms.tex
Submitted version

@@
     P_{D_e}  = \left(\frac{\partial{P}}{\partial{D_{e}}}\right)_{\tau, \epsilon}, \quad
-    P_{\tau}  = \left(\frac{\partial{P}}{\partial{\tau}}\right)_{\epsilon, D_{e}}.
+    P_{\tau}  = \left(\frac{\partial{P}}{\partial{\tau}}\right)_{\epsilon, D_{e}}
 %  \end{align}
 \eeq
+are the necessary thermodynamic derivatives. Expressions for these derivatives
+in terms of the table variables $\rho, T, \text{and}~y_{e}$ are given in Appendix~\ref{appendix:deriv}.
 The eigenvalues of the flux Jacobian are given by the diagonal matrix
@@
 presented in \citet{schaal:2015a}, but are considerably more complicated due to the
-restriction of a nuclear EOS.
+restriction of a nuclear EOS. Having an explicit form for these matrices is
+desirable as it eliminates the need to numerically diagonalize the flux
+Jacobian\footnote{These matrices and their derivations are available in Mathematica
+notebooks at \url{https://github.com/AstroBarker/dgHydro-nuc_charDecomp}}.
 
@@
 decomposition. Unless otherwise noted, we use the SFHo EOS of
-nuclear matter \citep{steiner:2013} now commonly used in high-fidelity CCSN simulations.
+nuclear matter \citep{steiner:2013} now commonly used in high-fidelity CCSN simulations
+due to its consistency with neutron star mass and radius observations.
 These tests serve to gauge the performance improvement of the
@@
 
-\subsection{Sod Shock Tube}
-We present the classic 1D Sod shock tube Riemann problem \citep{sod:1978} in
+\subsection{Riemann Problem}
+\label{sec:results:1}
+We present a 1D Riemann problem inspired by the classic 1D Sod shock tube Riemann problem \citep{sod:1978} in
 Cartesian coordinates. This is an ideal first test as it is a fairly simple
 scenario but contains both a contact discontinuity and a shock to test the
-efficiency of our limiting scheme. The computational domain is D = [-5,5] km
-with a discontinuity initially located at $x = 0$ km separating the left and right states
+efficiency of our limiting scheme, though due to the nature of our EOS, an analytic
+solution does not exist. The computational domain is D = [-5,5]~km
+with a discontinuity initially located at $x = 0$~km separating the left and right states
 
 \begin{align*}
-  \mathbf{U}_{L} &= (10^{12}~\text{g~cm}^{-3}, 0\,, 3.712*10^{32}~\text{ergs~cm}^{-3}, 0.4*10^{12})^T\,\,\, \\
-  \mathbf{U}_{R} &= (1.25*10^{11}~\text{g~cm}^{-3}, 0\, , 3.015*10^{31}~\text{ergs~cm}^{-3}, 0.375*10^{11})^T.
+  \mathbf{U}_{L} &= (10^{12}~\text{g~cm}^{-3}, 0\,, 2.703\text{x}10^{32}~\text{ergs~cm}^{-3}, 0.4\text{x}10^{12})^T\,\,\, \\
+  \mathbf{U}_{R} &= (1.25\text{x}10^{11}~\text{g~cm}^{-3}, 0\, , 2.822\text{x}10^{31}~\text{ergs~cm}^{-3}, 0.375\text{x}10^{11})^T.
 \end{align*}
-\footnote{\textbf{Fix e in the state vectors}}
-\noindent The test is run until $t = 0.025$ms with 100 elements
-using $C_{\TCI} = 0.2$ and $\beta_{\TVD} = 2.0$. Results are plotted in
-Figure~\ref{fig:SodSedovOptimal}. Using the characteristic limiting described
+
+\noindent The test is run until $t = 0.025$~ms with 100 elements
+using $C_{\TCI} = 0.2$ and $\beta_{\TVD} = 2.0$ with a third-order time
+integration SSP-RK3 scheme. Results for density, pressure, velocity, and electron fraction
+are plotted in the upper left, upper right, lower left, and lower right panels of
+Figure~\ref{fig:SodSedovOptimal}, respectively, compared to a reference run using
+10000 elements, third-order time integration, and first order spacial discretization.
+Using the characteristic limiting described
 above, the DG method captures the features of the solution,
-resolving well both the contact discontinuity and shock, without introducing
+resolving well the contact discontinuity, shock, and rarefraction wave
+located at about $x=2$~km, $x=4$~km, and from $x=-3$~km to $x=0$~km, respectively, without introducing
 noticeable oscillations in the numerical solutions near the discontinuities.
@@
   \centering
-  \caption{\label{fig:SodSedovOptimal} Numerical solution of Sod's problem using
-    100 elements compared with a reference solution using 10000 elements.}
+  \caption{\label{fig:SodSedovOptimal} Numerical solution of the Riemann problem using
+    100 elements and $C_{\TCI} = 0.2$ and $\beta_{\TVD} = 2.0$ with a third-order
+    time integration SSP-RK3 scheme for density (upper left), pressure (upper right),
+    velocity (lower left), and electron fraction (lower right)
+    compared with a reference solution (black) using 10000 elements.}
 \end{figure}
@@
   \includegraphics[width=18pc]{optimal_shock.png}
-  \caption{$xt$-plane of elements flagged for limiting by the troubled cell indicator.}
+  \caption{$xt$-plane of elements flagged for limiting by the troubled cell indicator
+  in the Riemann problem of Section~\ref{sec:results:1}.}
   \label{fig:shock}
@@
 $C_{\TCI}\in\left[0.0,0.2\right]$. For each $\left(\beta_{\TVD},C_{\TCI}\right)$ pair,
-we computed the relative error and total variation \textbf{\textcolor{red}{EQUATIONS}} in density and electron
-fraction at $t = 0.025$ms. Results are plotted in Figure~\ref{fig:landscape}.
+we computed the relative error and total variation \
+\beq
+  \varepsilon(\mathbf{U}) = \frac{1}{N}\sum_{\bK}\left\lvert\frac{\vect{U}_{h}^{\text{ref}}(x_{q}) - \vect{U}_{h}(x_{q})}{\vect{U}_{h}^{\text{ref}}(x_{q})}\right\rvert \quad\quad
+  TV = \sum_{k}\abs{ \bar{\vect{U}}_{k+1} - \bar{\vect{U}}_{k} }
+\eeq
+ in density and electron
+fraction at $t = 0.025$ms, where $\vect{U}_{h}^{\text{ref}} $ is the reference
+solution, $\bar{\vect{U}}_{k+1}$ is the cell average is cell $k$, and $x_{q}$ are
+numerical quadrature points.
+Results are plotted in Figure~\ref{fig:landscape}.
 We find that, as expected, increasing $\beta_{\TVD}$ tends to monotonically
@@
   \includegraphics[width=32pc, height=27.5pc]{./figures/error_tv_landscape.png}
-  \caption{\label{fig:landscape} Error - total variation landscapes for various
-  values of the limiting parameters $C_{TCI}$ and $\beta_{TVD}$.}
+  \caption{\label{fig:landscape} Error and total variation landscapes for various
+  values of the limiting parameters $C_{TCI}$ and $\beta_{TVD}$. The top row shows
+  total variation in density (left) and electron fraction (right). The bottom
+  row show relative error in density (left) and electron fraction (right).}
 \end{figure}
@@
 improvement from componentwise limiting. In Figure~\ref{fig:SodSedovCW} we plot the
-solution at $t = 0.025$ms using 100 elements employing both characteristic
-limiting (blue) and componentwise limiting (red) compared to a reference
+solution at $t = 0.025$~ms using 100 elements employing both characteristic
+limiting (blue) and componentwise limiting (red) compared to the reference
 solution computed with 10000 elements. We observe, for the componentwise limiting,
@@
 large oscillations at the left side of the contact discontinuity, up to
-$\pm 0.02$ in the cell averaged electron fraction. This makes characteristic
+$\pm 0.02$ in the cell averaged electron fraction. The solutions with
+characteristic limiting are very close to the reference solution, presenting
+a significant improvement over componentwise limiting.
+This makes characteristic
 limiting particularly appealing for CCSN simulations where the electron fraction
@@
   100 elements with characteristic limiting (blue) and componentwise limiting
-  (red) compared with a reference solution using 10000 elements.}
+  (red) for density (upper left), pressure (upper right),
+  velocity (lower left), and electron fraction (lower right)
+  compared to a reference solution (black) using 10000 elements.compared with a reference solution using 10000 elements.}
 \end{figure}
 
+\vspace{1cm}
 \subsection{High Density Regime}
-This test is included to verify the performance improvements of the characteristic
+This test is included to verify the performance improvements of characteristic
 limiting in a higher density regime. It is designed with the Sod problem in mind.
-The computational domain is $D = [-5,5]$ km
-with a discontinuity initially located at $x = 0$ km separating the left and right states
+The computational domain is $D = [-5,5]$~km
+with a discontinuity initially located at $x = 0$~km separating the left and right states
 
 \begin{align*}
-  \mathbf{U}_{L} &= (10^{13}~\text{g~cm}^{-3}, 0\,, 3.712*10^{32}~\text{ergs~cm}^{-3}, 0.15*10^{12})^T\,\,\, \\
-  \mathbf{U}_{R} &= (1.25*10^{12}~\text{g~cm}^{-3}, 0\, , 3.015*10^{31}~\text{ergs~cm}^{-3}, 0.169*10^{12})^T.
+  \mathbf{U}_{L} &= (10^{13}~\text{g~cm}^{-3}, 0\,, 3.712\text{x}10^{32}~\text{ergs~cm}^{-3}, 0.15\text{x}10^{12})^T\,\,\, \\
+  \mathbf{U}_{R} &= (1.25\text{x}10^{12}~\text{g~cm}^{-3}, 0\, , 3.015\text{x}10^{31}~\text{ergs~cm}^{-3}, 0.169\text{x}10^{12})^T.
 \end{align*}
 
-\noindent The test is run until $t = 0.05$ms with 100 elements
+\noindent The test is run until $t = 0.05$~ms with 100 elements
 using $C_{\TCI} = 0.1$ and $\beta_{\TVD} = 1.2$. We chose a lower $\beta_{\TVD}$
@@
 EOS table causing the test to fail, motivating the development of a
-positivity limiter compatible with a tabulated nuclear matter EOS.
+positivity limiter compatible with a tabulated nuclear matter EOS \citep{zhang:2010a}.
 Inital state values were chosen to be consistent with
-high-fidelity CCSN simulation code \chimera\, \citep{bruenn:2018} data.
+high-fidelity CCSN simulation code \chimera\, \citep{bruenn:2018} data by first
+choosing the left and right densities to be an order of magnitude larger than in
+Section~\ref{sec:results:1}, and then finding temperatures and electron fractions
+consistent with those densities from \chimera\, data, and computing the other
+state variables from the EOS table.
 Results are plotted in
@@
 10000 elements. Results for both tables are plotted in Figure~\ref{fig:SodSedovSFHoRes}.
-We find nearly no sensitivity to the table resolution, with the exception of
+We find no sensitivity to the table resolution, with the exception of
 minor differences in density to the left of the contact discontinuity.
@@
   Sod's problem using 100 elements with characteristic limiting and two
-  different EOS table resolutions.}
+  different EOS table resolutions (low: red, high: black) for density (upper left), pressure (upper right),
+  velocity (lower left), and electron fraction (lower right)
+  compared with a reference solution (black) using 10000 elements and the low resolution EOS table.}
 \end{figure}
@@
   Sod's problem using 100 elements with characteristic limiting and three
-  different EOS tables.}
+  different EOS tables for density (upper left), pressure (upper right),
+  velocity (lower left), and electron fraction (lower right).
+  }
 \end{figure}
@@
 Euler equations of gas dynamics in the {\bf t}oolkit for {\bf h}igh-{\bf or}der
-{\bf n}eutrino-r{\bf ad}iation hydr{\bf o}dynamics (\thornado). The results
-presented from a suite of 1D test problems, based around Sod's 1D shock
-tube problem, demonstrate the performance of the characteristic limiter
+{\bf n}eutrino-r{\bf ad}iation hydr{\bf o}dynamics (\thornado). We presented analytic
+forms for the diagonalizating matrices for the flux Jacobian of the Euler equations
+consistent with a tabulated nuclear EOS.
+The results presented from a suite of 1D test problems
+demonstrate the superior performance of the characteristic limiter
 compared to the componentwise limiter. Moreover, after performing a parameter
@@
 \thornado\, includes the development of a positivity limiter and extension of
-the characteristic limtier to handle relativistic and curvilinear problems.
+the characteristic limtier to handle multi-dimensional, relativistic, and curvilinear problems.
+Because of the superior performance of characteristic limiting, we hope that
+this work will improve the fidelity of CCSN simulations.
 
-\acknowledgements
-We thank people.
+\acknowledgments{
+Eirik Endeve and Anthony Mezzacappa acknowledge support from the NSF Gravitational Physics Program (NSF-GP 1505933 and 1806692).
+}
 
@@
 \section{Derivatives}
+\label{appendix:deriv}
 \input{derivatives}

### 2019-05-06 paper/ms.tex
Found an error.

@@
 the tests conducted in Section~\ref{sec:optimal} using three different EOS tables:
-the high resolution SFHo table from Section~\ref{sec:eosRes}, the SFHx table
-(\textbf{citation}), and the DD2 table (\textbf{citation}).
+the high resolution SFHo table from Section~\ref{sec:eosRes}, the SFHx table, and the DD2 table.
 As before, tests are computed using 100 elements until

### 2019-09-25 paper/ms.tex
Work on going multi-D
Pushed 3D stuff

@@
 The {\bf t}oolkit for {\bf h}igh-{\bf or}der {\bf n}eutrino-r{\bf ad}iation hydr{\bf o}dynamics
-(\thornado)\footnote{\url{https://github.com/endeve/thornado}}is being developed to simulate neutrino-radiation
+(\thornado)\citep{endeve:2019}\footnote{\url{https://github.com/endeve/thornado}}is being developed to simulate neutrino-radiation
 hydrodynamics in CCSNe and related applications in nuclear astrophysics.
@@
 
+% \subsection{Equation of State}
+
+% It is commonplace in the simulation of CCSNe to employ a tabulated EOS allowing for a wide 
+% range of physical conditions. The EOS will provide the thermodynamic quantities such as pressure, 
+% internal energy, and entropy as functions of the table variables density, temperature, and electron 
+% fraction. These quantities and their derivatives are crucial for the hydrodynamics, nuclear reactions, and 
+% neutrino transport as well as the characteristic limiting described in Section~\ref{sec:eigen}. To 
+% compute a thermodynamic quantity $\nu$, a thermodynamic grid in $\left( log\, \rho, log\, T, y_{e} \right)$ space 
+% is constructed, and then $log\, \nu$ may be computed by linear interpolation from the 
+% surrounding corners of the grid. It is then simple to determine the derivative of $\nu$ 
+% with respect to $\rho$, $T$, or $y_e$ by differentiating the linear interpolation 
+% equation with respect to the desired variable \citep[see e.g.,][for a more detailed description]{mezzacappa:1999, bruenn:2018}.
+
 \section{Jacobian and Eigensystem for the Nuclear Case}
@@
    v_{3} & 0 & 0 & 0 & 1 & v_{3} \\
-   h-c v_{1} & v_{2} & \beta & 0 & v_{3} & h+c v_{1} \\
+   h_{n}-c v_{1} & v_{2} & \beta & 0 & v_{3} & h_{n}+c v_{1} \\
    y_{e}  & 0 & 0 & \frac{\tau  \chi }{2 P_{D_{e}}} & 0 & y_{e}  \\
@@
 
+\section{3D}
+\label{appendix:3d}
+\input{3d}
+
 \bibliography{ms}

### 2020-07-10 paper/derivatives.tex
Updates to derivative appendix, prepping for param. study
First Draft?
fixed some bugs

@@
-Derivatives of pressure with respect to $\tau$, $\epsilon$, and n. \\
+We have an equation of state $P \,=\, P(\tau, \epsilon, D_{e})$ with $\tau \,=\, \rho^{-1}$, 
+$\epsilon \,=\, e/ \rho$, and $D_{e} = \rho \, Y_{e}$. On the other hand, 
+our tabulated equation of state constructs thermodynamic variables in terms of $\rho$, $T$, and $Y_{e}$. We 
+need to express the thermodynamic derivatives of pressure necessary for characteristic limiting in terms of the table variables.
+We start with the differential of pressure
+
+\begin{equation}
+ \label{eq:pressure_var1}
+  dP = (\partial_{\tau}P)_{\epsilon, D_{e}}d\tau + (\partial_{\epsilon}P)_{\tau, D_{e}}d\epsilon + (\partial_{D_{e}}P)_{\epsilon, \tau}dD_{e}.
+\end{equation}
+
+Similarly, we may express the differentials of $\tau$, $\epsilon$, and $D_{e}$ in terms of the table variables
+
+\begin{equation}
+\begin{split}
+  d\tau &= - \rho^{-2} d\rho\\
+  d\epsilon &= (\partial_{\rho}\epsilon)_{T,Y_{e}}d\rho + (\partial_{T}\epsilon)_{\rho,Y_{e}}dT + (\partial_{Y_{e}}\epsilon)_{T,\rho}dY_{e}\\
+  dD_{e} &= (\partial_{\rho}D_{e})_{T,Y_{e}}d\rho + (\partial_{T}D_{e})_{\epsilon,Y_{e}}dT + (\partial_{Y_{e}}D_{e})_{T,\epsilon}dY_{e}\\
+   & = Y_{e} d\rho + \rho dY_{e}
+\end{split}
+\end{equation}
+
+Inserting these differentials into Equation~\ref{eq:pressure_var1}, we find an expression for the pressure differential
+
+\begin{equation} \label{eq:pressure_dif}
+\begin{split}
+  dP = &\left[ -\rho^{2} (\partial_{\tau}P)_{\epsilon, D_{e}} + (\partial_{\rho}\epsilon)_{T,Y_{e}}\,(\partial_{\epsilon}P)_{\tau, D_{e}}
+    + Y_{e} (\partial_{D_{e}}P)_{\epsilon, \tau} \right] d\rho \\
+    & + (\partial_{T}\epsilon)_{\rho,Y_{e}} (\partial_{\epsilon}P)_{\tau, D_{e}} dT \\
+	& + \left[ (\partial_{Y_{e}}\epsilon)_{T,\rho} (\partial_{\epsilon}P)_{\tau, D_{e}} + \rho (\partial_{D_{e}}P)_{\epsilon, \tau}\right]dY_{e}   
+\end{split}
+\end{equation}
+
+On the other hand, we have the differential of pressure in terms of the table variables
+
+\begin{equation}
+ \label{eq:pressure_table}
+  dP = (\partial_{\rho}P)_{T, Y_{e}}d\rho + (\partial_{T}P)_{\rho, Y_{e}}dT + (\partial_{Y_{e}}P)_{T, \rho}dY_{e}.
+\end{equation}
+
+Comparing Equation~\ref{eq:pressure_dif} and Equation~\ref{eq:pressure_table}, we have the system of equations
+
+\begin{gather}
+  \begin{bmatrix} 
+ 	-\rho^{2} & (\partial_{\rho}\epsilon)_{T,Y_{e}} & Y_{e} \\ 
+ 	0 & (\partial_{T}\epsilon)_{\rho,Y_{e}} & 0 \\
+	 0 & (\partial_{Y_{e}}\epsilon)_{T,\rho} & \rho
+  \end{bmatrix}
+  \begin{bmatrix}
+	(\partial_{\tau}P)_{\epsilon, D_{e}} \\ (\partial_{\epsilon}P)_{\tau, D_{e}} \\ (\partial_{D_{e}}P)_{\epsilon, \tau}
+  \end{bmatrix}
+ =
+  \begin{bmatrix}
+	(\partial_{\rho}P)_{T, Y_{e}} \\ (\partial_{T}P)_{\rho, Y_{e}} \\ (\partial_{Y_{e}}P)_{T, \rho}
+  \end{bmatrix}
+\end{gather}
+
+Solving, with some simplifications, we find the derivatives of pressure with respect to $\tau$, $\epsilon$, and $D_{e}$
+in terms of the table variables $\rho$, $T$, $Y_{e}$. \\
+
 \begin{footnotesize}
 \begin{eqnarray}
-	\frac{\partial{P}}{\partial{\epsilon}} &=& \left(\frac{\partial{\epsilon}}{\partial{T}}\right)^{-1}\left(\frac{\partial{P}}{\partial{T}}\right) \\
-	\frac{\partial{P}}{\partial{n}} &=& \frac{m_B}{\rho} \left[ \frac{\partial{P}}{\partial{y_e}} -
-          \frac{\partial{\epsilon}}{\partial{y_e}}\left(\frac{\partial{\epsilon}}{\partial{T}}\right)^{-1}\left(\frac{\partial{P}}{\partial{T}}\right)\right]\\
-	\frac{\partial{P}}{\partial{\tau}} &=& -\tau^{-2} \left[ \frac{\partial{P}}{\partial{\rho}} - \frac{y_e}{\rho} \left( \frac{\partial{P}}{\partial{y_e}} -
-          \frac{\partial{\epsilon}}{\partial{y_e}} \left(\frac{\partial{\epsilon}}{\partial{T}}\right)^{-1}\left(\frac{\partial{P}}{\partial{T}}\right)\right) -
-          \frac{\partial{\epsilon}}{\partial{\rho}}\left(\frac{\partial{\epsilon}}{\partial{T}}\right)^{-1}\left(\frac{\partial{P}}{\partial{T}}\right)\right]
+	\left(\frac{\partial{P}}{\partial{\epsilon}}\right)_{\tau,D_{e}} &=& \left(\frac{\partial{\epsilon}}{\partial{T}}\right)^{-1}_{\rho, Y_{e}}\left(\frac{\partial{P}}{\partial{T}}\right)_{\rho, Y_{e}} \\
+	\left(\frac{\partial{P}}{\partial{D_{e}}}\right)_{\tau, \epsilon} &=& \tau \left[
+					\left(\frac{\partial{P}}{\partial{Y_{e}}}\right)_{\rho, T} -
+          \left(\frac{\partial{\epsilon}}{\partial{Y_{e}}}\right)_{\rho, T}
+					\left(\frac{\partial{P}}{\partial{\epsilon}}\right)_{\tau, D_{e}}\right]\\
+	\left(\frac{\partial{P}}{\partial{\tau}}\right)_{\epsilon, D_{e}} &=& \tau^{-2}
+					\left[ Y_{e}\left(\frac{\partial{P}}{\partial{D_{e}}}\right)_{\tau, \epsilon}
+					 + \left( \frac{\partial{\epsilon}}{\partial{\rho}} \right)_{Y_{e}, T}
+          \left(\frac{\partial{P}}{\partial{\epsilon}}\right)_{\tau, D_{e}} - 
+		  \left(\frac{\partial{P}}{\partial{\rho}}\right)_{T, Y_{e}} \right]
 \end{eqnarray}

### 2020-07-10 paper/ms.tex
fixed some bugs

--- 2019-05-06 paper/ms.tex
@@
 \received{\today}
-\shorttitle{Discontinuous Galerkin methods with a nuclear EOS}
+\shorttitle{Discontinuous Galerkin methods with a nuclear EoS}
 
@@
 We introduce the framework for the characteristic decomposition of the Euler equations
-with the inclusion of the nuclear EOS terms and
+with the inclusion of the nuclear EoS terms and
 present results from some initial tests using a third-order scheme.
@@
 in nature \citep[see e.g.,][]{blondin:2006, muller:2012, oconnor:2018b}. Alongside general
-reletavistic gravity, complex nuclear matter equations of state (EOS), and
+reletavistic gravity, microphysics through a complex nuclear matter equations of state (EoS), and
 neutrino transport, hydrodynamics must be accurately modelled,
-creating a very challenging and compelling problem. Hydrodynamic instabilities are critical
+creating a very challenging and compelling problem. These areas, constituting the 
+"four pillars" of CCSNe, each present a formidable theoretical and computational 
+problem in their own right and are deeply coupled in the context of CCSNe.
+Hydrodynamic instabilities are critical
 in aiding the explosion, with complex phenomena such as turbulence and
-convection playing key roles in the CCSN mechanism
-\citep{murphy:2011, murphy:2013, couch:2013, couch:2015a, radice:2016, mabanta:2018}.
-The accurate and efficient modelling of the supernova hydrodynamics is imperative if the
+convection playing key roles in driving the CCSN mechanism
+\citep{murphy:2011, murphy:2013, couch:2013, couch:2015a, radice:2016, couch:2020}. The 
+EoS provides many of the thermodynamic quantities (e.g., 
+pressure, internal energy, and entropy) that are necessary for the hydrodynamics 
+as well as nuclear reactions and neutrino transport.
+The accurate and efficient modelling of the supernova hydrodynamics and associated microphysics is imperative if the
 explosion is to be realistically modelled.
@@
 In this paper, we develop a chracteristic
-limiter for use with a general tabulated nuclear matter EOS in \thornado.
-In the case of the ideal EOS, the eigenvectors and eigenvalues which diagonalize
+limiter for use with a general tabulated nuclear matter EoS in \thornado.
+In the case of the ideal EoS, the eigenvectors and eigenvalues which diagonalize
 the flux Jacobian, which are necessary to transform to the characteristic variables,
 are known analytically. All of the thermodynamic derivatives have analytic forms.
-The inclusion of a tabulated nuclear EOS complicates matters, as derivatives
-must be appriximated by interpolating from the EOS table and the diagonalization
+The inclusion of a tabulated nuclear EoS complicates matters, as derivatives
+must be appriximated by interpolating from the EoS table and the diagonalization
 becomes non-trovial.
@@
 describe the Euler equations of gas dynamics including a conservation law
-appropriate for a nuclear EOS. Section~\ref{sec:DG} describes our
+appropriate for a nuclear EoS. Section~\ref{sec:DG} describes our
 numerical implementation of the DG method in the one dimensional (1D) case.
@@
 The non-relativistic Euler equations of gas dynamics
-\citep[see, e.g.,][for details in the case of an ideal EOS]{leveque:2002}
-in the absence of sources with a nuclear matter EOS are given by
+\citep[see, e.g.,][for details in the case of an ideal EoS]{leveque:2002}
+in the absence of sources with a nuclear matter EoS are given by
 the equations of conservation of mass,
@@
 where $\rho$ represents mass density, $\mathbf{v}$ the fluid velocity vector,
-$P$ the fluid pressure, $D_{e}=\rho y_e$ where $y_e$ is the electron fraction,
+$P$ the fluid pressure, $D_{e}=\rho Y_{e}$ where $Y_{e}$ is the electron fraction,
 $E=\epsilon \rho +\frac{1}{2}\rho v^2$ the total energy (internal plus kinetic),
@@
 \eqref{eq:massConservation}-\eqref{eq:electronConservation}. The inclusion of
-Equation \eqref{eq:electronConservation} is because we require a nuclear EOS.
-These equations are closed by a tabulated EOS where the pressure is given by a function of
-density, temperature $T$ and the electron fraction: $P = P(\rho, T, y_e)$.
+Equation \eqref{eq:electronConservation} is because we require a nuclear EoS.
+These equations are closed by a tabulated EoS where the pressure is given by a function of
+density, temperature $T$ and the electron fraction: $P = P(\rho, T, Y_{e})$.
 We may rewrite Equations \eqref{eq:massConservation}-\eqref{eq:electronConservation}
@@
 While this
-process of characteristic limiting has been done for the ideal EOS
+process of characteristic limiting has been done for the ideal EoS
 \citep[][]{cockburn:1998}, we want to extend this process to the case of a
-tabulated nuclear matter EOS.
+tabulated nuclear matter EoS.
+
+\section{Equation of State}
+\label{sec:eos}
+
+The EoS will provide the thermodynamic quantities such as pressure, 
+internal energy, and entropy as functions of the table variables density, temperature, and electron 
+fraction. These quantities and, as we will see, their derivatives are crucial for the hydrodynamics, nuclear reactions, and 
+neutrino transport as well as the characteristic limiting described in Section~\ref{sec:eigen}.
+
+\subsection{Tabulated EoS}
+
+For the conditions prevalent in stellar interiors, particularly in the high density regimes of stellar collapse, 
+a simple analytic form for the equation of state does not exist. For a CCSNe simulation, we seek to model a wide 
+range of conditions including matter at densities above nuclear saturation, the formation of a proto-neutron star, 
+and matter not in nuclear statistical equilibrium. These conditions correspond to temperatures ranging from 
+0.5 MeV or less up to around 100 MeV during black hole formation and densities convering many orders 
+of magnitude \citep[see, e.g.,][]{fischer:2011}. Constructing an equation of state for this wide array of conditions is no 
+small task, involving describing the many body nucleon-nucleon interactions and other relevant physics 
+\citep[for a detailed review, see][]{oertel:2017}.
+
+It is commonplace in the simulation of CCSNe to employ a tabulated EoS allowing for a wide 
+range of physical conditions. 
+To compute a thermodynamic quantity $\nu$, a thermodynamic grid in $\left( log\, \rho, log\, T, Y_{e} \right)$ space 
+is constructed, and then $log\, \nu$ may be computed by linear interpolation from the 
+surrounding corners of the grid. It is then simple to determine the derivative of $\nu$ 
+with respect to $\rho$, $T$, or $Y_{e}$ by differentiating the linear interpolation 
+equation with respect to the desired variable \citep[see e.g.,][for a more detailed description]{mezzacappa:1999, bruenn:2018}.
 
@@
 Let the vector of conserved variables be $\textbf{U} = \{\rho, m_1, m_2, m_3, E, D_{e}\}$,
-where $m_i = \rho v_i$ and $D_{e} = \rho y_{e}$.
+where $m_i = \rho v_i$ and $D_{e} = \rho Y_{e}$.
 The flux vector is $\textbf{F}(\textbf{U}) =
@@
 			v_{1}(-H - P_{\tau}\tau^{2} -P_{\epsilon}\tau(\epsilon - \frac{v^{2}}{2})) & H - P_{\epsilon}v_{1}^{2}\tau  & -P_{\epsilon}v_{1}v_{2}\tau & -P_{\epsilon}v_{1}v_{3}\tau  & v_{1}(1+P_{\epsilon}\tau) & v_{1} P_{D_{e}} \\
-			-v_{1} y_{e} & y_{e} & 0 & 0 & 0 & v_{1} \\
+			-v_{1} Y_{e} & Y_{e} & 0 & 0 & 0 & v_{1} \\
 		\end{array}
@@
 are the necessary thermodynamic derivatives. Expressions for these derivatives
-in terms of the table variables $\rho, T, \text{and}~y_{e}$ are given in Appendix~\ref{appendix:deriv}.
+in terms of the table variables $\rho, T, \text{and}~Y_{e}$ are given in Appendix~\ref{appendix:deriv}.
 The eigenvalues of the flux Jacobian are given by the diagonal matrix
@@
 where $c_{s} = \sqrt{\Gamma P \tau}$, with
-$\Gamma = \left(\tau (P P_{\epsilon} - P_{\tau}) + P_{D_e} y_{e} \tau^{-1}\right) P^{-1}$, is
-the local sound speed. In the case of the general EOS $P=P(\tau,\epsilon)$,
+$\Gamma = \left(\tau (P P_{\epsilon} - P_{\tau}) + P_{D_e} Y_{e} \tau^{-1}\right) P^{-1}$, is
+the local sound speed. In the case of the general EoS $P=P(\tau,\epsilon)$,
 this reduces to the sound speeds given in \citet{colella:1985}.
@@
    v_{3} & 0 & 0 & 0 & 1 & v_{3} \\
-   h-c v_{1} & v_{2} & \beta & 0 & v_{3} & h+c v_{1} \\
-   y_{e}  & 0 & 0 & \frac{\tau  \chi }{2 P_{D_{e}}} & 0 & y_{e}  \\
+   h_{n}-c v_{1} & v_{2} & \beta & 0 & v_{3} & h_{n}+c v_{1} \\
+   Y_{e}  & 0 & 0 & \frac{\tau  \chi }{2 P_{D_{e}}} & 0 & Y_{e}  \\
   \end{array}
@@
 where the following definitions have been used:
-$h_{n} = \frac{c^2}{P_{\epsilon}\tau} + k $, $k = \frac{-y_{e} P_{D_{e}} \tau^{-1}
+$h_{n} = \frac{c^2}{P_{\epsilon}\tau} + k $, $k = \frac{-Y_{e} P_{D_{e}} \tau^{-1}
 + P_{\epsilon} (\frac{1}{2}v^2 + \epsilon) + P_{\tau}\tau}{P_{\epsilon}}$,
@@
       \\
-   -\frac{y_{e} P_{D_{e}} \omega }{\chi \tau } & \frac{2 y_{e} P_{D_{e}} \phi_{1} }{\chi \tau } & \frac{2 y_{e} P_{D_{e}}
-     \phi_{2} }{\chi \tau} & \frac{2 y_{e} P_{D_{e}} \phi_{3} }{\chi \tau} & -\frac{2 y_{e}
-     P_{D_{e}} P_{\epsilon} }{\chi } & \frac{2 P_{D_{e}} \left(c^2-y_{e} P_{D_{e}} \right)}{\tau  \chi }
+   -\frac{Y_{e} P_{D_{e}} \omega }{\chi \tau } & \frac{2 Y_{e} P_{D_{e}} \phi_{1} }{\chi \tau } & \frac{2 Y_{e} P_{D_{e}}
+     \phi_{2} }{\chi \tau} & \frac{2 Y_{e} P_{D_{e}} \phi_{3} }{\chi \tau} & -\frac{2 Y_{e}
+     P_{D_{e}} P_{\epsilon} }{\chi } & \frac{2 P_{D_{e}} \left(c^2-Y_{e} P_{D_{e}} \right)}{\tau  \chi }
       \\
@@
 $\omega = \tau\, (P_{\epsilon}\,(v^2 - 2\epsilon) - 2\,P_{\tau}\,\tau)$, and
-$\alpha = 2 y_{e} P_{D_{e}} - \tau \chi$.
+$\alpha = 2 Y_{e} P_{D_{e}} - \tau \chi$.
 The left and right eigenvectors presented here are similar in form to those
 presented in \citet{schaal:2015a}, but are considerably more complicated due to the
-restriction of a nuclear EOS. Having an explicit form for these matrices is
+restriction of a nuclear EoS. Having an explicit form for these matrices is
 desirable as it eliminates the need to numerically diagonalize the flux
@@
 method implemented in \thornado\, leveraging use of the characteristic
-decomposition. Unless otherwise noted, we use the SFHo EOS of
+decomposition. Unless otherwise noted, we use the SFHo EoS of
 nuclear matter \citep{steiner:2013} now commonly used in high-fidelity CCSN simulations
@@
 scenario but contains both a contact discontinuity and a shock to test the
-efficiency of our limiting scheme, though due to the nature of our EOS, an analytic
+efficiency of our limiting scheme, though due to the nature of our EoS, an analytic
 solution does not exist. The computational domain is D = [-5,5]~km
@@
 %\clearpage
-\begin{wrapfigure}{L}{2.8in}
-  \includegraphics[width=18pc]{optimal_shock.png}
+\begin{wrapfigure}{L}{0.35\textwidth}
+  \includegraphics[width=0.35\textwidth]{optimal_shock.png}
   \caption{$xt$-plane of elements flagged for limiting by the troubled cell indicator
@@
 plays a critical role in the explosion dynamics.
+\vspace{3cm}
 
@@
   velocity (lower left), and electron fraction (lower right)
-  compared to a reference solution (black) using 10000 elements.compared with a reference solution using 10000 elements.}
+  compared to a reference solution (black) using 10000 elements.}
 \end{figure}
-
+\clearpage
 \vspace{1cm}
@@
 \begin{align*}
-  \mathbf{U}_{L} &= (10^{13}~\text{g~cm}^{-3}, 0\,, 3.712\text{x}10^{32}~\text{ergs~cm}^{-3}, 0.15\text{x}10^{12})^T\,\,\, \\
-  \mathbf{U}_{R} &= (1.25\text{x}10^{12}~\text{g~cm}^{-3}, 0\, , 3.015\text{x}10^{31}~\text{ergs~cm}^{-3}, 0.169\text{x}10^{12})^T.
+  \mathbf{U}_{L} &= (10^{13}~\text{g~cm}^{-3}, 0\,, 3.712\text{x}10^{32}~\text{ergs~cm}^{-3}, 0.15\text{x}10^{13})^T\,\,\, \\
+  \mathbf{U}_{R} &= (1.25\text{x}10^{12}~\text{g~cm}^{-3}, 0\, , 3.015\text{x}10^{31}~\text{ergs~cm}^{-3}, 0.169\text{x}10^{13})^T.
 \end{align*}
 
-\noindent The test is run until $t = 0.05$~ms with 100 elements
+\noindent The test is run until $t = 0.075$~ms with 400 elements
 using $C_{\TCI} = 0.1$ and $\beta_{\TVD} = 1.2$. We chose a lower $\beta_{\TVD}$
 here, as larger values resulted in unphysical internal energies outside of the
-EOS table causing the test to fail, motivating the development of a
-positivity limiter compatible with a tabulated nuclear matter EOS \citep{zhang:2010a}.
+EoS table causing the test to fail, motivating the development of a
+positivity limiter compatible with a tabulated nuclear matter EoS \citep{zhang:2010a}.
 Inital state values were chosen to be consistent with
@@
 consistent with those densities from \chimera\, data, and computing the other
-state variables from the EOS table.
+state variables from the EoS table.
 Results are plotted in
-Figure~\ref{fig:SodSedovHighDens}. Our method captures the main features
-of the solution, but introduces more oscillations than in the lower density
-case, even when using the characteristic limiter.
-We notice an undershooting of the right side
-of the contact discontinuity in the electron fraction not previously present
-warranting further investigation.
+Figure~\ref{fig:SodSedovHighDens}. Our method captures the solution well even 
+in these high density regimes relevant for CCSNe without introducing 
+any spurious oscillations.
 
@@
   \centering
-  \includegraphics[width=36pc]{./figures/highDens.png}
+  \includegraphics[width=32pc]{./figures/highDens.png}
   \caption{\label{fig:SodSedovHighDens} Results for the high density Sod-like problem
@@
 \end{figure}
-
-\subsection{EOS Resolution Dependence}
-\label{sec:eosRes}
-In order to test the sensitivity of the limiter on the EOS table resolution,
+\FloatBarrier
+\subsection{EoS Resolution Dependence}
+\label{sec:EoSRes}
+In order to test the sensitivity of the limiter on the EoS table resolution,
 we repeated the tests in Section~\ref{sec:optimal} with a higher resolution
-SFHo EOS table. As before, tests are computed using 100 elements until
+SFHo EoS table. As before, tests are computed using 100 elements until
 $t = 0.025$ms with characteristic limiting and the optimal limiting parameters determined in
@@
   \centering
-  \includegraphics[width=36pc]{./figures/eos_res.png}
+  \includegraphics[width=36pc]{./figures/EoS_res.png}
   \caption{\label{fig:SodSedovSFHoRes} Comparison of numerical solutions to
   Sod's problem using 100 elements with characteristic limiting and two
-  different EOS table resolutions (low: red, high: black) for density (upper left), pressure (upper right),
+  different EoS table resolutions (low: red, high: black) for density (upper left), pressure (upper right),
   velocity (lower left), and electron fraction (lower right)
-  compared with a reference solution (black) using 10000 elements and the low resolution EOS table.}
+  compared with a reference solution (black) using 10000 elements and the low resolution EoS table.}
 \end{figure}
+\FloatBarrier
 
-\subsection{EOS Table Sensitivity}
-In order to probe the sensitivity of the method to the EOS table, we repeated
-the tests conducted in Section~\ref{sec:optimal} using three different EOS tables:
-the high resolution SFHo table from Section~\ref{sec:eosRes}, the SFHx table, and the DD2 table.
+\subsection{EoS Table Sensitivity}
+In order to probe the sensitivity of the method to the EoS table, we repeated
+the tests conducted in Section~\ref{sec:optimal} using three different EoS tables:
+the high resolution SFHo table from Section~\ref{sec:EoSRes}, the SFHx table, and the DD2 table.
 As before, tests are computed using 100 elements until
 $t = 0.025$ms with characteristic limiting and the optimal limiting parameters determined in
-Section~\ref{sec:param}. We find nearly no sensitivity to the EOS table used in
+Section~\ref{sec:param}. We find nearly no sensitivity to the EoS table used in
 the low density regime, with the exception of very small variations in density
 to the left of the contact discontinuity. Future work includes testing the
-sensitivity to the EOS table in higher density regimes.
+sensitivity to the EoS table in higher density regimes.
 \begin{figure}[h!]
   \centering
-  \includegraphics[width=36pc]{./figures/eos_all.png}
-  \caption{\label{fig:SodSedovEOS} Comparison of numerical solutions to
+  \includegraphics[width=36pc]{./figures/EoS_all.png}
+  \caption{\label{fig:SodSedovEoS} Comparison of numerical solutions to
   Sod's problem using 100 elements with characteristic limiting and three
-  different EOS tables for density (upper left), pressure (upper right),
+  different EoS tables for density (upper left), pressure (upper right),
   velocity (lower left), and electron fraction (lower right).
@@
 \end{figure}
-%\FloatBarrier
+\FloatBarrier
 
@@
 forms for the diagonalizating matrices for the flux Jacobian of the Euler equations
-consistent with a tabulated nuclear EOS.
+consistent with a tabulated nuclear EoS.
 The results presented from a suite of 1D test problems
@@
 tend to be problem dependent, we hope to investigate if such an optimal
-parameter exists for CCSN applications. Tests with various EOS tables and
+parameter exists for CCSN applications. Tests with various EoS tables and
 table resolutions showed little to no sensitivity to the table or its
